### 2018-11-05 PAComp/comp.tex
minutia added

\documentclass{article}

\usepackage{aaai}
\usepackage{xcolor}

\begin{document}

\title{Statistical comparisons over species distribution modeling methods using presence-only data: MaxEnt vs. MaxLike}
\author{Shayan Amani, Marek Petrik}
\maketitle\

\section*{Abstract}
[Needs to be filled wisely!]

\section*{Introduction}
These two methods have distinct approach in distribution modeling. MaxLike estimates probability of occurrence, while MaxEnt estimates a relative suitability index. According to Fitzpatrick et al. (2013) there is no apparent correlation or proximity between MaxEnt’s indices and probability of occurrence. By default, the MaxEnt algorithm assumes a baseline species prevalence of 0.5 \cite{Phillips2008}. Therefore, MaxEnt assigns occurrence probability of 0.5 to most occurrence locations \cite{FitzpatrickMC;GotelliNJ;Ellison2013}.

MaxLike assigns substantially higher probabilities to locations with recorded presences … than did any of the implementations of MaxEnt \cite{FitzpatrickMC;GotelliNJ;Ellison2013}.

\subsection*{Occurrence Probability}
As Fithian et al. (2013) discussed occurrence probability can varies with altering quadrat size in the surveyed land area or other sampling parameters. Survey path length -a factor of sampling scheme- acts as a predictor in different species distribution models in a wide variety of sampling units which introduces a level of vagueness to the way of interpreting the models' resulted maps \cite{Fithian2013}.

\section*{MaxEnt}

\subsection*{Pros and Cons}
It is problematic that MaxEnt rarely predicts any areas with a high probability of occurrence ($p > 0.80$) and typically generates a relatively narrow distribution of occurrence probabilities of mean $p \approx 0.5$ for recorded presences. These probabilities depend on the assumed value of species prevalence (MaxEnt $default = 0.5$) \cite{FitzpatrickMC;GotelliNJ;Ellison2013}

\subsection*{Concepts and Methodology}
In MaxEnt literature true distribution of species $\pi$ in the study site $x$ is the probability distribution $\pi (x)$ which assigns non-negative values to each area $x$ with a total of one. Model constraints (features) are nothing more than functions of environmental \textcolor{red}{covariates}. In order to elude model under-specification which is most likely to happen when applying a number of constraints to the model we pick the feature having maximum entropy meaning most unconstrained one among the whole set. Applying Bayes’ rule to the conditional probability $P(x|y=1)$ ---the probability of being at $x$, given that the species is present--- proceeds as follows:

\begin{equation} \label{eq: Bayes' rule on MaxEnt's conditional probability}
P ( y = 1 |  x )= \frac {P ( x | y = 1 )  P ( y = | 1 ) }{  P (  x )} =  \pi ( x ) P ( y = 1 ) | X |
\end{equation}

Equation \ref{eq: Bayes' rule on MaxEnt's conditional probability} exhibits that $\pi$ is proportional to probability of presence. However, if we have only occurrence data, we cannot determine the species’ prevalence \cite{Phillips2006}; \cite{Ward2009}. Therefore, instead of estimating $P(x|y=1)$ directly, we estimate the distribution $\pi$. According to Section 2 of Dudík et al. (2004) the Maxent distribution belongs to the family of Gibbs distributions derived from the set of features. Gibbs distributions are exponential distributions:

\begin{equation} \label{eq:Gibbs distribution}
q_\lambda (x) = \frac{ \mathrm{exp} (\Sigma_{j=1}^{n} \lambda_j f_j(x))} {Z_\lambda}
\end{equation}

where $\lambda$ is vector of feature weights, $Z_\lambda$ is normalization constant to guarantee that all probabilities $q_\lambda (x)$ sum to one over the study region.

\section*{MaxLike}

\subsection*{Pros and Cons}
In contrast, MaxLike usually generates a broader range of occurrence probabilities, with generally higher occurrence probabilities \cite{Phillips2006}.

\section*{Inhomogeneous Poisson Process}
IPP, a relatively simple model which is defined by the intensity function of each point $Z$ in the area domain $D$ as:

\begin{equation} \label{eq:IPP intensity function}
\lambda : D \longrightarrow [0,\infty].
\end{equation}

\section*{Likelihood}

[To be completed soon!]

\bibliographystyle{aaai}
\bibliography{library}

\end{document}

### 2018-11-12 AML-RL/DiscountedReward/solutions_policyiteration_lin.py
added codebase and reinforcement learning AML course materials

@@ -1,0 +1,122 @@
+import os    # module that handles OS features
+import sys   # sys module needed to access command line information
+import re
+import argparse
+import numpy as np
+
+parser = argparse.ArgumentParser(description = 'Policy Iteration')
+parser.add_argument('csv_file')
+parser.add_argument('-o', default='out')
+
+args = parser.parse_args()
+
+csv_file = args.csv_file
+out_file = args.o
+
+idstatefrom = []
+idaction = []
+idstateto = []
+probability = []
+reward = []
+states = {}
+policy_a = []
+V = []
+
+GAMMA = 0.75
+
+'''
+Read file and properly setup stucture for policy iteration
+'''
+def read_csv_file():
+    global V
+    with open(csv_file) as infile:
+        i = 0
+        for line in infile:
+            if i == 0: 
+                i += 1
+                continue
+            columns = line.split(',')
+            state = int(columns[0])
+            action = int(columns[1])
+            if state in states:
+                if action in  states[state]:
+                    states[state][action].append(line)
+                else:
+                    states[state][action] = []
+                    states[state][action].append(line)
+            else:
+                states[state] = {}
+                states[state][action] = []
+                states[state][action].append(line)
+                V.append(0)
+                policy_a.append(0)
+
+            line = line.rstrip(os.linesep)
+            # print(line)
+    infile.close()
+
+'''
+Perform policy iteration
+'''
+def evaluate():
+    global V
+
+    i = 0
+    print('idstate,idaction,bestV')
+    policy_prev = [-1,-1,-1,-1]
+    while (policy_prev != policy_a):
+        print('Iteration ' + str(i))
+        policy_prev = policy_a.copy()
+        mV = np.matrix(V)
+        mV = mV.transpose()
+        
+        for state, state_value in states.items():
+            pol_best = 0            
+            for action, action_value in state_value.items():
+
+                T = []
+                R = []
+                
+                for a in action_value:
+                    columns = a.split(',')
+                    T.append(float(columns[3]))
+                    R.append(float(columns[4]))
+                
+                mT = np.matrix(T)
+                mR = np.matrix(R)
+                mR = mR.transpose()
+                
+                mB = np.add(mR, GAMMA*(mV))
+
+                improved = np.matmul(mT, mB)
+                improvedf = float(improved.item(0))
+
+                if  improvedf > pol_best:
[…]

### 2018-11-12 AML-RL/InheritedFarm.py

@@ -1,0 +1,34 @@
+import numpy as np
+
+offer = 500
+n = 10000
+
+print("sell farm", offer*1.02**(n-1))
+
+v = 0
+for i in range(n):
+    v = 1.02 * v + 100
+
+print("keep farm", v)
+
+print('**********')
+
+offer = 500
+n = 10000
+
+print("sell farm", offer)
+
+v = 0
+for i in range(n):
+    v = (1/1.02) * v + 100
+
+print("keep farm", v)
+
+#################################################
+
+gamma = 1/1.02
+
+P = np.array([[0.8, 0.2], [0.4, 0.6]])
+
+v = np.linalg.solve(np.eye(2) - gamma * P, [200, -20])
+print("value function", v)

### 2018-11-12 codebase/LogReg.py

@@ -1,0 +1,48 @@
+import sys
+import glob
+
+import pystan
+import pandas as pd
+import numpy as np
+
+import matplotlib.pyplot as plt
+# %matplotlib inline
+# %precision 4
+plt.style.use('ggplot')
+
+df = pd.read_csv('HtWt.csv')
+df.head()  # printing first 5 rows
+
+log_reg_code = """
+data {
+    int<lower=0> n;
+    int male[n];
+    real weight[n];
+    real height[n];
+}
+transformed data {}
+parameters {
+    real a;
+    real b;
+    real c;
+}
+transformed parameters {}
+model {
+    a ~ normal(0, 10);
+    b ~ normal(0, 10);
+    c ~ normal(0, 10);
+    for(i in 1:n) {
+        male[i] ~ bernoulli(inv_logit(a*weight[i] + b*height[i] + c));
+  }
+}
+generated quantities {}
+"""
+
+log_reg_dat = {
+             'n': len(df),
+             'male': df.male,
+             'height': df.height,
+             'weight': df.weight
+            }
+
+fit = pystan.stan(model_code=log_reg_code, data=log_reg_dat, iter=2000, chains=1)

### 2018-11-12 codebase/StanPg.py

@@ -1,0 +1,57 @@
+import pystan
+import numpy as np
+import matplotlib.pyplot as plt
+
+stan_code_model = """
+data {
+    int ns;  // num of samples
+    int nf;  // num of features
+    matrix[ns,nf] phi;  // feature matrix
+    vector[ns] y;  // response
+}
+parameters {
+    vector[nf] beta;
+    real<lower=0> epsilon; 
+}
+model {
+    y ~ normal(phi * beta, epsilon);
+}
+"""
+
+
+def distro():
+    # stan_data = {'num_samples': 30, 'y': d}
+
+    randNumGen = np.random.RandomState(7)
+    n_samp = 10
+    n_feat = 4
+
+    stan_model = pystan.StanModel(model_code=stan_code_model)
+
+    features = randNumGen.normal(0, 1, size=(n_samp, n_feat))
+    response = features @ np.array([-1, 0.4, 0, 0]) + randNumGen.normal(scale=0.2, size=n_samp)
+
+    stan_fit = stan_model.sampling(data={'ns': n_samp,
+                                         'nf': n_feat,
+                                         'phi': features,
+                                         'y': response},
+                                   iter=1000, chains=4, warmup=500, n_jobs=-1, seed=7)
+
+    chains = stan_fit.extract()
+
+    chains['beta'].shape
+
+    ft_plot =  stan_fit.plot()
+    stan_fit.plot()
+
+
+    lin.plot(chains['beta'][:,0])
+
+
+
+def main():
+    distro()
+
+
+if __name__ == '__name__':
+    main()

### 2018-11-12 codebase/Weighted_L1_Experiments.py

@@ -1,0 +1,89 @@
+import pandas as pd
+import random
+import math
+import numpy as np
+import matplotlib.pyplot as plt
+
+from sklearn import model_selection
+from sklearn.linear_model import LogisticRegression
+from sklearn import datasets
+
+import pystan
+
+### cd /home/reazul/PhD_Research/MaxEnt_PointProcess/Codes/Experiments
+df = pd.read_csv("datasets/barberry_sim_NewEngland.csv")
+num_sub_samples = 200
+sub_samples = df.sample(n=num_sub_samples)
+
+abundance_samples = pd.DataFrame()
+for index, row in sub_samples.iterrows():
+    abundance = max(row['N'], 1)  # sub_samples.loc[index, 'N']
+    if abundance > 1:
+        sub_samples.loc[index, 'N'] = 1
+    abundance_samples = abundance_samples.append([sub_samples.loc[index]] * 1)  # int(abundance)
+
+X = abundance_samples[['bio10_5', 'bio10_prMay']].values
+Y = abundance_samples['N'].values
+
+
+### Bayesian Logistic Regression model using Stan
+def fit_bayesian_logistic_regression(X, y):
+    stan_dat = {'N': len(X), 'detected': y.astype(int), 'bio10_5': X[:, 0], 'bio10_prMay': X[:, 1]}
+
+    sm = pystan.StanModel(file='model_logistic.stan')  # pystan.StanModel(model_code=m1)
+    # fit = sm.sampling(data=stan_dat, iter=20, chains=4, refresh=0)
+    fit = sm.sampling(data=stan_dat, iter=180, chains=3, refresh=0)
+
+    mean_alpha, std_alpha = np.mean(fit.extract()['alpha']), np.std(fit.extract()['alpha'])
+    mean_beta_bio10_5, std_beta_bio10_5 = np.mean(fit.extract()['beta_bio10_5']), np.std(fit.extract()['beta_bio10_5'])
+    mean_beta_bio10_prMay, std_beta_bio10_prMay = np.mean(fit.extract()['beta_bio10_prMay']), np.std(
+        fit.extract()['beta_bio10_prMay'])
+
+    return mean_alpha, mean_beta_bio10_5, mean_beta_bio10_prMay, std_alpha, std_beta_bio10_5, std_beta_bio10_prMay
+
+
+def inv_logit(p):
+    # with precision(100000):
+    return round(math.exp(p) / (1 + math.exp(p)))
+
[…]

### 2018-11-12 presentation/PAC Optimal MDP Planning/main.tex
presentations added

% https://github.com/geriom/beamercuerna

\documentclass{beamer}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\usetheme{Cuerna}
\usecolortheme{default}
% default, bluesimplex, lettuce, brick

\title{TITLE}
\author{Shayan Amani}

\date{DATE 2018}
\institute{Department of Computer Science, University of New Hampshire}

\begin{document}

  \begin{frame}
    \titlepage
  \end{frame}

\begin{frame}
    \frametitle{Pneumonoultramicroscopicsilicovolcanoconiosis}
    \framesubtitle{SUB}

  \end{frame}

	%\includegraphics[scale=0.45]{img.jpg}

\end{document}

### 2018-11-12 presentation/Summer Report/main.tex
presentations added

\documentclass{beamer}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\usetheme{Cuerna}
\usecolortheme{default}

\title{Summer Work - A Brief Report}
\author{Shayan Amani\\}

\date{Summer 2018}
\institute{Department of Computer Science, University of New Hampshire}

\begin{document}

  \begin{frame}
    \titlepage
  \end{frame}

\begin{frame}
    \frametitle{Pneumonoultramicroscopicsilicovolcanoconiosis}
    \framesubtitle{Goals}
    Discussed during weekly meetings to allocate time to:
    \begin{itemize}
    \item familiarity with Species Distribution Models (SDM)
    \item examining Convexity of those models
    \item learning tools such as Convex Optimization
  \end{itemize}
  \end{frame}

  \begin{frame}
  	\frametitle{Accomplishments}
  \end{frame}
  \begin{itemize}
    \item finding out the methods like MaxLike and MaxEnt and some hands on experience with them
    \item 
  \end{itemize}
  \begin{frame}
  	\frametitle{Rec. Activities}
    Daily:
    \begin{itemize}
    \item push-ups x20
    \item 8 min abs workout
    \item summer fruits
    \end{itemize}
    Occasionally:
    \begin{itemize}
    \item White Mountains group hiking
    \item New Castle's beach on 4th of July
    \end{itemize}
	\includegraphics[scale=0.45]{img.jpg}
  \end{frame}

\end{document}

### 2018-11-15 presentation/PAC Optimal MDP Planning/main.tex
Pareto and Dirichlet distributions added to minutia
presentation updated: PAC optimal MDP planning

@@
-% https://github.com/geriom/beamercuerna
+% Golden Spiral can be approximated with Golden ratio or Fibonacci sequence.
+% Golden Ratio: a+b/a = a/b
 
-\documentclass{beamer}
+\documentclass[11pt]{beamer}
 \usepackage[utf8]{inputenc}
 \usepackage[T1]{fontenc}
+\usepackage{ulem}
 
 \usetheme{Cuerna}
-\usecolortheme{default}
+\usecolortheme{bluesimplex}
 % default, bluesimplex, lettuce, brick
 
-\title{TITLE}
+\addtobeamertemplate{navigation symbols}{}{%
+    \usebeamerfont{footline}%
+    \usebeamercolor[fg]{footline}%
+    \hspace{1em}%
+      \insertframenumber/\inserttotalframenumber  
+}
+
+\title{PAC Optimal MDP Planning}
 \author{Shayan Amani}
 
-\date{DATE 2018}
+\date{fall 2018}
 \institute{Department of Computer Science, University of New Hampshire}
@@
 \begin{frame}
-    \frametitle{Pneumonoultramicroscopicsilicovolcanoconiosis}
-    \framesubtitle{SUB}
+\begin{center}
+    \textbf{PAC Optimal MDP Planning with Application to Invasive Species Management}
+\\\\
+Majid Alkaee Taleghan, Thomas G. Dietterich, Mark Crowley, Kim Hall, H. Jo Albers
+\end{center}
+\end{frame}
+
+\begin{frame}
+\frametitle{The Experiment}
+\textbf{Tamarisk}
+\begin{tabular}{c c}
+    \includegraphics[scale=0.2]{tamarisk.jpg} &   \includegraphics[scale=0.4]{tamarisk1.jpg} \\
+    Source: Getty Images & Source: National Park Service
+\end{tabular}
+\end{frame}
+
+\begin{frame}{Goals}
+    \begin{itemize}
+        \item Reducing simulator calls (expensive to run).
+        \item Terminating algorithm once a PAC solution is found.
+        \item Minimizing the infinite horizon discounted cumulative cost. In other words, maximizing cumulative reward (i.e. in the context of infinite horizon and discounted reward MDP) 
+    \end{itemize}
+\end{frame}
+
+\begin{frame}
+    \frametitle{Inspiration}
+    \framesubtitle{(chronologically sorted)}
+Previous works which influenced the current paper:
+\begin{enumerate}
+    \item Fiechter's PAC-RL algorithm [1994]
+    \item Model-Based Action Elimination algorithm (MBAE) [2002, 2006]
+    \item Model-Based Interval Estimation algorithm (MBIE) [2008]
+\end{enumerate}
+\end{frame}
+
+\begin{frame}{Contributions}
+The improvements can be identified in different spans as follows
+    \begin{itemize}
+        \item Selective exploration. Fiechter's algorithm visits all trajectories.
+        \item The occupancy measure as a guarantee for exploring reachable states from the start state.
+        \item Accuracy bound width as the termination condition.
+        \item Two more sophisticated intervals over Hoeffding bound. Weissman et al. [2008] (page 3884) and Good-Turing [1953] (page 3886).
+
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Primary Contribution}
+\includegraphics[width=0.9\columnwidth]{alg-ddv.png}
+\end{frame}
+
+\begin{frame}
+\frametitle{The Trade-off}
+\framesubtitle{time vs. simulator calls}
+
+\includegraphics[scale=0.2]{simcall-RiverSwim.png}
+
+\includegraphics[scale=0.2]{simcall-SixArms.png}
+\begin{table}
+    \resizebox{\textwidth}{!}{%
+    \begin{tabular}{c|c c c c}
+        Experiment  &   \textbf{DDV-UPPER}   &   \textbf{DDV-OUU} &  \textbf{ MBIE-reset}  & \textbf{Fiechter}  \\
+        \hline
+        \textbf{RiverSwim} & $22,152.9 \approx 6h$ &   $14,284.8 \approx 4h$ &   $15,106.5 \approx 4h$   &   $5,790.4 \approx 1.5h$   \\
+        \textbf{SixArms} & $621,600 \approx 7 days!$ &   $710,065 \approx 8 days!$  &   $203,229 \approx 56h$ &   $194,800 \approx 54h$   \\
+    \end{tabular}}
+    \caption{Total clock time in seconds needed for execution of each algorithm}
+    \label{tab:my_label}
+\end{table}
 
@@
 
+\begin{frame}
+\frametitle{Side Notes}
+\begin{itemize}
+    \item The authors repeatedly stressed that Fiechter's method is not PAC-RL but PAC-MDP.
+    \item They used boldface to emboss their results but with a delicate trickery reported clock time per simulator calls.
+
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Found Typos}
+
+        \begin{enumerate}
+            \item Page 3882: \sout{Figure 1} $ \rightarrow $ Algorithm 1.
+            \item Page 3898: Table 1 caption. This table demonstrating clock time per simulator calls for all 4 experiments.
+        \end{enumerate}
+\end{frame}
 	%\includegraphics[scale=0.45]{img.jpg}

### 2018-11-19 AML-RL/FinalProject/AMLproject.tex
Updates from ShareLaTeX

\documentclass[a4paper,12pt]{article}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}

\usepackage{graphicx}
\usepackage{csquotes}
\usepackage[style=verbose-ibid,backend=bibtex]{biblatex}
\bibliography{sample}

\title{Advanced Machine Learning - Final Project}

\author{Shayan Amani}

\begin{document}
\maketitle

\section{Questions}
\begin{itemize}
    \item 2000 samples for 10 times of 200 years. What kind of times?
    \item 
\end{itemize}

\section{Definition}
We have 2000 samples for 200 years.

\section{Methodology}
We can use either On-Policy or Off-Policy methods.
\end{document}

### 2018-11-26 summary/Safe Policy Improvement/main.tex
Updates from ShareLaTeX

\documentclass[a4paper,12pt]{article}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}

%
% For alternative styles, see the biblatex manual:
% http://mirrors.ctan.org/macros/latex/contrib/biblatex/doc/biblatex.pdf
%
% The 'verbose' family of styles produces full citations in footnotes, 
% with and a variety of options for ibidem abbreviations.
%
\usepackage{graphicx}
\usepackage{csquotes}
\usepackage[style=verbose-ibid,backend=bibtex]{biblatex}
\bibliography{sample}

\usepackage{lipsum} % for dummy text

\title{Safe Policy Improvement by Minimizing Robust Baseline Regret}

\author{Shayan Amani}

\date{\today}

\begin{document}
\maketitle

\section{Summary}
Safety is what you need to have at hand when you are trying to persuade stakeholders. As a matter of interest, ensuring about the model's performance is always favorable and in some cases essential. This paper tries to come up with a \textit{safe} policy which guarantees better performance than the \textit{baseline} policy. Authors explained how to go beyond regular robustness common in the literature with demonstrating proofs and proposing the algorithm.

Regret minimization is a topic of interest in RL which is discussed fairly well in this paper. Performance-wise talking, the proposed approach is promising over other available methods.

\includegraphics[width=1\columnwidth]{fig2.png}

\end{document}

### 2018-12-09 codebase/Albers_projections.py
Stan models for Bayesian learning and Albers projection added

@@ -1,0 +1,75 @@
+# Authored by Shayan Amani
+# @SHi-ON
+# 11/21/2018
+
+from mpl_toolkits.basemap import Basemap
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib
+import pandas as pd
+
+# in case of PROJ_LIB error related to Proj4 package, set the following Environment variable:
+# PROJ_LIB = "~/anaconda3/envs/<env_name>/share/proj"
+
+# shift offset to project data points in the actual area (US region).
+OFFSET_LAT = 23
+OFFSET_LON = -96
+
+
+def data_prep():
+    df = pd.read_csv("./datasets/barberry_sim_NewEngland.csv")
+    num_sub_samples = 600
+    # Pick random sub-samples
+    sub_samples = df.sample(n=num_sub_samples)
+    # lats = sub_samples['lat'].values / 100000 + OFFSET_LAT
+    # lons = sub_samples['lon'].values / 100000 + OFFSET_LON
+    lats = sub_samples['lat'].values
+    lons = sub_samples['lon'].values
+    alts = sub_samples['N'].values
+    return lats, lons, alts
+
+
+def bm_plotter(lats, lons, alts):
+    # setup Albers Equal Area conic projection basemap
+    # lat_1 is first standard parallel.
+    # lat_2 is second standard parallel.
+    # lon_0,lat_0 is central point.
+    m = Basemap(width=10000000, height=8000000,
+                resolution='l', projection='aea', \
+                lat_1=29.5, lat_2=45.5, lat_0=23, lon_0=-96)
+    m.drawcoastlines()
+    m.drawcountries()
+    m.fillcontinents(color='grey', lake_color='aqua')
+    # draw parallels, meridians and states boundaries
+    m.drawparallels(np.arange(-80., 81., 20.), labels=[1, 0, 0, 0])
+    m.drawmeridians(np.arange(-180., 181., 20.), labels=[0, 0, 0, 1])
+    m.drawmapboundary(fill_color='aqua')
+    m.drawstates()
+
+    ax = plt.gca()
+
+    # format colors for elevation range
+    alt_min = np.min(alts)
+    alt_max = np.max(alts)
+    cmap = plt.get_cmap('gist_earth')
+    normalize = matplotlib.colors.Normalize(vmin=alt_min, vmax=alt_max)
+
+    for ii in range(0, len(alts)):
[…]

### 2018-12-09 codebase/BayesianLearning/LogRegStan.py

@@ -10,7 +10,7 @@ import matplotlib.pyplot as plt
 # %precision 4
 plt.style.use('ggplot')
 
-df = pd.read_csv('HtWt.csv')
+df = pd.read_csv('../datasets/HtWt.csv')
 df.head()  # printing first 5 rows
 
 log_reg_code = """

### 2018-12-09 codebase/BayesianLearning/Weighted_L1_Experiments.py

@@ -1,17 +1,14 @@
 import pandas as pd
-import random
 import math
 import numpy as np
 import matplotlib.pyplot as plt
 
 from sklearn import model_selection
 from sklearn.linear_model import LogisticRegression
-from sklearn import datasets
 
 import pystan
 
-### cd /home/reazul/PhD_Research/MaxEnt_PointProcess/Codes/Experiments
-df = pd.read_csv("datasets/barberry_sim_NewEngland.csv")
+df = pd.read_csv("../datasets/barberry_sim_NewEngland.csv")
 num_sub_samples = 200
 sub_samples = df.sample(n=num_sub_samples)
 
@@ -30,7 +27,7 @@ Y = abundance_samples['N'].values
 def fit_bayesian_logistic_regression(X, y):
     stan_dat = {'N': len(X), 'detected': y.astype(int), 'bio10_5': X[:, 0], 'bio10_prMay': X[:, 1]}
 
-    sm = pystan.StanModel(file='model_logistic.stan')  # pystan.StanModel(model_code=m1)
+    sm = pystan.StanModel(file='./models/model_logistic.stan')  # pystan.StanModel(model_code=m1)
     # fit = sm.sampling(data=stan_dat, iter=20, chains=4, refresh=0)
     fit = sm.sampling(data=stan_dat, iter=180, chains=3, refresh=0)
 

### 2018-12-09 codebase/StanPg.py

@@ -1,57 +1,0 @@
-import pystan
-import numpy as np
-import matplotlib.pyplot as plt
-
-stan_code_model = """
-data {
-    int ns;  // num of samples
-    int nf;  // num of features
-    matrix[ns,nf] phi;  // feature matrix
-    vector[ns] y;  // response
-}
-parameters {
-    vector[nf] beta;
-    real<lower=0> epsilon; 
-}
-model {
-    y ~ normal(phi * beta, epsilon);
-}
-"""
-
-
-def distro():
-    # stan_data = {'num_samples': 30, 'y': d}
-
-    randNumGen = np.random.RandomState(7)
-    n_samp = 10
-    n_feat = 4
-
-    stan_model = pystan.StanModel(model_code=stan_code_model)
-
-    features = randNumGen.normal(0, 1, size=(n_samp, n_feat))
-    response = features @ np.array([-1, 0.4, 0, 0]) + randNumGen.normal(scale=0.2, size=n_samp)
-
-    stan_fit = stan_model.sampling(data={'ns': n_samp,
-                                         'nf': n_feat,
-                                         'phi': features,
-                                         'y': response},
-                                   iter=1000, chains=4, warmup=500, n_jobs=-1, seed=7)
-
-    chains = stan_fit.extract()
-
-    chains['beta'].shape
-
-    ft_plot =  stan_fit.plot()
-    stan_fit.plot()
-
-
-    lin.plot(chains['beta'][:,0])
-
-
-
-def main():
-    distro()
-
-
-if __name__ == '__name__':
-    main()

### 2018-12-09 codebase/playground/DQNcatch/qlearn.py

@@ -1,171 +1,0 @@
-import json
-import numpy as np
-from keras.models import Sequential
-from keras.layers.core import Dense
-from keras.optimizers import sgd
-
-
-class Catch(object):
-    def __init__(self, grid_size=10):
-        self.grid_size = grid_size
-        self.reset()
-
-    def _update_state(self, action):
-        """
-        Input: action and states
-        Ouput: new states and reward
-        """
-        state = self.state
-        if action == 0:  # left
-            action = -1
-        elif action == 1:  # stay
-            action = 0
-        else:
-            action = 1  # right
-        f0, f1, basket = state[0]
-        new_basket = min(max(1, basket + action), self.grid_size - 1)
-        f0 += 1
-        out = np.asarray([f0, f1, new_basket])
-        out = out[np.newaxis]
-
-        assert len(out.shape) == 2
-        self.state = out
-
-    def _draw_state(self):
-        im_size = (self.grid_size,) * 2
-        state = self.state[0]
-        canvas = np.zeros(im_size)
-        canvas[state[0], state[1]] = 1  # draw fruit
-        canvas[-1, state[2] - 1:state[2] + 2] = 1  # draw basket
-        return canvas
-
-    def _get_reward(self):
-        fruit_row, fruit_col, basket = self.state[0]
-        if fruit_row == self.grid_size - 1:
-            if abs(fruit_col - basket) <= 1:
-                return 1
-            else:
-                return -1
-        else:
-            return 0
-
-    def _is_over(self):
-        if self.state[0, 0] == self.grid_size - 1:
-            return True
-        else:
-            return False
-
-    def observe(self):
-        canvas = self._draw_state()
-        return canvas.reshape((1, -1))
-
-    def act(self, action):
-        self._update_state(action)
-        reward = self._get_reward()
-        game_over = self._is_over()
-        return self.observe(), reward, game_over
-
-    def reset(self):
-        n = np.random.randint(0, self.grid_size - 1, size=1)
-        m = np.random.randint(1, self.grid_size - 2, size=1)
-        self.state = np.asarray([0, n, m])[np.newaxis]
-
-
-class ExperienceReplay(object):
-    def __init__(self, max_memory=100, discount=.9):
-        self.max_memory = max_memory
-        self.memory = list()
-        self.discount = discount
-
-    def remember(self, states, game_over):
-        # memory[i] = [[state_t, action_t, reward_t, state_t+1], game_over?]
[…]

### 2018-12-09 codebase/playground/DQNcatch/test.py

@@ -1,45 +1,0 @@
-import json
-import matplotlib.pyplot as plt
-import numpy as np
-from keras.models import model_from_json
-from qlearn import Catch
-
-if __name__ == "__main__":
-    # Make sure this grid size matches the value used fro training
-    grid_size = 10
-
-    with open("./mdl/model.json", "r") as jfile:
-        model = model_from_json(json.load(jfile))
-    model.load_weights("./mdl/model.h5")
-    model.compile("sgd", "mse")
-
-    # Define environment, game
-    env = Catch(grid_size)
-    c = 0
-    for e in range(10):
-        loss = 0.
-        env.reset()
-        game_over = False
-        # get initial input
-        input_t = env.observe()
-
-        plt.imshow(input_t.reshape((grid_size,) * 2),
-                   interpolation='none', cmap='gray')
-        plt.show()
-        plt.savefig("./figs/%03d.png" % c)
-        c += 1
-        while not game_over:
-            input_tm1 = input_t
-
-            # get next action
-            q = model.predict(input_tm1)
-            action = np.argmax(q[0])
-
-            # apply action, get rewards and new state
-            input_t, reward, game_over = env.act(action)
-
-            plt.imshow(input_t.reshape((grid_size,) * 2),
-                       interpolation='none', cmap='gray')
-            plt.show()
-            plt.savefig("./figs/%03d.png" % c)
-            c += 1

### 2018-12-09 AML-RL/FinalProject/AMLproject.tex

@@
     \item 2000 samples for 10 times of 200 years. What kind of times?
-    \item 
+    \item What are states? How many?
 \end{itemize}
@@
 \section{Definition}
-We have 2000 samples for 200 years.
+We have 2000 samples for 200 years. In terms of games, this is like a already played game and we have the history of the gameplay.
+
+\subsection{State}
+each row
+
+\subsection{Action}
+2 actions: 0 and 10.
+$\epsilon$-greedy exploration is not applied, because we have no control over agent.
+
+\subsection{Reward}
+A summation according to it's definition.
 
@@
 We can use either On-Policy or Off-Policy methods.
+From the given samples it could be simply inferred that a model-free algorithm is needed and the reason is that the problem doesn't provide the transition probabilities.
 \end{document}

### 2018-12-18 presentation/AML Final Project/main.tex
presentation: AML final project added

--- 2018-11-19 AML-RL/FinalProject/AMLproject.tex
@@
-\documentclass[a4paper,12pt]{article}
-\usepackage[english]{babel}
+% Golden Spiral can be approximated with Golden ratio or Fibonacci sequence.
+% Golden Ratio: a+b/a = a/b
+
+\documentclass{beamer}
 \usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
 
-\usepackage{graphicx}
-\usepackage{csquotes}
-\usepackage[style=verbose-ibid,backend=bibtex]{biblatex}
-\bibliography{sample}
+\usetheme{Cuerna}
+\usecolortheme{default}
+% default, bluesimplex, lettuce, brick
 
-\title{Advanced Machine Learning - Final Project}
-
+\title{AML-RL Final Project}
 \author{Shayan Amani}
 
+\date{fall 2018}
+\institute{Department of Computer Science, University of New Hampshire}
+
 \begin{document}
-\maketitle
 
-\section{Questions}
+\begin{frame}
+    \titlepage
+\end{frame}
+
+\begin{frame}{Methodology}
+\textbf{Deep Q-Network}
+
+Why?
 \begin{itemize}
-    \item 2000 samples for 10 times of 200 years. What kind of times?
-    \item 
+    \item Applying a non-linear approximator such as neural networks seems reasonable to reveal unseen relations between among data samples.
+    \item Inquisitiveness 
+    \item Q-learning performed well in the literature.
+\end{itemize} \\\\
+In that sense: \\
+\centering  Q-learning + neural networks $\rightarrow$ DQN
+\end{frame}
+
+\begin{frame}{Ideas}
+
+  This approach take advantage of two key ideas:
+  \begin{enumerate}
+      \item Experience Replay: randomized sample drawing to breaks correlations between consecutive observations.
+      \item Frequent Q updates: adjusts action-values (Q) iteratively.
+  \end{enumerate}
+\end{frame}
+
+\begin{frame}{Algorithm}
+\includegraphics[scale=0.3]{alg-dqn.png}
+
+\end{frame}
+
+\begin{frame}{Architecture}
+    The neural network:
+    \begin{itemize}
+        \item 3 dense layers
+        \item ReLU activation function
+        \item features fed in as inputs
+        \item 2 action-values (Q values) as output
     \end{itemize}
 
-\section{Definition}
-We have 2000 samples for 200 years.
+\end{frame}
 
-\section{Methodology}
-We can use either On-Policy or Off-Policy methods.
+\begin{frame}{Training}
+
+\includegraphics[scale=0.35]{myplot.png}
+
+\end{frame}
+
+\begin{frame}{Other Proposed Techniques}
+    \item \textbf{$\epsilon$-greedy exploration}: batch learning vs online learning
+    \item \textbf{reward clipping}: doesn't remarkably affect the Q values.
+\end{frame}
+
+\begin{frame}{Applied Techniques}
+Random batch sampling from the memory of samples to break correlation between consecutive samples (i.e. consecutive years).
+
+\end{frame}
+
+	%\includegraphics[scale=0.45]{img.jpg}
+
+% \begin{frame}{To-Do}
+
+% \begin{itemize}
+%  \item add more layers
+%  \item describe ExpRep in slides
+%  \item PCA
+%  \item reward clipping
+%  \item other variation of DQN
+%  \item get rid of game over
+%  \item model training for 500 epochs
+%  \item q learning implementation vs qnet
+%  \item pytorch imp
+% \end{itemize}
+
+% \end{frame}
+
+\begin{frame}{Future Steps}
+\begin{itemize}
+    \item Applying Double Q-Networks and other acclaimed working variation of the configuration. 
+    \item Take a deeper look at the features interactions
+    \item Tuning hyper-parameters like learning rate in case of bigger set of samples.
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Acknowledgment}
+\begin{itemize}
+    \item UC Berkley’s course on deep reinforcement learning
+    \item Algorithm adapted from \href{http://artint.info/html/ArtInt_265.html}{\textit{Artificial Intelligence: Foundations of Computational Agents, second edition, Cambridge University Press 2017}}
+\end{itemize}
+\end{frame}
+
 \end{document}
+
+%     \section{Questions}
+% \begin{itemize}
+%     \item 2000 samples for 10 times of 200 years. What kind of times?
+%     \item What are states? How many?
+% \end{itemize}
+
+% \section{Definition}
+% We have 2000 samples for 200 years. In terms of games, this is like a already played game and we have the history of the gameplay.
+
+% \subsection{State}
+% each row
+
+% \subsection{Action}
+% 2 actions: 0 and 10.
+% $\epsilon$-greedy exploration is not applied, because we have no control over agent.
+
+% \subsection{Reward}
+% A summation according to it's definition.
+
+% We can use either On-Policy or Off-Policy methods.
+From the given samples it could be simply inferred that a model-free algorithm is needed and the reason is that the problem doesn't provide the transition probabilities.

### 2018-12-18 codebase/DQN/memory.py
adding DQN

@@ -1,0 +1,59 @@
+import numpy as np
+
+
+class ExperienceReplay(object):
+
+    def __init__(self, mem_len=100, disc=0.9):
+        self.mem_size = mem_len
+        self.disc = disc
+        self.memory = list()
+
+    def feature_dimension(self):
+        return self.memory[0][0][0].shape[1]
+
+    def store(self, record, game_over):
+        # memory[i] = [[s, a, r, s'], game_over]
+        self.memory.append([record, game_over])
+        if len(self.memory) > self.mem_size:
+            del self.memory[0]
+
+    # applying Bellman equation (update)
+    def get_batch(self, model, batch_size=10):
+        memory_len = len(self.memory)
+        # number of possible actions
+        num_actions = model.output_shape[-1]
+        # dimension of the input features
+        phi_dim = self.feature_dimension()
+        # min(len_memory, batch_size) <=> inputs.shape[0]
+        rows = min(memory_len, batch_size)
+
+        # initializing input features.
+        # min just in case that still not many states observed
+        inputs = np.zeros((rows, phi_dim))
+
+        # initializing target action-values (Q-table)
+        # target: Q(s,a) = r + gamma * max Q(s’,a’)
+        # Note that our target is a matrix, with possible fields not only for the action taken but also
+        # for the other possible actions. The actions not take the same value as the prediction to not affect them
+        targets = np.zeros((rows, num_actions))
+
+        for i, num in enumerate(np.random.randint(0, memory_len, size=rows)):
+            # random sampling to break correlations
+            # (s, a, r, s')
+            s, a, r, s_p = self.memory[num][0]
+            game_over = self.memory[num][1]
+
+            # storing state s in the inputs
+            inputs[i:i + 1] = s
+
+            # targets Qs
+            targets[i] = model.predict(s)[0]
+            # max_a' Q(s', a')
+            max_q_sp = np.max(model.predict(s_p)[0])
+            # if it's the last year
+            if game_over:
+                targets[i, a] = r
+            else:
+                # r + gamma * max_a' Q(s', a')
+                targets[i, a] = r + self.disc * max_q_sp
+        return inputs, targets

### 2018-12-18 codebase/DQN/qtrain.py

@@ -1,0 +1,114 @@
+import pandas as pd
+import numpy as np
+
+from keras.models import Sequential
+from keras.layers import Dense
+# Stochastic Gradient Descent optimizer
+# from keras.optimizers import sgd
+from keras.optimizers import rmsprop
+# from keras.optimizers import adam
+import json
+import time
+
+import matplotlib.pyplot as plt
+
+# sampling handler class and parameters
+from util import *
+import sample as samp
+# Experience Replay adapter class
+import memory as memo
+
+
+# import os
+# os.chdir("./codebase/DQN")
+
+
+def handler():
+    _df = pd.read_csv("./datasets/" + samples_name + ".csv", names=col_names, header=0)
+
+    print(_df.head())
+
+    _df.isnull().any()
+
+    return _df
+
+
+def network_builder():
+    model = Sequential()
+    # grid_size x 2 = 18 inputs
+    model.add(Dense(hidden_size, input_shape=(grid_size * 2,), activation='relu'))
+    model.add(Dense(hidden_size, activation='relu'))
+    # num_actions = 2 outputs
+    model.add(Dense(num_actions))
+    # lr: learning rate
+    model.compile(rmsprop(lr=alpha), "mse")
+
+    model.summary()
+    return model
+
+
+# store trained model's weights and configurations
+def model_export(mdl):
+    mdl.save_weights("./models/" + model_name + ".h5", overwrite=True)
+    with open("./models/" + model_name + ".json", "w") as out_handle:
+        json.dump(mdl.to_json(), out_handle)
+
+
+def train(df):
+
+    model = network_builder()
+
+    exp_rep = memo.ExperienceReplay(mem_len=mem_len, disc=gamma)
+    sample = samp.Sample(df, v1_max, reward_col)
+
+    epochs = sample.get_epochs()
+    loss_df = pd.DataFrame({'step': range(1, v1_max * epochs + 1), 'loss': range(1, v1_max * epochs + 1)})
+    for ep in range(epochs):
+        loss = 0.
+        next_state = sample.get_init_sample(ep)
+
+        # duplicate state prevention in initial step w/ start from 1
+        for i in range(1, v1_max):
+            state = next_state
+
+            q = model.predict(state)
+            action = np.argmax(q[0])
+            # print("Action in ", ep, "and step", i)
+
+            reward = sample.get_reward(ep, i)
+
+            next_state = sample.get_sample(ep, i)
+
+            game_over = sample.is_over(ep)
+
[…]

### 2018-12-18 codebase/DQN/run.py

@@ -1,0 +1,81 @@
+import numpy as np
+import pandas as pd
+from keras.models import model_from_json
+import matplotlib.pyplot as plt
+
+from qtrain import *
+from util import *
+import sample as samp
+
+
+def model_import():
+    with open("./models/" + model_name + ".json", "r") as m_file:
+        mdl = model_from_json(json.load(m_file))
+        mdl.load_weights("./models/" + model_name + ".h5")
+        mdl.compile("adam", "mse")  # TODO
+    return mdl
+
+
+def policy_store(df, plc):
+    plc_df = pd.DataFrame(plc)
+    new_df = pd.concat([df, plc_df], axis=1, sort=False)
+    new_df.columns = policy_col_names
+    new_df.to_csv("./datasets/" + policy_name + samples_name + ".csv", sep=",")
+
+
+def run_model():
+    d_f = handler()
+    smpl = samp.Sample(d_f, v1_max, reward_col)
+    epochs = smpl.get_epochs()
+
+    policy = np.zeros((epochs * v1_max, 1))
+
+    model = model_import()
+
+    for ep in range(epochs):
+        loss = 0.
+
+        next_state = smpl.get_init_sample(ep)
+        game_over = smpl.is_over(ep)
+
+        # plt.imshow(input_t.reshape((grid_size,) * 2), interpolation='none', cmap='gray')
+        # plt.imshow(next_state, interpolation='none', cmap='gray')
+        # plt.show()
+        # plt.savefig("./figs/%03d.png" % ep)
+
+        for i in range(1, v1_max):
+            state = next_state
+
+            q = model.predict(state)
+            # action = np.argmax(q[0])
+            action = np.argmax(q[0])
+            # if action == 1:
+            #     print("Action in ", ep, "and step", i)
+            policy[ep * v1_max + i] = pick_action(q[0])
+
+            next_state = smpl.get_sample(ep, i)
+            game_over = smpl.is_over(ep)
+
+            # plt.imshow(input_t.reshape((grid_size,) * 2), interpolation='none', cmap='gray')
+            # plt.imshow(next_state, interpolation='none', cmap='gray')
+            # plt.show()
+            # plt.savefig("./figs/%03d.png" % (ep + 1))
+
+    policy_store(d_f, policy)
+
+
+def moving_average_diff(a, n=100):
+    diff = np.diff(a)
+    ret = np.cumsum(diff, dtype=float)
[…]

### 2018-12-18 codebase/DQN/sample.py

@@ -1,0 +1,32 @@
+class Sample(object):
+
+    def __init__(self, df, yrs, r_col):
+        self.df = df
+        self.yrs = yrs
+        self.r_col = r_col
+
+    # return ndarray version of the data frame
+    def _nd_array(self):
+        return self.df.values
+
+    # return number of epochs exist in dataset
+    def get_epochs(self):
+        return self.df.loc[:, 'V1'].value_counts().to_dict()[self.yrs]
+
+    def is_over(self, ep):
+        return ep == self.yrs - 1
+
+    def get_reward(self, epoch, index):
+        values = self._nd_array()
+        row = epoch * self.yrs
+        return values[row + index, self.r_col]
+
+    def get_init_sample(self, epoch, start=1, end=19):
+        values = self._nd_array()
+        row = epoch * self.yrs
+        return values[row, start:end].reshape((1, -1))
+
+    def get_sample(self, epoch, index, start=1, end=19):
+        values = self._nd_array()
+        row = epoch * self.yrs
+        return values[row + index, start:end].reshape((1, -1))

### 2018-12-18 codebase/DQN/util.py

@@ -1,0 +1,59 @@
+import numpy as np
+
+grid_size = 9  # 3 x 3
+num_actions = 2  # [0, 10]
+v1_max = 200
+reward_col = 20
+
+samples_name = "project_sample1"
+# samples_name = "sample2"
+
+policy_name = "policy_"
+
+md_name = "dqn_model"
+md_appendix_10_1 = "_10_1"
+md_appendix_10_50 = "_10_50"
+md_appendix_500_1 = "_500_1"
+md_appendix_500_50 = "_500_50"
+# model_name = md_name
+# model_name = md_name + md_appendix_500_1
+model_name = md_name + md_appendix_10_1
+
+
+# hyper-parameters
+gamma = 0.9  # discount factor
+epsilon = .1  # epsilon-greedy exploration
+alpha = .2  # learning rate
+
+mem_len = 500  # Experience Replay length
+batch_size = 1  # number of experiences per batch
+# batch_size = 50  # number of experiences per batch
+hidden_size = 100  # size of the hidden layers
+
+col_names = ['V1', 'pop_1', 'pop_2', 'pop_3', 'pop_4', 'pop_5', 'pop_6',
+                 'pop_7', 'pop_8', 'pop_9', 'sbank_1', 'sbank_2', 'sbank_3',
+                 'sbank_4', 'sbank_5', 'sbank_6', 'sbank_7', 'sbank_8', 'sbank_9',
+                 'actions', 'rewards']
+
+policy_col_names = ['V1', 'pop_1', 'pop_2', 'pop_3', 'pop_4', 'pop_5', 'pop_6',
+                 'pop_7', 'pop_8', 'pop_9', 'sbank_1', 'sbank_2', 'sbank_3',
+                 'sbank_4', 'sbank_5', 'sbank_6', 'sbank_7', 'sbank_8', 'sbank_9',
+                 'actions', 'rewards', 'policy']
+
+
+def pick_action(q):
+    idx = np.argmax(q)
+    return 0 if idx == 0 else 10
+
+
+def stopwatch_log(start, end):
+    line = model_name + " for " + samples_name + " = " + str(np.round(end - start, 2)) + " seconds " + "\n"
+    with open("./DQN/time_sheet.txt", "a") as time_file:
+        time_file.write(line)
+
+
+def plotter(df, idx, loss):
+    if loss > 10e6:
+        df.iloc[idx, 1] = 10e6
+
+

### 2018-12-19 codebase/DQN/qtrain.py
codebase: DQN function organization updated

@@ -7,7 +7,6 @@ from keras.layers import Dense
 # from keras.optimizers import sgd
 from keras.optimizers import rmsprop
 # from keras.optimizers import adam
-import json
 import time
 
 import matplotlib.pyplot as plt
@@ -23,16 +22,6 @@ import memory as memo
 # os.chdir("./codebase/DQN")
 
 
-def handler():
-    _df = pd.read_csv("./datasets/" + samples_name + ".csv", names=col_names, header=0)
-
-    print(_df.head())
-
-    _df.isnull().any()
-
-    return _df
-
-
 def network_builder():
     model = Sequential()
     # grid_size x 2 = 18 inputs
@@ -45,13 +34,6 @@ def network_builder():
 
     model.summary()
     return model
-
-
-# store trained model's weights and configurations
-def model_export(mdl):
-    mdl.save_weights("./models/" + model_name + ".h5", overwrite=True)
-    with open("./models/" + model_name + ".json", "w") as out_handle:
-        json.dump(mdl.to_json(), out_handle)
 
 
 def train(df):
@@ -102,7 +84,7 @@ def train(df):
 
 
 if __name__ == '__main__':
-    data_frame = handler()
+    data_frame = dataset_import()
 
     a_time = time.time()
     train(data_frame)

### 2018-12-19 codebase/DQN/run.py

@@ -1,30 +1,13 @@
 import numpy as np
 import pandas as pd
-from keras.models import model_from_json
 import matplotlib.pyplot as plt
 
-from qtrain import *
 from util import *
 import sample as samp
 
 
-def model_import():
-    with open("./models/" + model_name + ".json", "r") as m_file:
-        mdl = model_from_json(json.load(m_file))
-        mdl.load_weights("./models/" + model_name + ".h5")
-        mdl.compile("adam", "mse")  # TODO
-    return mdl
-
-
-def policy_store(df, plc):
-    plc_df = pd.DataFrame(plc)
-    new_df = pd.concat([df, plc_df], axis=1, sort=False)
-    new_df.columns = policy_col_names
-    new_df.to_csv("./datasets/" + policy_name + samples_name + ".csv", sep=",")
-
-
 def run_model():
-    d_f = handler()
+    d_f = dataset_import()
     smpl = samp.Sample(d_f, v1_max, reward_col)
     epochs = smpl.get_epochs()
 
@@ -61,7 +44,7 @@ def run_model():
             # plt.show()
             # plt.savefig("./figs/%03d.png" % (ep + 1))
 
-    policy_store(d_f, policy)
+    policy_export(d_f, policy)
 
 
 def moving_average_diff(a, n=100):

### 2018-12-19 codebase/DQN/util.py

@@ -1,4 +1,7 @@
 import numpy as np
+import pandas as pd
+import json
+from keras.models import model_from_json
 
 grid_size = 9  # 3 x 3
 num_actions = 2  # [0, 10]
@@ -41,6 +44,36 @@ policy_col_names = ['V1', 'pop_1', 'pop_2', 'pop_3', 'pop_4', 'pop_5', 'pop_6',
                  'actions', 'rewards', 'policy']
 
 
+def dataset_import():
+    _df = pd.read_csv("./datasets/" + samples_name + ".csv", names=col_names, header=0)
+    print(_df.head())
+    _df.isnull().any()
+    return _df
+
+
+# open the already-trained model's weights and configurations
+def model_import():
+    with open("./models/" + model_name + ".json", "r") as m_file:
+        mdl = model_from_json(json.load(m_file))
+        mdl.load_weights("./models/" + model_name + ".h5")
+        mdl.compile("adam", "mse")  # TODO
+    return mdl
+
+
+# store trained model's weights and configurations
+def model_export(mdl):
+    mdl.save_weights("./models/" + model_name + ".h5", overwrite=True)
+    with open("./models/" + model_name + ".json", "w") as out_handle:
+        json.dump(mdl.to_json(), out_handle)
+
+
+def policy_export(df, plc):
+    plc_df = pd.DataFrame(plc)
+    new_df = pd.concat([df, plc_df], axis=1, sort=False)
+    new_df.columns = policy_col_names
+    new_df.to_csv("./datasets/" + policy_name + samples_name + ".csv", sep=",")
+
+
 def pick_action(q):
     idx = np.argmax(q)
     return 0 if idx == 0 else 10
@@ -55,5 +88,3 @@ def stopwatch_log(start, end):
 def plotter(df, idx, loss):
     if loss > 10e6:
         df.iloc[idx, 1] = 10e6
-
-

### 2018-12-23 minutia/AML-report/main.tex
Updates from ShareLaTeX

\documentclass[12pt]{report}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}

\usepackage[dvipsnames]{xcolor}

\usepackage{array}
\newcolumntype{L}{>{\centering\arraybackslash}m{3cm}}

\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{caption}

\usepackage{graphicx}
\usepackage{csquotes}

\title{Advanced Machine Learning\\Final Report}

\author{Shayan Amani}

\date{\today}

\begin{document}
\maketitle

\chapter{Problem Definition}
Presenting an optimum policy to control a specific invasive species, Frangula alnus so called, glossy buckthorn. For the sake of simplicity, a grid environment of 3x3 has been considered.

Simulator-generated samples of history of measures and consequences are provided in a dataset. Dataset is consisted of the following columns:
\begin{itemize}
    \item \textbf{V1}: year number of the sample. from 1 to 200 which means simulator run through a span of 200 years.
    \item: \textbf{pop\_1} to \textbf{pop\_9}: population of the plant in each cells.
    \item: \textbf{sbank\_1} to \textbf{sbank\_9}: number of plant's seeds in each cells.
    \item: \textbf{actions}: taken action for each year of samples.
    \item: \textbf{rewards}: observed reward after applying the action.
\end{itemize}

\chapter{Previous Works}
We have two different point of view to look at this problem. Firstly a batch reinforcement learning problem and secondly

\chapter{Proposed and Implemented Method}
Considering this sample-based problem I have picked and implemented a deep reinforcement learning approach, namely Deep Q-Networks (DQN). The elements of this method are explained in the following sections.

This needs to be mentioned that I have examined different method which are excelled in the literature. One of these methods is Deep Deterministic Policy Gradient (DDPG). According

\section{Algorithm}
The whole algorithm is described in Algorithm \ref{alg:DQN}.

\begin{algorithm}[H]
\caption{DQN algorithm in batch mode}
\label{alg:DQN}
\begin{algorithmic}[1]
    \State Init D \Comment{\textcolor{BlueViolet}{replay memory}}
    \State Init Q \Comment{\textcolor{BlueViolet}{Q-table w/ random weights}}
    \State \textcolor{OrangeRed}{Get} or Observe $s_0$ \Comment{\textcolor{BlueViolet}{the initial state}}
    \For{each episode}
        \For{samples in each episode}
            \State \textit{\textcolor{OrangeRed}{- skip $\epsilon$-greedy exploration}}
            \State $a = argmax_a Q(s,a)$
            \State \textit{\textcolor{OrangeRed}{- skip applying action $a$ to the environment}}
            \rlap{\smash{$\left.\begin{array}{@{}c@{}}\\{}\\{}\end{array}\color{BlueViolet}\right\}%
              \color{BlueViolet} experience$}}
            \State \textcolor{OrangeRed}{Get} or Observe $r, s'$
            \State Store the \textbf{experience} $<s, a, r, s'>$ in replay memory $D$
            \State \textit{Random} \textbf{sampling} from $D$ $<ss, aa, rr, ss'>$  \Comment{\textcolor{BlueViolet}{[mini]-batch}}
            \If{$ss' \neq $ terminal state} \Comment{\textcolor{BlueViolet}{target for each mini-batch}}
                \State $tt = rr + \gamma max_{aa'} Q(ss', aa')$
            \Else
                \State $tt = rr$
            \EndIf
            \State Train the network
        \EndFor    
    \EndFor
\end{algorithmic}
\end{algorithm}

\section{Experience Replay}
DQN is mostly relied on a stabilization technique with neural networks, called Experience replay. Experience replay is a memory abstraction which $n$ past samples are stored in this memory and then we use it later to draw samples from it on a randomized basis.

% \includegraphics[width=1\columnwidth]{elements.png}

\end{document}

### 2018-12-23 minutia/AML-report/main.tex
minutia: AML-report bibliography updated and contents added

@@
 \documentclass[12pt]{report}
-\usepackage[english]{babel}
+
 \usepackage[utf8]{inputenc}
@@
 \usepackage{csquotes}
+
+\renewcommand{\bibname}{References}
 
@@
 
+\chapter{Proposed Methods}
+As a problem in which the model (transition probabilities) is not provided, model-free design is what may first come to mind.
+Different options were available to choose among and each have their own pros and cons.
+From least squares temporal difference (LSTD) or least squares policy iteration (LSPI) algorithms to more sophisticated one such as asynchronous advantage actor-critic (A3C) \cite{Mnih2016b} are all available
+
 \chapter{Proposed and Implemented Method}
 Considering this sample-based problem I have picked and implemented a deep reinforcement learning approach, namely Deep Q-Networks (DQN).
+DQN is mostly relied on a stabilization technique with neural networks, called Experience replay.
  The elements of this method are explained in the following sections.
 
-This needs to be mentioned that I have examined different method which are excelled in the literature. One of these methods is Deep Deterministic Policy Gradient (DDPG). According
+This needs to be mentioned that I have examined different method which are excelled in the literature.
+One of these methods is Deep Deterministic Policy Gradient (DDPG).
+According to \citet{} such a methods are more applicable to continues action space problems which is not the case here.
+I have provided a one-to-one comparison chart in Table \ref{tab:dqn vs.
+ddpg} to discuss particular aspects of these two methods.
+
+\begin{table}
+    \centering
+    \begin{tabular}{|c|L|L|L|L|}
+        \hline
+        \textbf{Method}  &    \textbf{Model Dependence}    &   \textbf{Policy Dependence}   &   \textbf{Action Continuity}  & \textbf{Parent Algorithm(s)}\\
+        \hline
+        \hline
+        DQN &   model-free  &   on-policy   &   discrete    &   Q-learning, DL\\
+        \hline
+        DDPG    &   model-free  &   on-policy   &   continuous  &   Actor-critic\\
+        \hline
+    \end{tabular}
+    \caption{DRL methods comparison chart}
+    \label{tab:dqn vs. ddpg}
+\end{table}
 
@@
 
+\bibliographystyle{nips}
+\bibliography{library}
+
 \end{document}

### 2018-12-23 minutia/AML-report/main.tex
Updates from ShareLaTeX

@@
 \chapter{Problem Definition}
-Presenting an optimum policy to control a specific invasive species, Frangula alnus so called, glossy buckthorn. For the sake of simplicity, a grid environment of 3x3 has been considered.
+Presenting an optimum policy \cite{Sutton1998} to control a specific invasive species, Frangula alnus, so called glossy buckthorn.
+For the sake of simplicity, a grid environment of 3x3 has been considered.
 
@@
 
+Available actions are defined as a binary set of two actions, $0$ and $10$. Action $0$ means do nothing and action $10$ is equal to cutting and spraying all existing plants in the entire grid.
+
 \chapter{Previous Works}
-We have two different point of view to look at this problem. Firstly a batch reinforcement learning problem and secondly
+We have two different point of view to look at this problem.
+Firstly a batch reinforcement learning problem and secondly invasive species management problem.
+In retrospect, we can find numerous instances of trying different algorithms and approaches from the community \cite{Phillips2006}, \cite{Phillips2004}, and \cite{Warton2010}.
 
-\chapter{Proposed Methods}
+Such a inter-disciplinary problem is addressed by people active in various fields and they have their own novelties in their points of view.
+In this report I am elaborating on my approach to the given problem from a relatively new perspective, namely deep reinforcement learning which did not absorb too many attention from the community for presenting solutions to such cases.
+
+\chapter{Method Selection}
 As a problem in which the model (transition probabilities) is not provided, model-free design is what may first come to mind.
 Different options were available to choose among and each have their own pros and cons.
-From least squares temporal difference (LSTD) or least squares policy iteration (LSPI) algorithms to more sophisticated one such as asynchronous advantage actor-critic (A3C) \cite{Mnih2016b} are all available
+From least squares temporal difference (LSTD) \cite{Bradtke1996} or least squares policy iteration (LSPI) \cite{Lagoudakis2003} algorithms to more sophisticated one such as asynchronous advantage actor-critic (A3C) \cite{Mnih2016b} were all available.
+From a specific perspective, one can point out the diversity among these techniques by a component such as value function approximation technique.
 
-\chapter{Proposed and Implemented Method}
+According to the attributes that are observed from the given problem, the final implemented method has been selected. The significant characteristics of the given problem are as follows:
+\begin{itemize}
+    \item Model is not provided.
+    \item Dataset is only a set of generated samples.
+    \item Inaccessible agent therefore online learning is not possible.
+    \item Evidently high correlation between consequent year samples.
+\end{itemize}
+
+The mentioned issues are addressed in Chapter \ref{chap: implemented method}
+
+\chapter{Implemented Method}
+\label{chap: implemented method}
 Considering this sample-based problem I have picked and implemented a deep reinforcement learning approach, namely Deep Q-Networks (DQN).
-DQN is mostly relied on a stabilization technique with neural networks, called Experience replay.
+DQN is mostly relied on a stabilization technique with neural networks, called Experience replay \cite{Mnih2015}.
 The elements of this method are explained in the following sections.
 
-This needs to be mentioned that I have examined different method which are excelled in the literature.
-One of these methods is Deep Deterministic Policy Gradient (DDPG).
-According to \citet{} such a methods are more applicable to continues action space problems which is not the case here.
+DQN algorithm in a nutshell is a wrapper around Q-learning \cite{Watkins1992}.
+proposed as an online learning reinforcement learning algorithm, but in this case we need to implement it as a batch learning problem.
+
+One of the method which are excelled in the literature is Deep Deterministic Policy Gradient (DDPG).
+Such a methods are more applicable to continues action space problems which is not the case here \cite{Lillicrap2015}.
 I have provided a one-to-one comparison chart in Table \ref{tab:dqn vs.
-ddpg} to discuss particular aspects of these two methods.
+ddpg} to discuss particular aspects of DQN as the implemented method and DDPG as a studied method.
 
@@
 \section{Algorithm}
-The whole algorithm is described in Algorithm \ref{alg:DQN}.
+The implemented version of DQN algorithm is described in Algorithm \ref{alg:DQN}. Some variations and consideration are applied to fit this method to the problem.
 
@@
 
+\section{Neural Network}
+The concept of neural networks is a non-linear function approximator \cite{Bengio2009} which approximates action-value function (Q).
+The inputs to this network are 18 features (9 pop\_i plus 9 sbank\_i) and the outputs are Q-values for two actions 0 and 10.
+Input, output and hidden layers are selected to be dense (fully-connected) as it doesn't get involved with image processing tasks.
+
 \section{Experience Replay}
 DQN is mostly relied on a stabilization technique with neural networks, called Experience replay.
-Experience replay is a memory abstraction which $n$ past samples are stored in this memory and then we use it later to draw samples from it on a randomized basis.
+Experience replay is a memory abstraction which $n$ past samples are stored in this memory and then we use it later to draw a batch of samples from it on a randomized basis.
+
+Samples are accumulated in the memory while reading from the dataset and within a certain episode.
+
+\section{Random Exploration}
+One of the key steps in DQN algorithm is $\epsilon$-greedy exploration.
+The main purpose of this step is to address exploration-exploitation dilemma.
+However in this problem we have a batch RL problem and I've skipped this step in the main algorithm.
 

### 2018-12-27 minutia/AML-report/main.tex
writeup: clean commencement

@@
 
-\date{\today}
+\date{December 19, 2018}
 
@@
     \item Dataset is only a set of generated samples.
+    \item Episodic problem.
     \item Inaccessible agent therefore online learning is not possible.
@@
 
+\section{Results}
+Results are provided as dataset which in a policy column is appended.
+policy says which action should be taken at the end of each year for each episodes.
+From the loss values observed while training the network and the performance during the run-time, the results seems to be promising.
+I have also performed a set of measurements on after training like accuracy which shows the method works in operational regions.
+
 % \includegraphics[width=1\columnwidth]{elements.png}

### 2019-01-18 codebase/Gym/CartPole.py
OpenAI Gym - CartPole added

@@ -1,0 +1,13 @@
+import gym
+
+env = gym.make('CartPole-v0')
+for i_episode in range(20):
+    observation = env.reset()
+    for t in range(100):  # leave blank if you don't use step variable
+        env.render()
+        print(observation)
+        action = env.action_space.sample()
+        observation, reward, done, info = env.step(action)  # take a random action
+        if done:
+            print('Episode finished after {} timesteps'.format(t+1))
+            break

### 2019-01-31 presentation/Jan'19 Individual Meeting/main.tex

\documentclass{beamer}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\usetheme{Cuerna}
\usecolortheme{default}

\title{Progress Report}
\author{Shayan Amani\\}

\date{Jan 2019}
\institute{Department of Computer Science, University of New Hampshire}

\begin{document}

  \begin{frame}
    \titlepage
  \end{frame}

\begin{frame}
    \frametitle{Done}

    \begin{itemize}
        \item Laid out a sketch and implemented it to adapt a deep RL method works in offline setting. 
        \item Implemented Value-based Deep RL methods such as:
        \begin{itemize}
            \item Deep Q-Network (DQN) \cite{Mnih2015}
            \item Double DQN \cite{van2016a}
        \end{itemize}
        \item Studied some other proposed alternatives like Dueling Networks\cite{wang2015a}.
        \item Familiarized myself with Open AI's Gym RL frameworks and its popular environments.
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Deep RL Methods Schema}
    \includegraphics[width=1\columnwidth]{schema_deer.png}    
\end{frame}

\begin{frame}
    \frametitle{DQN Schema}
    \includegraphics[width=1\columnwidth]{dqn_schema.png}
\end{frame}

\begin{frame}
    \frametitle{To Be Done}
    The following need to be addressed and are proposed here as my future path:
    \begin{itemize}
        \item Random batch sampling from epochs ahead. Inspired by $\epsilon$-greedy approach.
        % \item finding a reliable alternative for $\epsilon$-greedy exploration.
        \item Trying other Loss calculation methods and compare them in practice.
    \end{itemize}
\end{frame}

\bibliographystyle{nips}
\bibliography{lib}

\end{document}

### 2019-02-05 codebase/MathOpt/plotter.py
resources: matplot lib notebook and Math. Opt. added

@@ -1,0 +1,36 @@
+import numpy as np
+from numpy import linalg as la
+from mpl_toolkits import mplot3d
+import matplotlib.pyplot as plt
+
+
+def f_func(a, b):
+    return 802 * a ** 2 - 400 * a * b + 200 * b ** 2
+
+
+if __name__ == "__main__":
+
+    x = np.linspace(-6, 6, 30)
+    y = np.linspace(-6, 6, 30)
+
+    X, Y = np.meshgrid(x, y)
+    Z = f_func(X, Y)
+
+    fig = plt.figure()
+    ax = plt.axes(projection='3d')
+
+    ax.contour3D(X, Y, Z, 200, cmap='magma')
+
+    fig.show()
+
+    # min finding
+    min_func = np.min(f_func(X, Y))
+    print("min=", min_func)
+
+    # eigenvalues
+    A = np.array([[802, -400], [-400, 200]])
+    eig_val, eig_vec = la.eig(A)
+    print("eigenvalues:", eig_val, la.eigvals(A))
+    print("eigenvectors:", eig_vec)
+
+

### 2019-02-18 codebase/MathOpt/plotter.py
math plot added

@@ -4,8 +4,8 @@ from mpl_toolkits import mplot3d
 import matplotlib.pyplot as plt
 
 
-def f_func(a, b):
-    return 802 * a ** 2 - 400 * a * b + 200 * b ** 2
+def f_rosenbrock(a, b):
+    return 100 * (b - a ** 2) ** 2 + (1 - a) ** 2
 
 
 if __name__ == "__main__":
@@ -14,17 +14,18 @@ if __name__ == "__main__":
     y = np.linspace(-6, 6, 30)
 
     X, Y = np.meshgrid(x, y)
-    Z = f_func(X, Y)
+    Z = f_rosenbrock(X, Y)
 
     fig = plt.figure()
     ax = plt.axes(projection='3d')
 
     ax.contour3D(X, Y, Z, 200, cmap='magma')
 
+    ax.view_init(60, 35)
     fig.show()
 
     # min finding
-    min_func = np.min(f_func(X, Y))
+    min_func = np.min(f_rosenbrock(X, Y))
     print("min=", min_func)
 
     # eigenvalues
@@ -32,5 +33,9 @@ if __name__ == "__main__":
     eig_val, eig_vec = la.eig(A)
     print("eigenvalues:", eig_val, la.eigvals(A))
     print("eigenvectors:", eig_vec)
+
+    # gradient
+    grad = np.gradient(f_rosenbrock(X, Y))
+    grad
 
 

### 2019-02-18 presentation/Adversarial Deep RL - group meeting presentation/main.tex

\documentclass{beamer}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\usetheme{Cuerna}
\usecolortheme{default}

% slide (frames) numbering
\addtobeamertemplate{navigation symbols}{}{
    \usebeamerfont{footline}%
    \usebeamercolor[fg]{footline}%
    \hspace{1em}%
    \insertframenumber
}

\title{Adversarial Attacks on Neural Network Policies}
\author{Shayan Amani\\}

\date{spring 2019}
\institute{RLRL Weekly Group Meeting}

\begin{document}

  \begin{frame}
    \titlepage
  \end{frame}

\begin{frame}{In This Paper}

Tampering input of the mentioned models with barely perceptible piece of information could lead to fooling the agent to take actions which amount to reduced final return (reward).

\begin{itemize}
    \item 3 type of attacks
    \item 3 methods
    \item 3 constraints
    \item 4 benchmark games
\end{itemize}

\end{frame}

\begin{frame}{Adversarial Attacks}
Adversarial attacks fall in either one of the following categories based upon the stage which in the adversary perturbs the model:
\begin{itemize}
    \item Train time: Data poisoning.
    \item Evaluation time: Adversarial examples.
    \item Deploy time: Black-box attacks.
\end{itemize}

\end{frame}

\begin{frame}{Contributions}
Contribution of this paper can be framed into two form of attacks to state-of-the-art methods, A3C, TRPO, and DQN. Inspired by cryptography and cyber attacks, they are named as follows:
\begin{itemize}
    \item \textbf{White-box attacks}: adversary has access to the model under-the-hood.
    \item \textbf{Black-box attacks}: adversary has no knowledge about the model.
\end{itemize}

\end{frame}

\begin{frame}
\includegraphics[width= 1\columnwidth]{fig1-l-inf-vs-l-1.png}
\end{frame}

\begin{frame}{Effect of Norm}
The paper has also reviewed effect of constraining three mentioned methods with $\ell_\infty$, $\ell_2$, and $\ell_1$ -norms.

\end{frame}

\begin{frame}
\includegraphics[width =1\columnwidth]{fig2-diff-norms.png}
\end{frame}

\begin{frame}
\includegraphics[width =1\columnwidth]{fig3-trans-A3C.png}
\end{frame}

\begin{frame}
\includegraphics[width =1\columnwidth]{fig4-trans-TRPO.png}
\end{frame}

\begin{frame}
\includegraphics[width =1\columnwidth]{fig5-trans-DQN.png}
\end{frame}

\end{document}

### 2019-02-25 codebase/MathOpt/line_search.py
Math opt problem set added and updated

@@ -1,33 +1,0 @@
-import numpy as np
-import numpy.linalg as la
-
-import scipy.optimize as sopt
-
-import matplotlib.pyplot as plt
-from mpl_toolkits.mplot3d import axes3d
-
-
-def f_rosenbrock(a, b):
-    return 100 * (b - a ** 2) ** 2 + (1 - a) ** 2
-
-
-def fd_rosenbrock(a, b):
-    fd_x = 400 * (a**3 - b * a) + 2 * a - 2
-    fd_y = 200 * (b - a ** 2)
-    return np.array([fd_x, fd_y])
-
-
-fig = plt.figure()
-ax = fig.gca(projection='3d')
-
-xmesh, ymesh = np.mgrid[-2:2:50j, -2:2:50j]
-fmesh = f_rosenbrock(xmesh, ymesh)
-
-ax.contour3D(xmesh, ymesh, fmesh, 200, cmap='magma')
-
-ax.view_init(20, 35)
-fig.show()
-fig.show()
-
-
-

### 2019-02-25 codebase/MathOpt/ps2/nt.py

@@ -1,0 +1,56 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from mpl_toolkits.mplot3d import Axes3D
+from numpy.linalg import inv
+
+
+def f_rosenbrock(x):
+    return 100 * np.square(np.square(x[0]) - x[1]) + np.square(x[0] - 1)
+
+
+# first order derivatives of the function
+def grad_rosenbrock(x):
+    df1 = 400 * x[0] * (np.square(x[0]) - x[1]) + 2 * (x[0] - 1)
+    df2 = -200 * (np.square(x[0]) - x[1])
+    return np.array([df1, df2])
+
+
+def inv_hess_rosenbrock(x):
+    df11 = 1200 * np.square(x[0]) - 400 * x[1] + 2
+    df12 = -400 * x[0]
+    df21 = -400 * x[0]
+    df22 = 200
+    hess = np.array([[df11, df12], [df21, df22]])
+    return inv(hess)
+
+
+def newton(x, max_int):
+    miter = 1
+    step = .5
+    vals = []
+    objectfs = []
+    # you can customize your own condition of convergence, here we limit the number of iterations
+    while miter <= max_int:
+        vals.append(x)
+        objectfs.append(f_rosenbrock(x))
+        temp = x - step * (inv_hess_rosenbrock(x).dot(grad_rosenbrock(x)))
+        if np.abs(f_rosenbrock(temp) - f_rosenbrock(x)) > 0.01:
+            x = temp
+        else:
+            break
+        print(x, f_rosenbrock(x), miter)
+        miter += 1
+    return vals, objectfs, miter
+
+
+start = [-1.2, 1]
+val, objectf, iters = newton(start, 50)
+
+x = np.array([i[0] for i in val])
+y = np.array([i[1] for i in val])
+z = np.array(objectf)
+fig = plt.figure()
+ax = fig.gca(projection='3d')
+ax.scatter(x, y, z, label='newton method')
+fig.show()
+fig.show()

### 2019-02-25 codebase/MathOpt/ps2/sd.py

@@ -1,0 +1,44 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from mpl_toolkits.mplot3d import Axes3D
+
+
+def f_rosenbrock(x):
+    return (1 - x[0]) ** 2 + 100 * (x[1] - x[0] ** 2) ** 2
+
+
+def grad_f_rosenbrock(x):
+    return np.array([2 * 100 * (x[1] - x[0] ** 2) * (-2 * x[0]) - 2 * (1. - x[0]), 2 * 100 * (x[1] - x[0] ** 2)])
+
+
+def grad(x, max_int):
+    miter = 1
+    step = .0001 / miter
+    vals = []
+    objectfs = []
+    while miter <= max_int:
+        vals.append(x)
+        objectfs.append(f_rosenbrock(x))
+        temp = x - step * grad_f_rosenbrock(x)
+        if np.abs(f_rosenbrock(temp) - f_rosenbrock(x)) > 0.01:
+            x = temp
+        else:
+            break
+        print(x, f_rosenbrock(x), miter)
+        miter += 1
+    return vals, objectfs, miter
+
+
+start = [-1.2, 1]
+val, objectf, iters = grad(start, 50)
+
+x = np.array([i[0] for i in val])
+y = np.array([i[1] for i in val])
+z = np.array(objectf)
+fig = plt.figure()
+ax = fig.gca(projection='3d')
+ax.scatter(x, y, z, label='gradient descent method', cmap='inferno')
+ax.legend()
+fig.show()
+fig.show()
+

### 2019-02-25 codebase/MathOpt/sd.py

@@ -1,47 +1,0 @@
-import numpy as np
-import numpy.linalg as la
-
-import scipy.optimize as sopt
-
-import matplotlib.pyplot as pt
-from mpl_toolkits.mplot3d import axes3d
-
-
-def f(x):
-    return 0.5*x[0]**2 + 2.5*x[1]**2
-
-
-def df(x):
-    return np.array([x[0], 5*x[1]])
-
-
-fig = pt.figure()
-ax = fig.gca(projection="3d")
-
-xmesh, ymesh = np.mgrid[-2:2:50j,-2:2:50j]
-fmesh = f(np.array([xmesh, ymesh]))
-ax.plot_surface(xmesh, ymesh, fmesh)
-
-pt.axis("equal")
-pt.contour(xmesh, ymesh, fmesh)
-
-guesses = [np.array([2, 2./5])]
-
-for i in range(100):
-    x = guesses[-1]
-    s = -df(x)
-
-    def f1d(alpha):
-        return f(x + alpha*s)
-
-    alpha_opt = sopt.golden(f1d)
-    next_guess = x + alpha_opt * s
-    guesses.append(next_guess)
-
-    print(next_guess)
-    pt.axis("equal")
-    pt.contour(xmesh, ymesh, fmesh, 50)
-    it_array = np.array(guesses)
-    pt.plot(it_array.T[0], it_array.T[1], "x-")
-    pt.figure()
-    pt.show()

### 2019-04-17 codebase/playground/pandas_cond_select.py
pg: conditional pandas selection

@@ -5,4 +5,3 @@ iris.head()
 iris
 
 rd = iris.loc[iris['sepal_length'] == 4.9]
-rd

### 2019-04-17 minutia/paper.tex

\documentclass[letterpaper,12pt]{article}
\usepackage[utf8]{inputenc}

% CSL style files for Mendeley and other referencing tools:
% https://github.com/citation-style-language/styles

%
% For alternative styles, see the biblatex manual:
% http://mirrors.ctan.org/macros/latex/contrib/biblatex/doc/biblatex.pdf
%
% The 'verbose' family of styles produces full citations in footnotes, 
% with and a variety of options for ibidem abbreviations.
%

\usepackage{graphicx}
\usepackage{csquotes}
\usepackage{amsfonts} % \mathbb in equations
% \usepackage[style=verbose-ibid,backend=bibtex]{biblatex}

\usepackage{cite}
% \usepackage{natbib}

\usepackage[dvipsnames]{xcolor}

\usepackage{array}
\newcolumntype{L}{>{\centering\arraybackslash}m{3cm}}

\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{caption}

\usepackage{amsmath} % \mathbb and \aligned
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

\newcommand{\norm}[1]{\left\lVert#1\right\rVert}

% \title{Markovian Approach to Data-driven Problems}
\title{Robust Deluding Adversarial Examples}

\author{Shayan Amani}

\begin{document}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
We propose a new algorithm to create robust adversarial examples to fool the decision-making models. We then show the amount of robustness we can achieve. Finally, an evaluation is provided to evaluate our proposed method.
\end{abstract}

\section{Introduction}
Deep reinforcement learning (Deep RL) achievements \cite{Mnih2013PlayingLearning} absorbed attentions to propose variety of algorithms \cite{Haarnoja2018SoftActor, Espeholt2018Impala:Architectures, Pong2018TemporalControl}. The promising methods ---similar to other aspects of machine learning--- have its own uncertainties which need to be addressed before the implementation. Robustness---as a property---elevates assurance about method performance and reliability to the foreground. We, in this paper, try to scrutinize state-of-the-art deep RL algorithms from a robustness point of view utilizing adversarial attacks in order to come up with a guarantee that we make sure which definitely works in the claimed criterion.

While the newly emerged methods are growing at a staggering pace, a parallel area of research has been formed wherein a variety of techniques have been utilized to make these methods contravene their promised functionality. These efforts would be served to develop robustness in those spectacular models and reduce their susceptibility to data tampering or dynamics of the environment.

\section{Related Work}
\citet{Goodfellow2014} stated and consequently depicted that the root cause of vulnerability to adversarial examples laid behind the piecewise linearity attributes of neural network models.

\section{Definition}
For the sake of clarity, we elaborate on some preliminary definitions here to have a consistent set of notations to the end of this paper.

\subsection{Adversarial attacks}
The designed attack may have different levels of access to the underlying training policy which have motivated the community to classify adversarial attacks into two distinct categories, namely \textit{white-box} or \textit{black-box} attacks. White-box attacks illustrating the scenario in which the... Adversarial attacks fall in either one of the following categories based upon the stage which in the adversary perturbs the model:
\begin{itemize}
    \item Train time: Data poisoning.
    \item Evaluation time: Adversarial examples.
    \item Deploy time: Black-box attacks.
\end{itemize}

\section{Devising Adversarial Attacks}
Neural networks are vulnerable to adversarial attacks either in supervised learning or reinforcement learning settings \cite{Huang2017}. Such weakness could possibly result in unrecognizable differences for the unaided eye in the presence of perturbation which leads to misclassification for neural network based model \cite{Szegedy2014, Goodfellow2014}.

\subsection{Objective}
The ultimate goal is deceiving the model either in choosing the worst possible action in an RL task or misclassifying an input in a classification task in a way which is not perceptible for unaided eye. In terms of a Markovian RL task we show how minimum set of worst-case actions are needed to result in the minimum final reward.

\subsection{Attacks Based upon Fast Gradient Sign Method}

The main goal of adversarial training and examples can be expressed briefly in form of the following equation:
\begin{equation}
\begin{aligned}
& \underset{\theta}{\min}
& & \sum_{x,y \in S} \underset{\delta \in \Delta}{\max}
& & & \textbb{Loss}(x + \delta, y; \theta)
\end{aligned}
\end{equation}

where in we trying to find an adversarial example (inner maximization) in order to adversarially train a robust model (outer minimization).

We define $s_\eta$ as the perturbed observation:
\begin{equation}
    s_\eta = (1 + \epsilon) s
\end{equation}

Let $s_\eta$ be the perturbed observation and $s_c$ scaled observation, so we are trying to find the following:
\begin{equation}
\begin{aligned}
& \underset{s}{\argmin}
& & {\norm{s_\eta - s_c}}
\end{aligned}
\end{equation}
which is the state observation that have the minimum distance between real-world [scaled] observation and the perturbed one.

In order to expand the applicability of designed adversarial attacks to an assortment of algorithms [models], we rely on \textit{transferability property} of RL-specific adversarial attacks \cite{Szegedy2014, Papernot2016, Goodfellow2014a}. \citet{Huang2017} clearly assert [or asserts?!] attacks are deployable to a group of trained RL models aimed at the same task.

\section{Adversarial attacks on different methodologies}
Here we review adversarial attacks on a variety of value function based state-of-art algorithms proposed by the community so far. As \citet{Pattanaik2018} pointed out, the concept of achieving a successful adversarial attack is unalike in reinforcement learning tasks compared to image classification tasks. Worst \textit{action} is tangible and feasible but not worst \textit{image}.

\subsection{Deep Q Learning (DQN)}
Utilizing experience replay and the leveraging target network, DQN take advantage of them to gain more stability \cite{Pattanaik2018}.

\section{Experiment}
In order to achieve more robustness, we need to omit loosely correlated features in training the model.

\subsection{Benchmarks}
When it comes to benchmarking, an acute shortage of a comprehensive testbed is felt. Two available evaluation frameworks are as follows:
\begin{itemize}
    \item CleverHans \cite{Papernot2016}
    \item Robust ML 
\end{itemize}

\section{Future Works}
We developed the presented approach with having reproducibility in mind. It could be served as a launching pad for the future contributors who want to surf this newly emerged field deeper. Future ideas to pursue is a generative model to be able to generate adversaries from a network.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{nips}
\bibliography{library}

\end{document}

### 2019-05-04 minutia/paper.tex
Updates from Overleaf

@@
 In terms of a Markovian RL task we show how minimum set of worst-case actions are needed to result in the minimum final reward.
+We try to show that the model is not robust to other configuration of each benchmarks with different dynamics.
 
@@
 
+\subsection{Reward Perturbation Attacks}
+In an adversarial environment we try to perturb reward to interfere in learning process of the RL agent.
+We considered this attack also in scenarios with a single reward which we will show the process later.
+
 \section{Adversarial attacks on different methodologies}

### 2019-05-06 codebase/adrel/simulate.py
adrel: noise introduction to observations added

@@ -9,6 +9,35 @@ import matplotlib.pyplot as plt
 import seaborn as sns
 
 
+"""
+     Observation: 
+        Type: Box(4)
+        Num	Observation                 Min         Max
+        0	Cart Position             -4.8            4.8
+        1	Cart Velocity             -Inf            Inf
+        2	Pole Angle                 -24 deg        24 deg
+        3	Pole Velocity At Tip      -Inf            Inf
+
+    Actions:
+        Type: Discrete(2)
+        Num	Action
+        0	Push cart to the left
+        1	Push cart to the right
+
+        Note: The amount the velocity that is reduced or increased is not fixed; it depends on the angle the pole is pointing. This is because the center of gravity of the pole increases the amount of energy needed to move the cart underneath it
+    Reward:
+        Reward is 1 for every step taken, including the termination step
+    Starting State:
+        All observations are assigned a uniform random value in [-0.05..0.05]
+    Episode Termination:
+        Pole Angle is more than 12 degrees
+        Cart Position is more than 2.4 (center of the cart reaches the edge of the display)
+        Episode length is greater than 200
+    Solved Requirements:
+        Considered solved when the average reward is greater than or equal to 195.0 over 100 consecutive trials.
+    """
+
+
 def select_action(s):
     state_q = q_values.loc[q_values['idstate'] == s]
     a = state_q.loc[state_q['qvalue'] == state_q.max()[2]]['idaction']
@@ -28,7 +57,7 @@ scales = pd.read_csv("qvals/scales.csv").values  # converting to an ndarray
 # policy = pd.read_csv("codebase/adrel/marek/policy_nn.csv")
 # scales = pd.read_csv("codebase/adrel/marek/scales.csv").values
 
-statefeatures = policy.values[:, 0:4]
+state_features = policy.values[:, 0:4]
 stateids = np.array(policy["State"])
 actions = np.array(policy["Action"])
 values = np.array(policy["Value"])
@@ -43,34 +72,33 @@ else:
     probabilities = None
 
 # number of runs to determine how good is the policy
-trials = 1000
+trials = 100
+eps = 1
 is_q = True
-total_reward = 0
 
-# n = 0
-# m = 0
 rewards = dict()
 info_list = []
+total_reward = 0
 
 # will not include sleep time as time.perf_counter will
 t_perf = time.perf_counter()
 t_elapsed = time.process_time()
 
-eps = 0.09
-
 for trial in tqdm.trange(trials):
     env.reset()
[…]

### 2019-05-06 codebase/GradDesc_OptAlgs/comparison.py
gradient descent optimization algorithms comparison added

@@ -1,0 +1,600 @@
+import numpy as np
+from sklearn.utils import shuffle
+from tensorflow.examples.tutorials.mnist import input_data
+import matplotlib.pyplot as plt
+import matplotlib
+
+np.random.seed(42)
+
+'''
+Implementation of 8 GD algorithms: 
+a. Stochastic Gradient Descent (One Line)
+b. Momentum
+c. Nesterov accelerated gradient (NAG)
+d. AdaGrad
+e. AdaDelta
+f. RMSprop
+g. Adam
+h. NAdam
+'''
+
+
+def log(x):
+    return 1 / (1 + np.exp(-1 * x))
+
+
+def d_log(x):
+    return log(x) * (1 - log(x))
+
+
+def tanh(x):
+    return np.tanh(x)
+
+
+def d_tanh(x):
+    return 1 - np.tanh(x) ** 2
+
+
+def ReLu(x):
+    mask = (x > 0.0) * 1.0
+    return x * mask
+
+
+def d_ReLu(x):
+    mask = (x > 0.0) * 1.0
+    return mask
+
+
+def elu(matrix):
+    mask = (matrix <= 0) * 1.0
+    less_zero = matrix * mask
+    safe = (matrix > 0) * 1.0
+    greater_zero = matrix * safe
+    final = 3.0 * (np.exp(less_zero) - 1) * less_zero
+    return greater_zero + final
+
+
+def d_elu(matrix):
+    safe = (matrix > 0) * 1.0
+    mask2 = (matrix <= 0) * 1.0
+    temp = matrix * mask2
+    final = (3.0 * np.exp(temp)) * mask2
+    return (matrix * safe) + final
+
+
+# 0. Declare Training Data and Labels
+mnist = input_data.read_data_sets("../MNIST_data/", one_hot=False)
+train = mnist.test
+images, labels = train.images, train.labels
+
+only_zero_index, only_one_index = np.where(labels == 0)[0], np.where(labels == 1)[0]
+only_zero_image, only_zero_label = images[[only_zero_index]], np.expand_dims(labels[[only_zero_index]], axis=1)
+only_one_image, only_one_label = images[[only_one_index]], np.expand_dims(labels[[only_one_index]], axis=1)
+
+images = np.vstack((only_zero_image, only_one_image))
+labels = np.vstack((only_zero_label, only_one_label))
+images, label = shuffle(images, labels)
+
[…]

### 2019-05-06 codebase/GradDesc_OptAlgs/comparison.py
GD algorithms updated comparison

@@ -581,7 +581,7 @@ for iter in range(num_epoch):
         w1_nadam = w1_nadam - (learn_rate / (np.sqrt(Nadam_v_1_hat) + Nadam_e)) * (
                 Nadam_Beta_1 * Nadam_m_1_hat + (((1 - Nadam_Beta_1) * grad_1) / (1 - Nadam_Beta_1)))
     if iter % 10 == 0:
-        print("h. Nadam current Iter: ", iter, " Total Cost: ", total_cost)
+        print("h. NAdam current Iter: ", iter, " Total Cost: ", total_cost)
     cost_temp_array.append(total_cost)
     total_cost = 0
 cost_array.append(cost_temp_array)

### 2019-05-07 minutia/MathOpt-report/main.tex
MathOpt: report added.

\documentclass[12pt]{article}

\usepackage[utf8]{inputenc}

\usepackage[dvipsnames]{xcolor}

\usepackage{array}
\newcolumntype{L}{>{\centering\arraybackslash}m{3cm}}

\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{caption}

\usepackage{graphicx}
\usepackage{csquotes}

\renewcommand{\bibname}{References}

\title{Mathematical Optimization\\Final Project Report}

\author{Shayan Amani}

\date{\today}

\begin{document}
\maketitle

\section{Introduction}
Among available optimization algorithms, group of gradient descent have shown better performance in case of large data sets. This project is to compare these set of algorithms with a focus on understanding the approach of each.

\section{Motivation}
Advanced and popular differentiable programming frameworks are using gradient descent optimization variations at the core level. By exploring the documentation of these frameworks, it is visible that a specific set of the algorithms are in common. Understanding and getting a hands-on experience in implementation of these algorithms was the key factor to select this topic as my final project.

\section{Algorithms}
Gradient descent as an optimization technique includes three variations:
\begin{itemize}
    \item batch gradient descent,
    \item stochastic gradient descent, and 
    \item mini-batch gradient descent.
\end{itemize}
The examined algorithms are as follows:
\begin{itemize}
    \item Momentum \cite{NQianOnAlgorithms}
    \item Nesterov Accelerated Gradient \cite{NESTEROV1983AO1/k2}
    \item AdaGrad \cite{DuchiAdaptiveOptimization}
    \item AdaDelta \cite{Zeiler2012ADADELTA:Method}
    \item RMSprop \cite{HintonNeuralDescent}
    \item Adam \cite{KingmaAdam:Optimization}
    \item NAdam \cite{Dozat2016IncorporatingAdam}
\end{itemize}

\section{Implementation}
I have implemented the algorithm in python using numpy package. My idea is mimicking a set of neural network layers by only incorporating activation functions. The functions such as log, tanh, elu are considered here to test out the algorithms. MNIST dataset is used to calculate error in learning (cost function) by sequentially calculating functions value for each image of the dataset and plugging them into the algorithms' update equation.

\section{Results}
As figure \ref{fig:algorithms_cost_function} shows, the cost (error) of learning during 100 epochs varies by using a differnet algorithm. AdaDelta went to the no-return point and indicates \textit{no free lunch theorem} for optimization.
\begin{figure}
    \centering
    \includegraphics[width=1\columnwidth]{presentation/MathOpt-proj/myplot.png}
    \caption{Minimizing the error function in learning for 100 trials calculated using different algorithms.}
    \label{fig:algorithms_cost_function}
\end{figure}

\bibliographystyle{plain}
\bibliography{library}

\end{document}

### 2019-05-08 codebase/GradDesc_OptAlgs/comparison.py
simplex updated

@@ -1,3 +1,7 @@
+# Gradient Descent-based algorithms comparison
+# Shayan Amani
+# http://shayanamani.com
+
 import numpy as np
 from sklearn.utils import shuffle
 from tensorflow.examples.tutorials.mnist import input_data

### 2019-05-08 codebase/adrel/qlearn.py

@@ -1,52 +1,0 @@
-import random
-
-class QLearn:
-    def __init__(self, actions, epsilon, alpha, gamma):
-        self.q = {}
-        self.epsilon = epsilon  # exploration constant
-        self.alpha = alpha  # discount constant
-        self.gamma = gamma  # discount factor
-        self.actions = actions
-
-    def get_q(self, state, action):
-        return self.q.get((state, action), 0.0)
-
-    def learn_q(self, state, action, reward, value):
-        '''
-        Q-learning:
-            Q(s, a) += alpha * (reward(s,a) + max(Q(s') - Q(s,a))
-        '''
-        oldv = self.q.get((state, action), None)
-        if oldv is None:
-            self.q[(state, action)] = reward
-        else:
-            self.q[(state, action)] = oldv + self.alpha * (value - oldv)
-
-    def chooseAction(self, state, return_q=False):
-        q = [self.get_q(state, a) for a in self.actions]
-        maxQ = max(q)
-
-        if random.random() < self.epsilon:
-            minQ = min(q);
-            mag = max(abs(minQ), abs(maxQ))
-            # add random values to all the actions, recalculate maxQ
-            q = [q[i] + random.random() * mag - .5 * mag for i in range(len(self.actions))]
-            maxQ = max(q)
-
-        count = q.count(maxQ)
-        # In case there're several state-action max values
-        # we select a random one among them
-        if count > 1:
-            best = [i for i in range(len(self.actions)) if q[i] == maxQ]
-            i = random.choice(best)
-        else:
-            i = q.index(maxQ)
-
-        action = self.actions[i]
-        if return_q:  # if they want it, give it!
-            return action, q
-        return action
-
-    def learn(self, state1, action1, reward, state2):
-        maxqnew = max([self.get_q(state2, a) for a in self.actions])
-        self.learn_q(state1, action1, reward, reward + self.gamma * maxqnew)

### 2019-05-08 codebase/adrel/player.py
adrel: player class reorg

@@ -1,0 +1,101 @@
+import numpy as np
+import gym
+import tqdm
+import time
+
+
+class CartPolePlayer:
+
+    def __init__(self, policy, scale, q_values=None, render=False):
+        self.policy = policy
+        self.scale = scale
+        self.q_values = q_values
+
+        self.state_features = policy.values[:, 0:4]
+        self.state_ids = np.array(policy["State"])
+        self.actions = np.array(policy["Action"])
+        self.values = np.array(policy["Value"])
+        if "Probability" in policy.columns:
+            print("Using a randomized policy")
+            self.policy_randomized = True
+            self.probabilities = np.array(policy["Probability"])
+        else:
+            print("Using a deterministic policy")
+            self.policy_randomized = False
+            self.probabilities = None
+
+        self.render = render
+
+    def select_action(self, state):
+        state_q = self.q_values.loc[self.q_values['idstate'] == state]
+        action = state_q.loc[state_q['qvalue'] == state_q.max()[2]]['idaction']
+        return int(action)
+
+    def play(self, trials, eps):
+
+        # gym/envs/__init__.py'
+        # different max_episode_steps 'CartPole-v0' and 'CartPole-v1'
+        env = gym.make('CartPole-v1')
+        max_episode_steps = 200
+
+        total_rewards = dict()
+
+        for trial in tqdm.trange(trials):
+            env.reset()
+            done = False
+
+            # take an initial random action and observe the state
+            action = env.action_space.sample()
+
+            [state, reward, done, info] = env.step(action)
+            if done:
+                continue
+
+            for i in range(1, max_episode_steps):
+                # faster with no graphical output
+                if self.render:
+                    env.render()
+                    time.sleep(0.05)
+
+                # scale the observed state
+                state_scaled = state @ self.scale
+                # TODO: multiplication w/o eye matrix
+                perturbation = np.eye(self.scale.shape[0]) * (1 + eps)
+                state_scaled = state_scaled @ perturbation
+                # ||(Sc + eps) - Sc||
+                distance = np.linalg.norm(
+                    self.state_features - np.repeat(np.atleast_2d(state_scaled), self.state_features.shape[0], 0),
+                    axis=1)
+                # find the closest state to the observed state in the policy
+                # state index in the file, not the state number
+                state_index = np.argmin(distance)
+
+                if self.policy_randomized:
+                    id_state = self.state_ids[state_index]
+                    # find all relevant state ids
[…]

### 2019-05-08 codebase/adrel/simulate.py

@@ -1,13 +1,10 @@
-import time
-
-import gym
 import numpy as np
 import pandas as pd
-import tqdm
 
 import matplotlib.pyplot as plt
 import seaborn as sns
 
+from player import CartPolePlayer
 
 """
      Observation: 
@@ -35,133 +32,42 @@ import seaborn as sns
         Episode length is greater than 200
     Solved Requirements:
         Considered solved when the average reward is greater than or equal to 195.0 over 100 consecutive trials.
-    """
+"""
 
 
-def select_action(s):
-    state_q = q_values.loc[q_values['idstate'] == s]
-    a = state_q.loc[state_q['qvalue'] == state_q.max()[2]]['idaction']
-    return int(a)
+if __name__ == "__main__":
 
+    policies = [{'dir': 'initial', 'q': False},
+                {'dir': 'qvals', 'q': True}]
+    files_name = {'policy': '/policy_nn.csv', 'scale': '/scales.csv', 'q_values': '/qvalues.csv'}
 
-env = gym.make('CartPole-v1')
+    trials = 1000
+    render = False
+    epsilon = 1
 
-# policy = pd.read_csv("initial/policy_nn.csv")
-policy = pd.read_csv("qvals/policy_nn.csv")
+    for p in policies:
+        path_policy = p['dir'] + files_name['policy']
+        path_scale = p['dir'] + files_name['scale']
+        path_q_values = p['dir'] + files_name['q_values']
 
-q_values = pd.read_csv("qvals/qvalues.csv")
+        policy = pd.read_csv(path_policy)
+        scale = pd.read_csv(path_scale).values
+        q_values = pd.read_csv(path_q_values) if p['q'] else None
 
-# scales = pd.read_csv("initial/scales.csv").values # converting to an ndarray
-scales = pd.read_csv("qvals/scales.csv").values  # converting to an ndarray
+        cart_pole = CartPolePlayer(policy, scale, q_values, render)
+        total_rewards = cart_pole.play(trials, epsilon)
 
-# policy = pd.read_csv("codebase/adrel/marek/policy_nn.csv")
-# scales = pd.read_csv("codebase/adrel/marek/scales.csv").values
+        average_reward = sum(total_rewards.values()) / trials
+        print("Average reward per trial", average_reward)
 
-state_features = policy.values[:, 0:4]
-stateids = np.array(policy["State"])
-actions = np.array(policy["Action"])
-values = np.array(policy["Value"])
[…]

### 2019-05-08 minutia/paper.tex

@@
 
-\section{Devising Adversarial Attacks}
-Neural networks are vulnerable to adversarial attacks either in supervised learning or reinforcement learning settings \cite{Huang2017}.
-Such weakness could possibly result in unrecognizable differences for the unaided eye in the presence of perturbation which leads to misclassification for neural network based model \cite{Szegedy2014, Goodfellow2014}.
-
 \subsection{Objective}
@@
 We try to show that the model is not robust to other configuration of each benchmarks with different dynamics.
-
-\subsection{Attacks Based upon Fast Gradient Sign Method}
 
@@
 
+\section{Devising Adversarial Attacks}
+Neural networks are vulnerable to adversarial attacks either in supervised learning or reinforcement learning settings \cite{Huang2017}.
+Such weakness could possibly result in unrecognizable differences for the unaided eye in the presence of perturbation which leads to misclassification for neural network based model \cite{Szegedy2014, Goodfellow2014}.
+
+\subsection{Observation-perturbed attacks}
+
 where in we trying to find an adversarial example (inner maximization) in order to adversarially train a robust model (outer minimization).
@@
 \begin{equation}
-    s_\eta = (1 + \epsilon) s
+    s_\eta = s . (1 + \epsilon) 
 \end{equation}
@@
 & \underset{s}{\argmin}
-& & {\norm{s_\eta - s_c}}
+& & {\norm{s_\eta - s_c}} = \norm{s . (1 + \eps) - s_c}
 \end{aligned}
@@
 
-\subsection{Reward Perturbation Attacks}
+\subsection{Reward-Perturbed attacks}
 In an adversarial environment we try to perturb reward to interfere in learning process of the RL agent.

### 2019-05-18 codebase/adrel/agent_dqn.py
adrel: agent dqn and policy added. reorganized.

@@ -1,0 +1,73 @@
+import numpy as np
+import random
+from collections import deque
+
+from keras.models import Sequential
+from keras.layers import Dense
+from keras.optimizers import Adam
+
+
+class AgentDQN:
+
+    def __init__(self, state_size, action_size):
+        self.state_size = state_size
+        self.action_size = action_size
+        self.memory = deque(maxlen=2000)  # arbitrary size
+
+        self.gamma = 0.95  # reward discount rate
+
+        # exploration rates
+        self.epsilon = 1.0
+        self.epsilon_decay = 0.995
+        self.epsilon_min = 0.01
+
+        self.alpha = 0.001  # learning rate
+
+        self.model = self._build_model()
+
+    def _build_model(self):
+        model = Sequential()
+        # first argument of Dense is dimension of output.
+        # last layer's output dimension makes more sense
+        model.add(Dense(24, input_dim=self.state_size, activation='relu'))
+        model.add(Dense(24, activation='relu'))
+        model.add(Dense(self.action_size, activation='linear'))
+        model.compile(loss='mse', optimizer=Adam(lr=self.alpha))
+        return model
+
+    def remember(self, state, action, reward, next_state, done):
+        self.memory.append((state, action, reward, next_state, done))
+
+    # exploration - exploitation
+    # returns a random exploratory action or
+    # a predicted action from the network
+    def act(self, state):
+        if np.random.rand() <= self.epsilon:
+            return random.randrange(self.action_size)
+        # TODO: try to see the output shape
+        predicted_action_values = self.model.predict(state)[0]
+        return np.argmax(predicted_action_values)  # returns action
+
+    def replay(self, batch_size):
+        mini_batch = random.sample(self.memory, batch_size)
+        for state, action, reward, next_state, done in mini_batch:
+            target = reward
+            if not done:
+                next_action_values = self.model.predict(next_state)[0]
+                # q_target = r + gamma * q_s'
+                target = (reward + self.gamma * np.amax(next_action_values))
+            # updates q network
+            target_f = self.model.predict(state)
+            target_f[0][action] = target
+            # learn the experience
+            self.model.fit(state, target_f, epochs=1, verbose=0)
+        # decay exploration and towards exploitation
+        if self.epsilon > self.epsilon_min:
+            self.epsilon *= self.epsilon_decay
+
+    # TODO: change to a better arg name
+    def load(self, name):
+        self.model.load_weights(name)
+
+    def save(self, name):
+        self.model.save_weights(name)

### 2019-05-18 codebase/adrel/agent_policy.py

@@ -1,0 +1,63 @@
+import numpy as np
+import pandas as pd
+
+FILES_NAME = {'policy': '/policy_nn.csv',
+              'scale': '/scales.csv',
+              'q_values': '/qvalues.csv'}
+
+
+class AgentPolicy:
+
+    def __init__(self, policy_info):
+        self.policy, self.scale, self.q_values = self.load_policy(policy_info['dir'],
+                                                                  policy_info['q'])
+
+        self.state_features = self.policy.values[:, 0:4]
+        self.state_ids = np.array(self.policy['State'])
+        self.actions = np.array(self.policy['Action'])
+        self.values = np.array(self.policy['Value'])
+        if 'Probability' in self.policy.columns:
+            print('Stochastic policy loaded')
+            self.probabilities = np.array(self.policy['Probability'])
+        else:
+            print('Deterministic policy loaded')
+            self.probabilities = None
+
+    def load_policy(self, directory, is_q):
+        policy = pd.read_csv(directory + FILES_NAME['policy'])
+        scale = pd.read_csv(directory + FILES_NAME['scale']).values
+        q_values = pd.read_csv(directory + FILES_NAME['q_values']) if is_q else None
+        return policy, scale, q_values
+
+    def select_state(self, state, pert_noise):
+        # scale the observed state
+        state_scaled = state @ self.scale
+        # TODO: multiplication w/o eye matrix
+        perturbed = np.eye(self.scale.shape[0]) * (1 + pert_noise)
+        state_scaled = state_scaled @ perturbed
+        # ||(Sc + eps) - Sc||
+        distance = np.linalg.norm(
+            self.state_features - np.repeat(np.atleast_2d(state_scaled), self.state_features.shape[0], 0),
+            axis=1)
+        # find the closest state to the observed state in the policy
+        # state index in the file, not the state number
+        return np.argmin(distance)
+
+    def select_action(self, state_index):
+        if self.probabilities is not None:
+            # find all relevant state ids
+            all_state_ids = np.where(self.state_ids == self.state_ids[state_index])[0]
+            all_probs = self.probabilities[all_state_ids]
+            all_actions = self.actions[all_state_ids]
+            assert (abs(1 - sum(all_probs)) < 0.01)
+            action = int(np.random.choice(all_actions, p=all_probs))
+        else:
+            # assume that there is a single action for each state
+            # switch between q-values or the policy
+            if self.q_values is not None:
+                # TODO: np.where rather than .loc[]
[…]

### 2019-05-18 codebase/adrel/player.py

@@ -1,101 +1,0 @@
-import numpy as np
-import gym
-import tqdm
-import time
-
-
-class CartPolePlayer:
-
-    def __init__(self, policy, scale, q_values=None, render=False):
-        self.policy = policy
-        self.scale = scale
-        self.q_values = q_values
-
-        self.state_features = policy.values[:, 0:4]
-        self.state_ids = np.array(policy["State"])
-        self.actions = np.array(policy["Action"])
-        self.values = np.array(policy["Value"])
-        if "Probability" in policy.columns:
-            print("Using a randomized policy")
-            self.policy_randomized = True
-            self.probabilities = np.array(policy["Probability"])
-        else:
-            print("Using a deterministic policy")
-            self.policy_randomized = False
-            self.probabilities = None
-
-        self.render = render
-
-    def select_action(self, state):
-        state_q = self.q_values.loc[self.q_values['idstate'] == state]
-        action = state_q.loc[state_q['qvalue'] == state_q.max()[2]]['idaction']
-        return int(action)
-
-    def play(self, trials, eps):
-
-        # gym/envs/__init__.py'
-        # different max_episode_steps 'CartPole-v0' and 'CartPole-v1'
-        env = gym.make('CartPole-v1')
-        max_episode_steps = 200
-
-        total_rewards = dict()
-
-        for trial in tqdm.trange(trials):
-            env.reset()
-            done = False
-
-            # take an initial random action and observe the state
-            action = env.action_space.sample()
-
-            [state, reward, done, info] = env.step(action)
-            if done:
-                continue
-
-            for i in range(1, max_episode_steps):
-                # faster with no graphical output
-                if self.render:
-                    env.render()
-                    time.sleep(0.05)
-
-                # scale the observed state
-                state_scaled = state @ self.scale
-                # TODO: multiplication w/o eye matrix
-                perturbation = np.eye(self.scale.shape[0]) * (1 + eps)
-                state_scaled = state_scaled @ perturbation
-                # ||(Sc + eps) - Sc||
-                distance = np.linalg.norm(
-                    self.state_features - np.repeat(np.atleast_2d(state_scaled), self.state_features.shape[0], 0),
-                    axis=1)
-                # find the closest state to the observed state in the policy
-                # state index in the file, not the state number
-                state_index = np.argmin(distance)
-
-                if self.policy_randomized:
-                    id_state = self.state_ids[state_index]
-                    # find all relevant state ids
[…]

### 2019-05-18 codebase/adrel/simulate.py

@@ -1,12 +1,29 @@
 import numpy as np
-import pandas as pd
+import tqdm
+
+import gym
 
 import matplotlib.pyplot as plt
 import seaborn as sns
 
-from player import CartPolePlayer
+from agent_policy import AgentPolicy
+from agent_dqn import AgentDQN
 
-"""
+# gym/envs/__init__.py'
+# different max_episode_steps 'CartPole-v0' and 'CartPole-v1'
+GYM_ENV = 'CartPole-v1'
+EPISODES = 1000
+MAX_EPISODE_STEPS = 200
+BATCH_SIZE = 32
+
+
+IS_RENDER = False
+
+POLICIES = [{'dir': 'initial', 'q': False},
+            {'dir': 'qvals', 'q': True}]
+
+""" CartPole-v0:
+
      Observation: 
         Type: Box(4)
         Num	Observation                 Min         Max
@@ -31,43 +48,95 @@ from player import CartPolePlayer
         Cart Position is more than 2.4 (center of the cart reaches the edge of the display)
         Episode length is greater than 200
     Solved Requirements:
-        Considered solved when the average reward is greater than or equal to 195.0 over 100 consecutive trials.
+        Considered solved when the average reward is greater than or equal to 195.0 over 100 consecutive EPISODES.
 """
+
+
+def play_dqn():
+    env = gym.make(GYM_ENV)
+    state_size = env.observation_space.shape[0]
+    action_size = env.action_space.n
+    agent = AgentDQN(state_size, action_size)
+    # agent.load('./model/cartpole-dqn.h5')
+
+    game_reward = dict()
+    for episode in range(EPISODES):
+        state = env.reset()
+        state = np.reshape(state, [1, state_size])
+        for step in range(MAX_EPISODE_STEPS):
+            if IS_RENDER:
+                env.render()
+            action = agent.act(state)
+            next_state, reward, done, _ = env.step(action)
+            reward = reward if not done else -10
+            next_state = np.reshape(next_state, [1, state_size])
+            agent.remember(state, action, reward, next_state, done)
+            state = next_state
+            if done:
+                print("episode: {}/{}, score: {}, e: {:.2}"
+                      .format(episode, EPISODES, step, agent.epsilon))
+                game_reward[episode] = step + 1
+                break
+            if len(agent.memory) > BATCH_SIZE:
+                agent.replay(BATCH_SIZE)
+            if step == MAX_EPISODE_STEPS - 1:
[…]

### 2019-05-18 codebase/GradDesc_OptAlgs/comparison.py
GradDesc: algorithm comparison added

@@ -240,7 +240,7 @@ for iter in range(num_epoch):
         grad_1_part_3 = current_image
         grad_1 = grad_1_part_3.T.dot(grad_1_part_1 * grad_1_part_2)
 
-        # ------- FAKE GRADIENT --------
+        # -------  --------
         fake_w3_ng = w3_ng - alpha * v3
         fake_w2_ng = w2_ng - alpha * v2
         fake_w1_ng = w1_ng - alpha * v1
@@ -268,7 +268,7 @@ for iter in range(num_epoch):
         grad_1_part_2 = d_elu(l1)
         grad_1_part_3 = current_image
         grad_1_fake = grad_1_part_3.T.dot(grad_1_part_1 * grad_1_part_2)
-        # ------- FAKE GRADIENT --------
+        # -------  --------
 
         v3 = v3 * alpha + learn_rate * grad_3_fake
         v2 = v2 * alpha + learn_rate * grad_2_fake

### 2019-10-24 minutia/archive/StateAggregation-survey/survey.tex
Latex minutia folder reorganized - migrated from Overleaf

--- 2019-04-17 minutia/paper.tex
@@
 
-% \title{Markovian Approach to Data-driven Problems}
-\title{Robust Deluding Adversarial Examples}
+\title{State Aggregation in Robust MDPs, A Quick Survey}
 
-\author{Shayan Amani}
+% \author{Shayan Amani}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{abstract}
-We propose a new algorithm to create robust adversarial examples to fool the decision-making models.
-We then show the amount of robustness we can achieve.
-Finally, an evaluation is provided to evaluate our proposed method.
-\end{abstract}
+As in real world scenarios, handling large state-spaces is a challenge in reinforcement learning.
+Value functions resulted from such problems could be solved by set of approximation methods to yield the policy.
+One of those methods is state aggregation which maps states in a vicinity to a certain action.
 
-\section{Introduction}
-Deep reinforcement learning (Deep RL) achievements \cite{Mnih2013PlayingLearning} absorbed attentions to propose variety of algorithms \cite{Haarnoja2018SoftActor, Espeholt2018Impala:Architectures, Pong2018TemporalControl}.
-The promising methods ---similar to other aspects of machine learning--- have its own uncertainties which need to be addressed before the implementation.
-Robustness---as a property---elevates assurance about method performance and reliability to the foreground.
-We, in this paper, try to scrutinize state-of-the-art deep RL algorithms from a robustness point of view utilizing adversarial attacks in order to come up with a guarantee that we make sure which definitely works in the claimed criterion.
+\section{Related works}
 
-While the newly emerged methods are growing at a staggering pace, a parallel area of research has been formed wherein a variety of techniques have been utilized to make these methods contravene their promised functionality.
-These efforts would be served to develop robustness in those spectacular models and reduce their susceptibility to data tampering or dynamics of the environment.
+\subsection{RAAM}
+Give a \textbf{tighter} bound on performance loss by reducing sensitivity to approximation error $\epsilon(.)$.
 
-\section{Related Work}
-\citet{Goodfellow2014} stated and consequently depicted that the root cause of vulnerability to adversarial examples laid behind the piecewise linearity attributes of neural network models.
-
-\section{Definition}
-For the sake of clarity, we elaborate on some preliminary definitions here to have a consistent set of notations to the end of this paper.
-
-\subsection{Adversarial attacks}
-The designed attack may have different levels of access to the underlying training policy which have motivated the community to classify adversarial attacks into two distinct categories, namely \textit{white-box} or \textit{black-box} attacks.
-White-box attacks illustrating the scenario in which the...
-Adversarial attacks fall in either one of the following categories based upon the stage which in the adversary perturbs the model:
+\subsection{Kernel-based RL in RMDP}
+Robust policy learning in case of problems with massive or continuous state-spaces. By using \textit{kernel averagers} --as a function approximator-- to:
     \begin{itemize}
-    \item Train time: Data poisoning.
-    \item Evaluation time: Adversarial examples.
-    \item Deploy time: Black-box attacks.
+        \item approximate reward function $r$ and/or transition function $P$ based on training state-action samples.
+        \item approximate the value function $\nu$.
     \end{itemize}
-
-\section{Devising Adversarial Attacks}
-Neural networks are vulnerable to adversarial attacks either in supervised learning or reinforcement learning settings \cite{Huang2017}.
-Such weakness could possibly result in unrecognizable differences for the unaided eye in the presence of perturbation which leads to misclassification for neural network based model \cite{Szegedy2014, Goodfellow2014}.
-
-\subsection{Objective}
-The ultimate goal is deceiving the model either in choosing the worst possible action in an RL task or misclassifying an input in a classification task in a way which is not perceptible for unaided eye.
-In terms of a Markovian RL task we show how minimum set of worst-case actions are needed to result in the minimum final reward.
-
-\subsection{Attacks Based upon Fast Gradient Sign Method}
-
-The main goal of adversarial training and examples can be expressed briefly in form of the following equation:
-\begin{equation}
-\begin{aligned}
-& \underset{\theta}{\min}
-& & \sum_{x,y \in S} \underset{\delta \in \Delta}{\max}
-& & & \textbb{Loss}(x + \delta, y; \theta)
-\end{aligned}
-\end{equation}
-
-where in we trying to find an adversarial example (inner maximization) in order to adversarially train a robust model (outer minimization).
-
-We define $s_\eta$ as the perturbed observation:
-\begin{equation}
-    s_\eta = (1 + \epsilon) s
-\end{equation}
-
-Let $s_\eta$ be the perturbed observation and $s_c$ scaled observation, so we are trying to find the following:
-\begin{equation}
-\begin{aligned}
-& \underset{s}{\argmin}
-& & {\norm{s_\eta - s_c}}
-\end{aligned}
-\end{equation}
-which is the state observation that have the minimum distance between real-world [scaled] observation and the perturbed one.
-
-In order to expand the applicability of designed adversarial attacks to an assortment of algorithms [models], we rely on \textit{transferability property} of RL-specific adversarial attacks \cite{Szegedy2014, Papernot2016, Goodfellow2014a}.
-\citet{Huang2017} clearly assert [or asserts?!] attacks are deployable to a group of trained RL models aimed at the same task.
-
-\section{Adversarial attacks on different methodologies}
-Here we review adversarial attacks on a variety of value function based state-of-art algorithms proposed by the community so far.
-As \citet{Pattanaik2018} pointed out, the concept of achieving a successful adversarial attack is unalike in reinforcement learning tasks compared to image classification tasks.
-Worst \textit{action} is tangible and feasible but not worst \textit{image}.
-
-\subsection{Deep Q Learning (DQN)}
-Utilizing experience replay and the leveraging target network, DQN take advantage of them to gain more stability \cite{Pattanaik2018}.
-
-\section{Experiment}
-In order to achieve more robustness, we need to omit loosely correlated features in training the model.
-
-\subsection{Benchmarks}
-When it comes to benchmarking, an acute shortage of a comprehensive testbed is felt. Two available evaluation frameworks are as follows:
-\begin{itemize}
-    \item CleverHans \cite{Papernot2016}
-    \item Robust ML 
-\end{itemize}
-
-\section{Future Works}
-We developed the presented approach with having reproducibility in mind.
-It could be served as a launching pad for the future contributors who want to surf this newly emerged field deeper.
-Future ideas to pursue is a generative model to be able to generate adversaries from a network.
 
@@
 
-\bibliographystyle{nips}
-\bibliography{library}
+% \bibliographystyle{nips}
+% \bibliography{library}
 

### 2019-10-24 minutia/paper/paper.tex
Latex minutia folder reorganized - migrated from Overleaf

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2019

% ready for submission
% \usepackage{neurips_2019}

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
    \usepackage[preprint]{neurips_2019}

% to compile a camera-ready version, add the [final] option, e.g.:
    %  \usepackage[final]{neurips_2019}

% to avoid loading the natbib package, add option nonatbib:
%     \usepackage[nonatbib]{neurips_2019}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

 \usepackage[sort]{natbib}

\title{Feature Extraction and Aggregation}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.

\author{%
  Shayan Amani, Marek Petrik \\ %\thanks{Use footnote for providing further information
%    about author (webpage, alternative address)---\emph{not} for acknowledging
%    funding agencies.}%
  Department of Computer Science\\
  University of New Hampshire \\
  \texttt{\{shayan\}\{mpetrik\}@cs.unh.edu} \\
  % examples of more authors
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \AND
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
}

\begin{document}

\maketitle

\begin{abstract}
  Promising advances in deep reinforcement learning suggests several application to the algorithms. However robustness of these algorithms remains doubts to trust these algorithms in practice. In this paper we study resistance of the state-of-the-art algorithms to environment dynamics. We then show the amount of robustness we can achieve as a guarantee of performance in presence of those changes.
\end{abstract}

\section{Introduction}

Reassuring achievements in the Deep RL in a variety of fields absorb the attention from the community \citep{Mnih2015, Schulman2015TrustOptimization, Lillicrap2015ContinuousDDPG, silver2016a}. The cross fertilization in using approximation techniques by Deep Neural Networks in solving immense state space problems in RL is a key. While the newly emerged methods are growing at a staggering pace, a parallel area of research has been formed wherein a variety of techniques have been utilized to make these methods contravene their promised functionality. These efforts would be served to develop robustness in those spectacular models and reduce their susceptibility to data tampering or dynamics of the environment. Robustness ---as a property in machine learning--- elevates assurance about methods performance and their reliability to the foreground. In a few studies \cite{Goodfellow2014}  vulnerabilities of DNN's to adversarially-manipulated inputs are revealed. Therefore, robust algorithms, have to be of paramount importance in applying deep RL to real world scenarios to mitigate risks of implementation. Adversarial attacks as a way to perturbing these algorithms, are introduced in recent works \cite{Huang2017, Kos2017} from different aspects to degrade algorithms' performance. One of such potentials are changes in the environment dynamics.

In this work, we scrutinize state-of-the-art deep RL algorithms resiliency to environment parameters alterations and then introduced our approach to overcome those. Then a guarantee of performance to these attacks is introduced.

\section{Related Work}
\cite{Goodfellow2014} stated and consequently depicted that the root cause of vulnerability to adversarial examples laid behind the piecewise linearity attributes of neural network models.

\section{Definition}
For the sake of clarity, we elaborate on some preliminary definitions here to have a consistent set of notations to the end of this paper.

\subsection{Adversarial attacks}
The designed attack may have different levels of access to the underlying training policy which have motivated the community to classify adversarial attacks into two distinct categories, namely \textit{white-box} or \textit{black-box} attacks. White-box attacks illustrating the scenario in which the... Adversarial attacks fall in either one of the following categories based upon the stage which in the adversary perturbs the model:
\begin{itemize}
    \item Train time: Data poisoning.
    \item Evaluation time: Adversarial examples.
    \item Deploy time: Black-box attacks.
\end{itemize}

\subsection{Objective}
The ultimate goal is deceiving the model either in choosing the worst possible action in an RL task or misclassifying an input in a classification task in a way which is not perceptible for unaided eye. In terms of a Markovian RL task we show how minimum set of worst-case actions are needed to result in the minimum final reward. We try to show that the model is not robust to other configuration of each benchmarks with different dynamics.

The main goal of adversarial training and examples can be expressed briefly in form of the following equation:
\begin{equation}
\begin{aligned}
& \underset{\theta}{\min}
& & \sum_{x,y \in S} \underset{\delta \in \Delta}{\max}
& & & \textbb{Loss}(x + \delta, y; \theta)
\end{aligned}
\end{equation}

\section{Devising Adversarial Attacks}
Neural networks are vulnerable to adversarial attacks either in supervised learning or reinforcement learning settings \cite{Huang2017}. Such weakness could possibly result in unrecognizable differences for the unaided eye in the presence of perturbation which leads to misclassification for neural network based model \cite{Szegedy2014, Goodfellow2014}.

\subsection{Observation-perturbed attacks}

where in we trying to find an adversarial example (inner maximization) in order to adversarially train a robust model (outer minimization).

We define $s_\eta$ as the perturbed observation:
\begin{equation}
    s_\eta = s . (1 + \epsilon) 
\end{equation}

Let $s_\eta$ be the perturbed observation and $s_c$ scaled observation, so we are trying to find the following:
\begin{equation}
\begin{aligned}
& \underset{s}{\argmin}
& & {\norm{s_\eta - s_c}} = \norm{s . (1 + \eps) - s_c}
\end{aligned}
\end{equation}
which is the state observation that have the minimum distance between real-world [scaled] observation and the perturbed one.

In order to expand the applicability of designed adversarial attacks to an assortment of algorithms [models], we rely on \textit{transferability property} of RL-specific adversarial attacks \cite{Szegedy2014, Papernot2016, Goodfellow2014a}. \cite{Huang2017} clearly assert [or asserts?!] attacks are deployable to a group of trained RL models aimed at the same task.

\subsection{Reward-Perturbed attacks}
In an adversarial environment we try to perturb reward to interfere in learning process of the RL agent. We considered this attack also in scenarios with a single reward which we will show the process later.

\section{Adversarial attacks on different methodologies}
Here we review adversarial attacks on a variety of value function based state-of-art algorithms proposed by the community so far. As \cite{Pattanaik2018} pointed out, the concept of achieving a successful adversarial attack is unalike in reinforcement learning tasks compared to image classification tasks. Worst \textit{action} is tangible and feasible but not worst \textit{image}.

\subsection{Deep Q Learning (DQN)}
Utilizing experience replay and the leveraging target network, DQN take advantage of them to gain more stability \cite{Pattanaik2018}.

\section{Experiment}
We apply the approach to two MDPs proposed by \cite{Strehl2004AnProcesses}. \textit{RiverSwim} an MDP consists of six states and \textit{SixArms} one with seven states. In \textit{RiverSwim} flow of the river is to left and the swimmer picks probabilistically equal either state 1 or 2 as the start point. All states but two terminal states have zero reward to land at while the right most one rewards substantially higher than the other one. The agent in SixArms selects between six distinct actions which pulls different arms of a multi-armed bandit. By pulling an arm the agent traverses to a new state. Although the agent does not obtain reward by pulling the arms but going to the connected state to that arm is highly rewarding in which the transition probability is in inverse relationship with the reward value.

As both problems favor keeping the agent in states with lower reward, exploration is paramount to maximize the accumulated reward. This is also confirmed by the authors in \cite{Strehl2004AnProcesses} that the agent will fail to learn when a rudimentary $\epsilon$-greedy algorithm.

Our implementation uses Deep Q-Network (DQN) \cite{Mnih2015} as the policy optimizer. OpenAI Gym \cite{BrockmanOpenAIGym} is the testbed in the benchmarks in this paper. In training phase, models are trained in the environment. Online methods such as DQN learn by interacting directly in the environment, however, CRAAM

% \begin{figure}
%     \centering
%     \includegraphics[width=1\columnwidth]{minutia/cartpole-v1_batchsize-32_episodes-100_full-step.png}
%     \caption{Average reward of DQN training to 100 episodes in cartpole.}
%     \label{fig:my_label}
% \end{figure}

\section{Future Works}
We developed the presented approach with having reproducibility in mind. It could be served as a launching pad for the future contributors who want to surf this newly emerged field deeper. Future ideas to pursue is a generative model to be able to generate adversaries from a network.

\section*{Appendix}

\subsection{Implementation note}
* Reshape is needed after getting each state from Gym models.

\bibliographystyle{plain}
\bibliography{library}
% \section*{References}

% [1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms for
% connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and T.K.\ Leen
% (eds.), {\it Advances in Neural Information Processing Systems 7},
% pp.\ 609--616. Cambridge, MA: MIT Press.

% [2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS: Exploring
%   Realistic Neural Models with the GEneral NEural SImulation System.}  New York:
% TELOS/Springer--Verlag.

% [3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of learning and
% recall at excitatory recurrent synapses and cholinergic modulation in rat
% hippocampal region CA3. {\it Journal of Neuroscience} {\bf 15}(7):5249-5262.

\end{document}

### 2019-10-24 texlean.sh
Shell script for trashing Latex generated files

@@ -1,0 +1,4 @@
+# trashes the Latex generated peripheral files
+# Shayan Amani - Oct 2019
+
+find . -type f -name '*.aux' -o -name '*.gz' -o -name '*.log' -o -name '*.out' | xargs -L1 trash

### 2019-10-24 texlean.sh
Latex trash Shell updated

@@ -1,4 +1,5 @@
 # trashes the Latex generated peripheral files
 # Shayan Amani - Oct 2019
 
+trash ./out
 find . -type f -name '*.aux' -o -name '*.gz' -o -name '*.log' -o -name '*.out' | xargs -L1 trash

### 2019-10-29 minutia/paper/paper.tex
paper LSPI added

@@
 
- \usepackage[sort]{natbib}
-
-\title{Feature Extraction and Aggregation}
+\title{Adversarial State Aggregation}
 
@@
 \author{%
-  Shayan Amani, Marek Petrik \\ %\thanks{Use footnote for providing further information
+Shayan Amani \\ %\thanks{Use footnote for providing further information
 %    about author (webpage, alternative address)---\emph{not} for acknowledging
@@
     \begin{abstract}
-  Promising advances in deep reinforcement learning suggests several application to the algorithms.
-However robustness of these algorithms remains doubts to trust these algorithms in practice.
-In this paper we study resistance of the state-of-the-art algorithms to environment dynamics.
-We then show the amount of robustness we can achieve as a guarantee of performance in presence of those changes.
+        Promising advances in deep reinforcement learning suggests several application to the algorithms. However
+        robustness of these algorithms remains doubts to trust these algorithms in practice. In this paper we study
+        resistance of the state-of-the-art algorithms to environment dynamics. We then show the amount of robustness
+        we can achieve as a guarantee of performance in presence of those changes.
     \end{abstract}
@@
     \section{Introduction}
+    Although state aggregation may not generalize as well as methods based on neural networks, the techniques has
+    both analytical and practical advantages. In camparison, state aggregation takes less computation power to
+    compute representative features. Moreover, analysis and troubleshooting in a state aggregation process is more
+    obvious than a black-box neural network. Approximation is a corner stone in all successfull RL methods to hold a
+    "good" mapping between a large state space to a much smaller one where in value function is computable or at
+    lesast feasible. These "good" approximators end up with "good" features to represent the projected state space.
 
-Reassuring achievements in the Deep RL in a variety of fields absorb the attention from the community \citep{Mnih2015, Schulman2015TrustOptimization, Lillicrap2015ContinuousDDPG, silver2016a}.
-The cross fertilization in using approximation techniques by Deep Neural Networks in solving immense state space problems in RL is a key.
-While the newly emerged methods are growing at a staggering pace, a parallel area of research has been formed wherein a variety of techniques have been utilized to make these methods contravene their promised functionality.
-These efforts would be served to develop robustness in those spectacular models and reduce their susceptibility to data tampering or dynamics of the environment.
-Robustness ---as a property in machine learning--- elevates assurance about methods performance and their reliability to the foreground.
-In a few studies \cite{Goodfellow2014}  vulnerabilities of DNN's to adversarially-manipulated inputs are revealed.
-Therefore, robust algorithms, have to be of paramount importance in applying deep RL to real world scenarios to mitigate risks of implementation.
-Adversarial attacks as a way to perturbing these algorithms, are introduced in recent works \cite{Huang2017, Kos2017} from different aspects to degrade algorithms' performance.
-One of such potentials are changes in the environment dynamics.
+    Reassuring achievements in the Deep RL in a variety of fields absorb the attention from the community
+    \citep{Mnih2015, Schulman2015TrustOptimization, Lillicrap2015ContinuousDDPG, silver2016a}. The cross
+    fertilization in using approximation techniques by Deep Neural Networks in solving immense state space problems
+    in RL is a key. While the newly emerged methods are growing at a staggering pace, a parallel area of research
+    has been formed wherein a variety of techniques have been utilized to make these methods contravene their
+    promised functionality. These efforts would be served to develop robustness in those spectacular models and
+    reduce their susceptibility to data tampering or dynamics of the environment. Robustness ---as a property in
+    machine learning--- elevates assurance about methods performance and their reliability to the foreground. In a
+    few studies \cite{Goodfellow2014}  vulnerabilities of DNN's to adversarially-manipulated inputs are revealed.
+    Therefore, robust algorithms, have to be of paramount importance in applying deep RL to real world scenarios to
+    mitigate risks of implementation. Adversarial attacks as a way to perturbing these algorithms, are introduced in
+    recent works \cite{Huang2017, Kos2017} from different aspects to degrade algorithms' performance. One of such
+    potentials are changes in the environment dynamics.
 
-In this work, we scrutinize state-of-the-art deep RL algorithms resiliency to environment parameters alterations and then introduced our approach to overcome those.
-Then a guarantee of performance to these attacks is introduced.
+    In this work, we scrutinize state-of-the-art deep RL algorithms resiliency to environment parameters alterations
+    and then introduced our approach to overcome those. Then a guarantee of performance to these attacks is introduced.
 
     \section{Related Work}
-\cite{Goodfellow2014} stated and consequently depicted that the root cause of vulnerability to adversarial examples laid behind the piecewise linearity attributes of neural network models.
+    \cite{Goodfellow2014} stated and consequently depicted that the root cause of vulnerability to adversarial
+    examples laid behind the piecewise linearity attributes of neural network models.
 
-\section{Definition}
-For the sake of clarity, we elaborate on some preliminary definitions here to have a consistent set of notations to the end of this paper.
+    \section{Method}
+    For the sake of clarity, we elaborate on some preliminary definitions here to have a consistent set of notations
+    to the end of this paper. We use model-free LSPI and LSTDQ \citep{Lagoudakis2003} to calculate and
+    evaluate the policy of actions, respectively.
 
     \subsection{Adversarial attacks}
-The designed attack may have different levels of access to the underlying training policy which have motivated the community to classify adversarial attacks into two distinct categories, namely \textit{white-box} or \textit{black-box} attacks.
-White-box attacks illustrating the scenario in which the...
-Adversarial attacks fall in either one of the following categories based upon the stage which in the adversary perturbs the model:
+    The designed attack may have different levels of access to the underlying training policy which have motivated
+    the community to classify adversarial attacks into two distinct categories, namely \textit{white-box} or
+    \textit{black-box} attacks. White-box attacks illustrating the scenario in which the... Adversarial attacks fall
+    in either one of the following categories based upon the stage which in the adversary perturbs the model:
     \begin{itemize}

### 2019-11-01 codebase/adrel/craam/riverswim.py
LSPI RiverSwim coded

@@ -1,4 +1,4 @@
-# Riverswim
+# RiverSwim
 
 import pandas as pd
 import numpy as np

### 2019-11-01 codebase/lspi/basis_functions.py

@@ -2,6 +2,7 @@
 """Abstract Base Class for Basis Function and some common implementations."""
 
 import abc
+from functools import reduce
 
 import numpy as np
 
@@ -290,7 +291,7 @@ class OneDimensionalPolynomialBasis(BasisFunction):
 
         phi = np.zeros((self.size(), ))
 
-        offset = (self.size()/self.num_actions)*action
+        offset = int((self.size()/self.num_actions)*action)
 
         value = state[0]
 

### 2019-11-01 codebase/lspi/domains.py

@@ -323,7 +323,7 @@ class RiverSwimDomain(Domain):
         return self._state
 
     def apply_action(self, action):
-        """Apply the action to the MDP.
+        """finds the current and state and action in MDP and returns the corresponding sample.
 
            Parameters
            ----------
@@ -341,17 +341,24 @@ class RiverSwimDomain(Domain):
                If the action index is outside of the range [0, num_actions)
 
            """
-        if action < 0 or action > 1:
-            raise ValueError('Action index outside of bounds [0, %d)' %
-                             self.num_actions)
+        if action < 0 or action > self.num_actions - 1:
+            raise ValueError('Action index outside of bounds [0, %d)' % self.num_actions)
 
-        sample = self.mdp[(self.mdp['idstatefrom'] == self._state[0]) & (self.mdp['idaction'] == action)]
+        curr_sa = (self.mdp['idstatefrom'] == self._state.item(0)) & (self.mdp['idaction'] == action)
+        curr_candidates = self.mdp[curr_sa]
 
-        rnd = random()
-        if len(sample) > 1:
-            s = (sample['probability'] - rnd).apply(abs).argmin
-            samp_final = sample.loc[s]
-        return Sample(self._state.copy(), action, np.array(samp_final.idstateto[0]), np.array(samp_final.reward[0]))
+        # select the sample with the closest probability
+        rand = np.random.random()
+        if len(curr_candidates) > 1:
+            curr_sample = curr_candidates.loc[(np.abs(curr_candidates['probability'] - rand)).argmin]
+        else:
+            curr_sample = curr_candidates.copy(deep=True)
+
+        sample = Sample(self._state.copy(), action, np.array([float(curr_sample.reward)]), np.array([float(curr_sample.idstateto)]))
+
+        self._state = np.array([float(curr_sample.idstateto)])
+
+        return sample
 
     def reset(self, initial_state=None):
         if initial_state is None:
@@ -367,4 +374,11 @@ class RiverSwimDomain(Domain):
             self._state = state
 
     def __init_random_state(self):
+        """
+        RiverSwim starts at either state 1 or 2.
+
+        :returns: The starting state
+
+        :raises: The 
+        """
         return np.array([np.random.randint(1, 2 + 1)])

### 2019-11-01 codebase/lspi/main.py

@@ -2,7 +2,7 @@ import numpy as np
 import pandas as pd
 
 from domains import ChainWalkDomain, RiverSwimDomain
-from basis_functions import FakeBasis
+from basis_functions import FakeBasis, OneDimensionalPolynomialBasis
 from policy import Policy
 from solvers import LSTDQSolver
 from lspi import learn
@@ -23,8 +23,10 @@ for i in range(1, 101):
     a = samples[-1].action
     samples.append(domain.apply_action(a))
 
-# basis = OneDimensionalPolynomialBasis(3, 2)
 basis = FakeBasis(2)
+poly_basis = OneDimensionalPolynomialBasis(3, 2)
+poly_basis.evaluate(np.array([2]), 1)
+
 policy = Policy(basis)
 print('initial policy weights:', policy.weights)
 
@@ -32,14 +34,13 @@ solver = LSTDQSolver()
 
 ret = learn(samples, policy, solver, max_iterations=100)
 print('final policy weights:', ret.weights)
+policy.weights
 
 if False: del policy, ret
 
 
-##### Riverswim
+##### RiverSwim
 rs_mdp = pd.read_csv("codebase/adrel/craam/riverswim_mdp.csv")
-# sample = rs_mdp[(rs_mdp['idstatefrom'] == 0) & (rs_mdp['idaction'] == 0)]
-# sample.idaction + 4 == 4
 
 samples = []
 

### 2019-11-01 minutia/paper/paper.tex
paper updated

@@
 \author{%
-Shayan Amani \\ %\thanks{Use footnote for providing further information
+Authors \\ %\thanks{Use footnote for providing further information
 %    about author (webpage, alternative address)---\emph{not} for acknowledging
@@
 University of New Hampshire \\
-\texttt{\{shayan\}\{mpetrik\}@cs.unh.edu} \\
+\texttt{\{authors\}@cs.unh.edu} \\
 % examples of more authors
@@
     For the sake of clarity, we elaborate on some preliminary definitions here to have a consistent set of notations
-    to the end of this paper. We use model-free LSPI and LSTDQ \citep{Lagoudakis2003} to calculate and
-    evaluate the policy of actions, respectively.
+    to the end of this paper. We use model-free LSPI and LSTDQ \citep{Lagoudakis2003} to update and
+    evaluate the policy of actions, respectively. CRAAM, on the other hand, as a exact method tries to
 

### 2019-11-05 codebase/lspi/domains.py
sixarms added to lspi

@@ -338,7 +338,6 @@ class RiverSwimDomain(Domain):
             print("state: ", self._state, "action:", action)
             raise KeyError('Couldn\'t find the corresponding sample!')
 
-
         # select the sample with the closest probability
         rand = np.random.random()
         if len(curr_candidates) == 1:
@@ -395,4 +394,107 @@ class RiverSwimDomain(Domain):
 
         :raises: The 
         """
-        return np.array([np.random.randint(1, self.num_states + 1)])
+        return np.array([np.random.randint(1, 2 + 1)])
+
+
+class SixArmsDomain(Domain):
+    __num_states = 7
+    __num_actions = 6
+    __action_names = ['zero', 'one', 'two', 'three', 'four', 'five']
+
+    def __init__(self,
+                 mdp,
+                 initial_state=None):
+
+        self.num_states = SixArmsDomain.__num_states
+        self.num_actions = SixArmsDomain.__num_actions
+
+        self.mdp = mdp
+        self.reset(initial_state)
+
+    def current_state(self):
+        return self._state
+
+    def apply_action(self, action):
+        """finds the current and state and action in MDP and returns the corresponding sample.
+
+           Parameters
+           ----------
+           action: int
+               Action index. Must be in range [0, num_actions)
+
+           Returns
+           -------
+           sample.Sample
+               The sample for the applied action.
+
+           Raises
+           ------
+           ValueError
+               If the action index is outside of the range [0, num_actions)
+
+           """
+        if action < 0 or action > self.num_actions - 1:
+            raise IndexError('Action index outside of bounds [0, %d)' % self.num_actions)
+
+        curr_sa = (self.mdp['idstatefrom'] == self._state.item(0)) & (self.mdp['idaction'] == action)
+        curr_candidates = self.mdp[curr_sa]
+
+        if len(curr_candidates) == 0:
+            print("state: ", self._state, "action:", action)
+            raise KeyError('Couldn\'t find the corresponding sample!')
+
+        # select the sample with the closest probability
+        rand = np.random.random()
+        if len(curr_candidates) == 1:
+            curr_sample = curr_candidates.iloc[0]  # get a Series for type consistency
+        else:
+            curr_sample = curr_candidates.loc[(np.abs(curr_candidates['probability'] - rand)).argmin]
+
+        next_state = np.array([curr_sample.values[2].astype(int)])
+        reward = np.array([curr_sample.values[4].astype(int)])
[…]

### 2019-11-05 minutia/paper/paper.tex
paper updated

@@
     \subsection{Objective}
+    The aggregate problem is stochastic even if the original problem is deterministic. Once we lay out the aggregate
+    problem framework we can solve the problem by any exact methods either value- or policy-space solutions.
+
     The ultimate goal is deceiving the model either in choosing the worst possible action in an RL task or misclassifying an input in a classification task in a way which is not perceptible for unaided eye.

### 2019-11-05 minutia/paper/paper.tex
itermediate update

--- 2019-11-01 minutia/paper/paper.tex
@@
     to the end of this paper. We use model-free LSPI and LSTDQ \citep{Lagoudakis2003} to update and
-    evaluate the policy of actions, respectively. CRAAM, on the other hand, as a exact method tries to
+    evaluate the policy of actions, respectively. CRAAM, on the other hand, as a exact method tries to approximate
+    the value function by dynamic programming optimization.
+
+    We developed an aggregation scheme to lower the state space dimensionality by assigning same action to the state
+    in a neighbourhood, so called nearest neighbour/ piecewise linear aggregation.
 

### 2019-11-05 codebase/lspi/domains.py

@@ -1,10 +1,9 @@
 # -*- coding: utf-8 -*-
 """Contains example domains that LSPI works on."""
 import abc
-from random import randint, random
+from random import random
 
 import numpy as np
-import pandas as pd
 
 from sample import Sample
 
@@ -197,7 +196,7 @@ class ChainWalkDomain(Domain):
                              self.num_actions)
 
         action_failed = False
-        if random() < self.failure_probability:
+        if np.random.random() < self.failure_probability:
             action_failed = True
 
         # this assumes that the state has one and only one occupied location
@@ -288,7 +287,7 @@ class ChainWalkDomain(Domain):
     @staticmethod
     def __init_random_state(num_states):
         """Return randomly initialized state of the specified size."""
-        return np.array([randint(0, num_states - 1)])
+        return np.array([np.random.randint(0, num_states)])
 
 
 class RiverSwimDomain(Domain):
@@ -395,4 +394,4 @@ class RiverSwimDomain(Domain):
 
         :raises: The 
         """
-        return np.array([np.random.randint(1, self.num_states + 1)])
+        return np.array([np.random.randint(1, 2 + 1)])

### 2019-11-05 codebase/adrel/util.py
debug util added, action evaluation fixed

@@ -43,3 +43,7 @@ def pickle_load(file_name):
     with open(file_path, 'rb') as in_handle:
         return pickle.load(in_handle)
 
+
+def erwin_log(report):
+    print('ErwinLog: An exception has been caught:\n', report)
+

### 2019-11-05 codebase/lspi/basis_functions.py

@@ -6,6 +6,8 @@ from functools import reduce
 
 import numpy as np
 
+from adrel.util import erwin_log
+
 
 class BasisFunction(object):
 
@@ -284,6 +286,7 @@ class OneDimensionalPolynomialBasis(BasisFunction):
 
         """
         if action < 0 or action >= self.num_actions:
+            erwin_log('action: ' + str(action) + ', state: ' + str(state))
             raise IndexError('Action index out of bounds')
 
         if state.shape != (1, ):

### 2019-11-05 codebase/lspi/main.py

@@ -22,8 +22,6 @@ def chain_walk(n_samples):
 
     # basis = FakeBasis(2)
     poly_basis = OneDimensionalPolynomialBasis(3, 2)
-    poly_basis.evaluate(np.array([2]),1)
-    # poly_basis.evaluate(np.array([2]), 1)
     # policy = Policy(basis)
     policy = Policy(poly_basis)
     policy.weights
@@ -49,7 +47,7 @@ def mdps(domain, n_samples):
         samples.append(domain.apply_action(a))
 
     # basis = FakeBasis(2)
-    basis = OneDimensionalPolynomialBasis(3, 2)
+    basis = OneDimensionalPolynomialBasis(3, domain.num_actions)
     policy = Policy(basis)
     print('initial policy weights:', policy.weights)
 

### 2019-11-06 minutia/paper/paper.tex
paper troubleshooting and references

@@
 \usepackage{microtype}      % microtypography
+
+% SHi-ON packages
+% argmin argmax begin
+\usepackage{amsmath}
+\DeclareMathOperator*{\argmax}{arg\,max}
+\DeclareMathOperator*{\argmin}{arg\,min}
+% argmin argmax end
+\newcommand{\norm}[1]{\left\lVert#1\right\rVert}        % norm using amsmath
 
@@
     Reassuring achievements in the Deep RL in a variety of fields absorb the attention from the community
-    \citep{Mnih2015, Schulman2015TrustOptimization, Lillicrap2015ContinuousDDPG, silver2016a}. The cross
+    \citep{Mnih2015, schulman2015a, Lillicrap2015, silver2016a}. The cross
     fertilization in using approximation techniques by Deep Neural Networks in solving immense state space problems
@@
     The main goal of adversarial training and examples can be expressed briefly in form of the following equation:
-    \begin{equation}
-        \begin{aligned}
-            & \underset{\theta}{\min}
-            & & \sum_{x,y \in S} \underset{\delta \in \Delta}{\max}
-            & & & \textbb{Loss}(x + \delta, y; \theta)
-        \end{aligned}
-    \end{equation}
+    %    \begin{equation}
+    %        \begin{aligned}
+    %            & \underset{\theta}{\min}
+    %            & & \sum_{x,y \in S} \underset{\delta \in \Delta}{\max}
+    %            & & & \textbb{Loss}(x + \delta, y; \theta)
+    %        \end{aligned}
+    %    \end{equation}
 
@@
             & \underset{s}{\argmin}
-            & & {\norm{s_\eta - s_c}} = \norm{s . (1 + \eps) - s_c}
+            & & {\norm{s_\eta - s_c}} = \norm{s . (1 + \epsilon) - s_c}
         \end{aligned}
@@
 
-    \section{Adversarial attacks on different methodologies}
-    Here we review adversarial attacks on a variety of value function based state-of-art algorithms proposed by the community so far.
-As \cite{Pattanaik2018} pointed out, the concept of achieving a successful adversarial attack is unalike in reinforcement learning tasks compared to image classification tasks.
-Worst \textit{action} is tangible and feasible but not worst \textit{image}.
-
-    \subsection{Deep Q Learning (DQN)}
-    Utilizing experience replay and the leveraging target network, DQN take advantage of them to gain more stability \cite{Pattanaik2018}.
-
     \section{Experiment}
-    We apply the approach to two MDPs proposed by \cite{Strehl2004AnProcesses}.
+    We apply the approach to two MDPs proposed by\cite{Strehl2004}.
 \textit{RiverSwim} an MDP consists of six states and \textit{SixArms} one with seven states.
@@
     As both problems favor keeping the agent in states with lower reward, exploration is paramount to maximize the accumulated reward.
-This is also confirmed by the authors in \cite{Strehl2004AnProcesses} that the agent will fail to learn when a rudimentary $\epsilon$-greedy algorithm.
+This is also confirmed by the authors in \cite{Strehl2004} that the agent will fail to learn when a rudimentary $\epsilon$-greedy algorithm.
 
-    Our implementation uses Deep Q-Network (DQN) \cite{Mnih2015} as the policy optimizer.
-OpenAI Gym \cite{BrockmanOpenAIGym} is the testbed in the benchmarks in this paper.
+    Our implementation uses Deep Q-Network (DQN)\cite{Mnih2015} as the policy optimizer.
+OpenAI Gym \cite{Brockman2016} is the testbed in the benchmarks in this paper.
 In training phase, models are trained in the environment.

### 2019-11-22 codebase/adrel/craam/riverswim_mdp_gg.R
riverswim r code setwd

@@ -2,6 +2,8 @@ library(rcraam)
 library(dplyr)
 library(readr)
 
+setwd('/Volumes/DevCamp/PyCharmProjects/Erwin/codebase/adrel/craam/')
+
 mdp <- read_csv("riverswim_mdp.csv")
 
 sol <- solve_mdp(mdp, 0.99)

### 2019-11-22 minutia/paper/paper.tex

@@
     \section{Introduction}
-    Although state aggregation may not generalize as well as methods based on neural networks, the techniques has
-    both analytical and practical advantages. In camparison, state aggregation takes less computation power to
-    compute representative features. Moreover, analysis and troubleshooting in a state aggregation process is more
-    obvious than a black-box neural network. Approximation is a corner stone in all successfull RL methods to hold a
-    "good" mapping between a large state space to a much smaller one where in value function is computable or at
-    lesast feasible. These "good" approximators end up with "good" features to represent the projected state space.
+    Real-world problems embody high-dimensional state spaces where in computational issues
+    arise.
+    Various approximation methods are proposed in the literature to reduce the resulted value function
+    dimensionality \cite{Sutton1998, Francois-Lavet2018}.
+    State aggregation is one of the simplest and most analytically transparent state representation approximation.
 
-    Reassuring achievements in the Deep RL in a variety of fields absorb the attention from the community
-    \citep{Mnih2015, schulman2015a, Lillicrap2015, silver2016a}. The cross
-    fertilization in using approximation techniques by Deep Neural Networks in solving immense state space problems
-    in RL is a key. While the newly emerged methods are growing at a staggering pace, a parallel area of research
-    has been formed wherein a variety of techniques have been utilized to make these methods contravene their
-    promised functionality. These efforts would be served to develop robustness in those spectacular models and
-    reduce their susceptibility to data tampering or dynamics of the environment. Robustness ---as a property in
-    machine learning--- elevates assurance about methods performance and their reliability to the foreground. In a
-    few studies \cite{Goodfellow2014}  vulnerabilities of DNN's to adversarially-manipulated inputs are revealed.
-    Therefore, robust algorithms, have to be of paramount importance in applying deep RL to real world scenarios to
-    mitigate risks of implementation. Adversarial attacks as a way to perturbing these algorithms, are introduced in
-    recent works \cite{Huang2017, Kos2017} from different aspects to degrade algorithms' performance. One of such
-    potentials are changes in the environment dynamics.
+    Although state aggregation may not generalize as well as neural network-based methods, the techniques has both
+    analytical and practical advantages \cite{Lagoudakis2003}.
+    In comparison, state aggregation takes less computation power to calculate representative features.
+    Moreover, analysis and troubleshooting in a state aggregation process is more obvious than a black-box neural
+    network.
+    Approximation is a corner stone in all successful RL methods to hold a "good" mapping between a large state
+    space to a much smaller one where in value function is computable or at least feasible.
+    These "good" approximators can end up with "good" features to represent the projected state space.
+
+    The cross-fertilization in using state aggregation and deep neural networks in solving immense state
+    space problems in RL is also profound.
+    While the newly emerged methods are growing at a staggering pace, a parallel area of research has been formed
+    wherein a variety of techniques have been utilized to make these methods contravene their promised functionality.
+    These efforts would be served to develop robustness in those spectacular models and reduce their
+    susceptibility to data tampering or dynamics of the environment.
 
@@
     and then introduced our approach to overcome those. Then a guarantee of performance to these attacks is introduced.
-
-    \section{Related Work}
-    \cite{Goodfellow2014} stated and consequently depicted that the root cause of vulnerability to adversarial
-    examples laid behind the piecewise linearity attributes of neural network models.
 
@@
     \textit{black-box} attacks. White-box attacks illustrating the scenario in which the... Adversarial attacks fall
-    in either one of the following categories based upon the stage which in the adversary perturbs the model:
+    in either one of the following categories based upon the stage, which in the adversary perturbs the model:
     \begin{itemize}

### 2019-12-26 minutia/paper/paper.tex
paper added MDP section

@@
     and then introduced our approach to overcome those. Then a guarantee of performance to these attacks is introduced.
+
+    \section{Markov Decision Processes}
+    Reinforcement learning problems can be formulated as an Markov decision process (MDP)\cite{Puterman1994}.
+    Discrete discounted reward criterion is the center of focus for ease of exposition in this paper.
+    For an environment and a decision maker, state space $S$ and action space $A$ are defined, respectively.
+    The transition probabilities $T$ describe how likely is the next state $s'$ by taking a specific action $a$ by
+    the agent in a certain state $s$ which will receive $r$ as reward.
+    For a taken action $a_i$ in the state $s_i$, the decision maker will be ended up in state $s_{i+1}$ by the
+    transition probabilities $P : S \times A \times S \rightarrow [0, 1]$ and wil receive reward $r_{i+1}$ from the
+    rewards $R : S \times A \rightarrow \mathbb{R}$: \\
+    \center $s_{i+1} = T(s_{i+1}|s_i, a_i)$ \\
+    \center $r_{i+1} = R(s_i, a_i)$ \\
 

### 2019-12-28 minutia/paper/paper.tex
paper: state aggregation section added

@@
     rewards $R : S \times A \rightarrow \mathbb{R}$: \\
-    \center $s_{i+1} = T(s_{i+1}|s_i, a_i)$ \\
-    \center $r_{i+1} = R(s_i, a_i)$ \\
+    \begin{equation}
+        s_{i+1} = T(s_{i+1}|s_i, a_i)
+    \end{equation}
+
+    \begin{equation}
+        r_{i+1} = R(s_i, a_i)
+    \end{equation}
+
+    \section{State Aggregation}
+    State aggregation is an approximation approach to make a problem more tractable by a reduction in dimensionality in
+    state space \footnote{aggregation in action space although is possible, has less significance comparatively.}.
+    State aggregation, in particular, is a parametric feature-based approximation where the features are membership
+    functions of 0-1 form.
+    Aggregation groups similar states together, so called representative states, which lets the \textit{aggregate}
+    problem to be
+    solved by an exact
+    method.
+    The computed value function from the exact solution of the aggregate problem can be then used in the original
+    problem.
+
+    Now we formulate the aggregated problem by introducing the dynamics of the aggregate problem.
+    The transition probabilities between two representative states:
+
+    \begin{equation}
+        \hat{p}_{x y}(u)=\sum_{j=1}^{n} p_{x j}(u) \phi_{j y},
+    \end{equation}
+
+    where $p_ {x j}$ is the transition probabilities from a representative state, $x$, to an original state, $j$
+    and $\phi_{j y}$ is the aggregation probabilities from an original state, $j$, to a representative state, $y$.
+    The rewards are defined as:
+
+    \begin{equation}
+        \hat{g}(x, u)=\sum_{j=1}^{n} p_{x j}(u) g(x, u, j)
+    \end{equation}
 

### 2019-12-30 minutia/paper/paper.tex
updated paper

@@
     Aggregation groups similar states together, so called representative states, which lets the \textit{aggregate}
-    problem to be
-    solved by an exact
-    method.
+    problem to be solved by an exact method.
     The computed value function from the exact solution of the aggregate problem can be then used in the original
@@
     \begin{equation}
-        \hat{p}_{x y}(u)=\sum_{j=1}^{n} p_{x j}(u) \phi_{j y},
+        \hat{P}_{s s^\prime}(a)=\sum_{j=1}^{n} P_{s j}(a) \phi_{j y},
     \end{equation}
 
-    where $p_ {x j}$ is the transition probabilities from a representative state, $x$, to an original state, $j$
-    and $\phi_{j y}$ is the aggregation probabilities from an original state, $j$, to a representative state, $y$.
+    where $p_ {s j}$ is the transition probabilities from a representative state, $s$, to an original state, $j$
+    and $\phi_{j s^\prime}$ is the aggregation probabilities from an original state, $j$, to a representative state,
+    $s^\prime$.
     The rewards are defined as:
@@
     \begin{equation}
-        \hat{g}(x, u)=\sum_{j=1}^{n} p_{x j}(u) g(x, u, j)
+        \hat{r}(s, a)=\sum_{j=1}^{n} P_{s j}(a) r(s, a, j),
     \end{equation}
+
+    where $\hat{r}(s, a)$ is the received reward at state $s$ by taking action $a$.
+
+    Aggregation then helps us to derive the value function.
+    Approximated optimal value function is a weighted sum of optimal rewards and can be computed by
+    equation\ref{approx val func} where $r_{s^\prime}^{\ast}$ is optimal reward at represented state $s$.
+
+    \begin{equation} \label{approx val func}
+        \tilde{V}(j)=\sum_{s^\prime \in \mathcal{U}} \phi_{j s^\prime} r_{s^\prime}^{\ast},
+    \end{equation}
+
+    In this paper, we lay down the aggregation framework and then solve the aggregate problem by a tabular method
+    such as value iteration (VI), policy iteration (PI).
 
@@
     in a neighbourhood, so called nearest neighbour/ piecewise linear aggregation.
+    We then solve directly the lower dimension problem which relatively takes less computation.
 

### 2019-12-30 codebase/adrel/craam/portfolio.py
MDPs abtract base class fixed

@@ -1,106 +1,0 @@
-#!/usr/bin/python
-
-# Copyright 2019, Gurobi Optimization, LLC
-
-# Portfolio selection: given a sum of money to invest, one must decide how to
-# spend it amongst a portfolio of financial securities.  Our approach is due
-# to Markowitz (1959) and looks to minimize the risk associated with the
-# investment while realizing a target expected return.  By varying the target,
-# one can compute an 'efficient frontier', which defines the optimal portfolio
-# for a given expected return.
-#
-# Note that this example reads historical return data from a comma-separated
-# file (../data/portfolio.csv).  As a result, it must be run from the Gurobi
-# examples/python directory.
-#
-# This example requires the pandas (>= 0.20.3), NumPy, and Matplotlib
-# Python packages, which are part of the SciPy ecosystem for
-# mathematics, science, and engineering (http://scipy.org).  These
-# packages aren't included in all Python distributions, but are
-# included by default with Anaconda Python.
-
-from gurobipy import *
-from math import sqrt
-import pandas as pd
-import numpy as np
-import matplotlib.pyplot as plt
-
-# Import (normalized) historical return data using pandas
-data = pd.read_csv('codebase/dataset/portfolio.csv', index_col=0)
-stocks = data.columns
-
-# Calculate basic summary statistics for individual stocks
-stock_volatility = data.std()
-stock_return = data.mean()
-
-# Create an empty model
-m = Model('portfolio')
-
-# Add a variable for each stock
-vars = pd.Series(m.addVars(stocks), index=stocks)
-
-# Objective is to minimize risk (squared).  This is modeled using the
-# covariance matrix, which measures the historical correlation between stocks.
-sigma = data.cov()
-portfolio_risk = sigma.dot(vars).dot(vars)
-m.setObjective(portfolio_risk, GRB.MINIMIZE)
-
-# Fix budget with a constraint
-m.addConstr(vars.sum() == 1, 'budget')
-
-# SHi-ON
-m.getVars()
-m.getConstrs()
-
-# Optimize model to find the minimum risk portfolio
-m.setParam('OutputFlag', 0)
-m.optimize()
-
-# Create an expression representing the expected return for the portfolio
-portfolio_return = stock_return.dot(vars)
-
-# Display minimum risk portfolio
-print('Minimum Risk Portfolio:\n')
[…]

### 2019-12-30 codebase/rlessential/mdps.py

@@ -45,7 +45,7 @@ class MDP(object):
         pass  # pragma: no cover
 
 
-class TwoStateMDP:
+class TwoStateMDP(MDP):
 
     def __init__(self, mdp):
         self.mdp = mdp
@@ -81,7 +81,7 @@ class TwoStateMDP:
         return probs
 
 
-class TwoStateParametricMDP:
+class TwoStateParametricMDP(MDP):
 
     def __init__(self, mdp, param):
         self.mdp = mdp

### 2019-12-30 codebase/rlessential/mdps.py
MDPs abtract base class fixed

@@ -4,6 +4,7 @@ import numpy as np
 
 
 class MDP(object):
+    __metaclass__ = abc.ABCMeta
 
     @abc.abstractmethod
     def state_count(self):
@@ -90,7 +91,6 @@ class TwoStateParametricMDP(MDP):
         self.num_actions = self.action_count()
         self.rewards = TwoStateParametricMDP.parametrize_rewards(param)
         self.probabilities = TwoStateParametricMDP.parametrize_probabilities(param)
-
 
     def state_count(self):
         unique_from = self.mdp['idstatefrom'].nunique()

### 2020-01-13 minutia/paper/paper.tex
paper updated

@@
     \section{State Aggregation}
+    Value function approximation alleviates the computational hurdle of a large state-space problem.
+    We introduce an approximate policy iteration method to solve an aggregate problem.
     State aggregation is an approximation approach to make a problem more tractable by a reduction in dimensionality in
@@
 
-    \begin{equation} \label{approx val func}
+    \begin{equation}
+        \label{approx val func}
         \tilde{V}(j)=\sum_{s^\prime \in \mathcal{U}} \phi_{j s^\prime} r_{s^\prime}^{\ast},

### 2020-01-13 codebase/rlessential/main.py
RAAM agent's state aggregation finished

@@ -1,8 +1,25 @@
-from agents import QLearningCartPoleAgent, QLearningMountainCarAgent
+import numpy as np
 
-if __name__ == '__main__':
+from agents import QLearningCartPoleAgent, QLearningMountainCarAgent, CartPoleAgent
+
+
+def q_learning():
     # discounted-reward problem
-    # agent = QLearningCartPoleAgent(num_buckets=(1, 2, 6, 12), num_episodes=1000, discount=0.99)
-    agent = QLearningMountainCarAgent(num_buckets=(6, 6), num_episodes=10000, discount=0.99)
+    agent = QLearningCartPoleAgent(num_buckets=(1, 2, 6, 12), num_episodes=1000, discount=0.99)
+    # agent = QLearningMountainCarAgent(num_buckets=(6, 6), num_episodes=10000, discount=0.99)
     agent.train()
     agent.run()
+
+
+
+
+if __name__ == '__main__':
+    q_learning()
+    # agent = CartPoleAgent()
+    # agent.run()
+    # samples = agent.samples
+    # samples = np.array(samples)
+
+
+
+

### 2020-01-19 codebase/rlessential/agents.py
aggregation experiments

@@ -86,18 +86,24 @@ class QLearningAgent(Agent):
 
         print('Training finished!')
 
-    def run(self):
-        while True:
+    def run(self, render=True):
+        total_reward = 0
+        for e in tqdm.trange(self.num_episodes):
             current_state = self.env.reset()
             current_state = self.aggregate(current_state)
 
             done = False
             while not done:
-                self.env.render()
+                if render:
+                    self.env.render()
                 action = np.argmax(self.q_table[current_state])
                 observation, reward, done, _ = self.env.step(action)
+                total_reward += reward
                 next_state = self.aggregate(observation)
                 current_state = next_state
+
+        avg_reward = total_reward / self.num_episodes
+        print('Average reward: {0} in {1} episodes'.format(avg_reward, self.num_episodes))
 
 
 class CartPoleAgent(Agent):
@@ -172,6 +178,27 @@ class QLearningCartPoleAgent(QLearningAgent):
                              -np.radians(50) / 1]
 
         self.q_table = np.zeros(self.num_buckets + (self.env.action_space.n,))
+        self.batch = list()
+
+    def simulate(self):
+        for e in range(self.num_episodes):
+            current_observation = self.env.reset()
+            current_state = self.aggregate(current_observation)
+
+            done = False
+            while not done:
+                sample = list()
+                sample.append(current_observation)
+                action = self.choose_action(current_state)
+                sample.append(action)
+                next_observation, reward, done, _ = self.env.step(action)
+                sample.append(reward)
+                sample.append(next_observation)
+                next_state = self.aggregate(next_observation)
+                current_observation = next_observation
+                current_state = next_state
+                self.batch.append(tuple(sample))
+
 
 
 class QLearningMountainCarAgent(QLearningAgent):
@@ -252,6 +279,7 @@ class RAAMAgent:
             # in the same bucket will update the action and also outcomes.
             self.A[current_state] = sample[1]
             self.B[current_state] = sample[3]
+
 
 # showcase the agent performance
 if __name__ == '__main__':

### 2020-01-19 codebase/rlessential/main.py

@@ -8,13 +8,11 @@ def q_learning():
     agent = QLearningCartPoleAgent(num_buckets=(1, 2, 6, 12), num_episodes=1000, discount=0.99)
     # agent = QLearningMountainCarAgent(num_buckets=(6, 6), num_episodes=10000, discount=0.99)
     agent.train()
-    agent.run()
+    # agent.run(False)
+    agent.simulate()
+    return agent.batch
 
-
-
-
-if __name__ == '__main__':
-    # q_learning()
+def robust_learning():
     agent = CartPoleAgent()
     agent.run()
     samples = agent.samples
@@ -27,6 +25,20 @@ if __name__ == '__main__':
     len(raam.S)
     raam.A.shape
     raam.B.shape
+
+
+if __name__ == '__main__':
+    samples = q_learning()
+    sample_size = len(samples)
+    print('Sample size:', sample_size)
+    # step per episode
+    avg_spe = (sample_size / 1000)
+    print('Average step per episode:', avg_spe)
+
+
+    # robust_learning()
+
+
 
 
 

### 2020-01-19 minutia/paper/paper.tex
paper updated

@@
 
+    \subsection{Aggregation}
+    Find the best policy then find optimal number of buckets.
+    Train the agent and retrieve the best policy then by running the agent with the best policy \textit{generate}
+    samples.
+    These collected samples are not based on a random policy but the best trained policy (baseline).
+    Then we can test our hypothesis on the collected samples to see how discretization affect another trained agent.
+    One way to measure the effect of discretization is average reward received by the trained agent.
+
     \subsection{Adversarial attacks}

### 2020-02-17 minutia/paper/paper.tex
paper refinement

@@
 
-    \section{Devising Adversarial Attacks}
-    Neural networks are vulnerable to adversarial attacks either in supervised learning or reinforcement learning settings \cite{Huang2017}.
-Such weakness could possibly result in unrecognizable differences for the unaided eye in the presence of perturbation which leads to misclassification for neural network based model \cite{Szegedy2014, Goodfellow2014}.
-
-    \subsection{Observation-perturbed attacks}
-
-    where in we trying to find an adversarial example (inner maximization) in order to adversarially train a robust model (outer minimization).
-
-    We define $s_\eta$ as the perturbed observation:
-    \begin{equation}
-        s_\eta = s . (1 + \epsilon)
-    \end{equation}
-
-    Let $s_\eta$ be the perturbed observation and $s_c$ scaled observation, so we are trying to find the following:
-    \begin{equation}
-        \begin{aligned}
-            & \underset{s}{\argmin}
-            & & {\norm{s_\eta - s_c}} = \norm{s . (1 + \epsilon) - s_c}
-        \end{aligned}
-    \end{equation}
-    which is the state observation that have the minimum distance between real-world [scaled] observation and the perturbed one.
-
-    In order to expand the applicability of designed adversarial attacks to an assortment of algorithms [models], we rely on \textit{transferability property} of RL-specific adversarial attacks \cite{Szegedy2014, Papernot2016, Goodfellow2014a}.
-\cite{Huang2017} clearly assert [or asserts?!] attacks are deployable to a group of trained RL models aimed at the same task.
-
-    \subsection{Reward-Perturbed attacks}
-    In an adversarial environment we try to perturb reward to interfere in learning process of the RL agent.
-We considered this attack also in scenarios with a single reward which we will show the process later.
-
     \section{Experiment}

### 2020-02-17 codebase/rlessential/agents.py
range aggregation initialized

@@ -239,21 +239,6 @@ class QLearningMountainCarAgent(QLearningAgent):
         self.q_table = np.zeros(self.num_buckets + (self.env.action_space.n,))
 
 
-class MachineReplacementAgent(Agent):
-
-    def __init__(self, domain):
-        self.domain = domain
-
-    def choose_action(self, state):
-        return
-
-    def train(self):
-        pass
-
-    def run(self):
-        pass
-
-
 class RAAMAgent:
 
     def __init__(self, samples, env, num_buckets=(1, 40, 40, 40)):
@@ -327,6 +312,20 @@ class RAAMAgent:
                         # TODO: check to be correct
                         self.r[a,b,s] = (1 / len(transition_reward)) * sum_rewards
 
+
+class MachineReplacementAgent(Agent):
+
+    def __init__(self, domain):
+        self.domain = domain
+
+    def choose_action(self, state):
+        return
+
+    def train(self):
+        pass
+
+    def run(self):
+        pass
 
 # showcase the agent performance
 if __name__ == '__main__':

### 2020-02-17 codebase/rlessential/domains.py

@@ -153,40 +153,6 @@ class TwoStateParametricMDP(MDP):
         return [a / 2, 1 - a / 2, 1]
 
 
-class MachineReplacementMDP(MDP):
-
-    def __init__(self, mdp):
-        self.mdp = mdp
-        self.num_states = self.state_count()
-        self.num_actions = self.action_count()
-
-    def state_count(self):
-        unique_from = self.mdp['idstatefrom'].nunique()
-        unique_to = self.mdp['idstateto'].nunique()
-        return max(unique_from, unique_to)
-
-    def action_count(self):
-        unique_action = self.mdp['idaction'].nunique()
-        return unique_action
-
-    def get_rewards(self):
-        rewards = np.zeros((self.num_states * self.num_actions, 1))
-        for  index, row in self.mdp.iterrows():
-            s = row['idstatefrom'].astype(int)
-            a = row['idaction'].astype(int)
-            rewards[s * self.num_actions + a] = row['reward']
-        return rewards
-
-    def get_probabilities(self):
-        probs = np.zeros((self.num_states * self.num_actions, self.num_states))
-        for index, row in self.mdp.iterrows():
-            s = row['idstatefrom'].astype(int)
-            a = row['idaction'].astype(int)
-            sp = row['idstateto'].astype(int)
-            probs[s * self.num_actions + a][sp] = row['probability']
-        return probs
-
-
 class RAAMMDP(MDP):
     """
     Three-state deterministic MDP in the `RAAM Paper <http://www.cs.unh.edu/~mpetrik/pub/Petrik2014_appendix.pdf>`_.
@@ -234,3 +200,36 @@ class RAAMMDP(MDP):
             probs[s * self.num_actions + a][sp] = row['probability']
         return probs
 
+
+class MachineReplacementMDP(MDP):
+
+    def __init__(self, mdp):
+        self.mdp = mdp
+        self.num_states = self.state_count()
+        self.num_actions = self.action_count()
+
+    def state_count(self):
+        unique_from = self.mdp['idstatefrom'].nunique()
+        unique_to = self.mdp['idstateto'].nunique()
+        return max(unique_from, unique_to)
+
+    def action_count(self):
+        unique_action = self.mdp['idaction'].nunique()
+        return unique_action
+
+    def get_rewards(self):
+        rewards = np.zeros((self.num_states * self.num_actions, 1))
+        for  index, row in self.mdp.iterrows():
+            s = row['idstatefrom'].astype(int)
[…]

### 2020-02-17 codebase/rlessential/main.py

@@ -44,11 +44,7 @@ def robust_learning():
     raam.B.shape
 
 
-if __name__ == '__main__':
-    # q_learning()
-
-    # robust_learning()
-
+def cartpole_iqr_compare():
     print("loading samples")
 
     samples = pickle.load(open('samples.p', 'rb'))
@@ -69,7 +65,7 @@ if __name__ == '__main__':
 
     n = len(smp)
     iq_range = iqr(smp, axis=0)
-    h = 2 * iq_range * (n ** (-1/3))
+    h = 2 * iq_range * (n ** (-1 / 3))
 
     hi = np.max(smp, axis=0)
     lo = np.min(smp, axis=0)
@@ -91,5 +87,14 @@ if __name__ == '__main__':
     agent_hist.run(False)
 
 
-    # training the agent to learn the baseline Q-table
+if __name__ == '__main__':
+    # q_learning()
+
+    # robust_learning()
+
+    # cartpole_iqr_compare()
+
+    pass
+
+
 

### 2020-02-17 codebase/rlessential/solvers.py

@@ -128,5 +128,6 @@ def main():
     value_iteration(mdp, threshold=tau, discount=gamma)
     policy_iteration(mdp, threshold=tau, discount=gamma)
 
+
 if __name__ == '__main__':
     main()

### 2020-02-17 minutia/paper/paper.tex
paper updated with Freedman-Diaconis

@@
 
-\title{Adversarial State Aggregation}
+\title{Distributional State Aggregation}
 
@@
     \begin{abstract}
-        Promising advances in deep reinforcement learning suggests several application to the algorithms. However
-        robustness of these algorithms remains doubts to trust these algorithms in practice. In this paper we study
-        resistance of the state-of-the-art algorithms to environment dynamics. We then show the amount of robustness
-        we can achieve as a guarantee of performance in presence of those changes.
+        Promising advances in deep reinforcement learning suggests several application to the algorithms.
+        However robustness of these algorithms remains doubts to trust these algorithms in practice.
+        In this paper we study resistance of the state-of-the-art algorithms to environment dynamics.
+        We then show the amount of robustness we can achieve as a guarantee of performance in presence of those changes.
     \end{abstract}
@@
 
-    In this work, we scrutinize state-of-the-art deep RL algorithms resiliency to environment parameters alterations
-    and then introduced our approach to overcome those. Then a guarantee of performance to these attacks is introduced.
-
     \section{Markov Decision Processes}
@@
     State aggregation is an approximation approach to make a problem more tractable by a reduction in dimensionality in
-    state space \footnote{aggregation in action space although is possible, has less significance comparatively.}.
+    state space \footnote{aggre
+    gation in action space although is possible, has less significance comparatively.}.
     State aggregation, in particular, is a parametric feature-based approximation where the features are membership
@@
     \section{Method}
+    Splitting the observations into neighborhoods of similar observations is one way to interpret an aggregation
+    problem.
+    Inspired by histograms and the statistics behind finding the proper number of bins for a given distribution, we
+    build our proposal on top of such basics.
+    Finding proper number of bins for a given distribution could be calculated by Freedman-Diaconis rule. As the rule
+    states:
+
+    \begin{equation}
+        \text { Bin width }=2 \frac{\operatorname{IQR}(x)}{\sqrt[3]{n}}
+    \end{equation}
+
+    where n is number of observations and IQR is the interquartile range of data.
+
     For the sake of clarity, we elaborate on some preliminary definitions here to have a consistent set of notations
     to the end of this paper. We use model-free LSPI and LSTDQ \citep{Lagoudakis2003} to update and
-    evaluate the policy of actions, respectively. CRAAM, on the other hand, as a exact method tries to approximate
-    the value function by dynamic programming optimization.
+    evaluate the policy of actions, respectively.
 
@@
 
-    \subsection{Aggregation}
-    Find the best policy then find optimal number of buckets.
-    Train the agent and retrieve the best policy then by running the agent with the best policy \textit{generate}
-    samples.
-    These collected samples are not based on a random policy but the best trained policy (baseline).
-    Then we can test our hypothesis on the collected samples to see how discretization affect another trained agent.
-    One way to measure the effect of discretization is average reward received by the trained agent.
-
-    \subsection{Adversarial attacks}
-    The designed attack may have different levels of access to the underlying training policy which have motivated
-    the community to classify adversarial attacks into two distinct categories, namely \textit{white-box} or
-    \textit{black-box} attacks. White-box attacks illustrating the scenario in which the... Adversarial attacks fall
-    in either one of the following categories based upon the stage, which in the adversary perturbs the model:
-    \begin{itemize}
-        \item Train time: Data poisoning.
-        \item Evaluation time: Adversarial examples.
-        \item Deploy time: Black-box attacks.
-    \end{itemize}
-
-    \subsection{Objective}
-    The aggregate problem is stochastic even if the original problem is deterministic. Once we lay out the aggregate
-    problem framework we can solve the problem by any exact methods either value- or policy-space solutions.
-
-    The ultimate goal is deceiving the model either in choosing the worst possible action in an RL task or misclassifying an input in a classification task in a way which is not perceptible for unaided eye.
-In terms of a Markovian RL task we show how minimum set of worst-case actions are needed to result in the minimum final reward.
-We try to show that the model is not robust to other configuration of each benchmarks with different dynamics.
-
-    The main goal of adversarial training and examples can be expressed briefly in form of the following equation:
-    %    \begin{equation}
-    %        \begin{aligned}
-    %            & \underset{\theta}{\min}
-    %            & & \sum_{x,y \in S} \underset{\delta \in \Delta}{\max}
-    %            & & & \textbb{Loss}(x + \delta, y; \theta)
-    %        \end{aligned}
-    %    \end{equation}
+    The aggregate problem is stochastic even if the original problem is deterministic.
+    Once we lay out the aggregate problem framework we can solve the problem by any exact methods either value- or
+    policy-space solutions.
 
@@
 This is also confirmed by the authors in \cite{Strehl2004} that the agent will fail to learn when a rudimentary $\epsilon$-greedy algorithm.
-
-    Our implementation uses Deep Q-Network (DQN)\cite{Mnih2015} as the policy optimizer.
-OpenAI Gym \cite{Brockman2016} is the testbed in the benchmarks in this paper.
-In training phase, models are trained in the environment.
-Online methods such as DQN learn by interacting directly in the environment, however, CRAAM
-
-    % \begin{figure}
-    %     \centering
-    %     \includegraphics[width=1\columnwidth]{minutia/cartpole-v1_batchsize-32_episodes-100_full-step.png}
-    %     \caption{Average reward of DQN training to 100 episodes in cartpole.}
-    %     \label{fig:my_label}
-    % \end{figure}
 

### 2020-02-24 codebase/rlessential/agents.py
machine replacement IQR preliminary

@@ -319,7 +319,7 @@ class MachineReplacementAgent(Agent):
         self.domain = domain
 
     def choose_action(self, state):
-        return
+
 
     def train(self):
         pass

### 2020-02-24 codebase/rlessential/domains.py

@@ -2,6 +2,8 @@ from abc import ABC, abstractmethod
 
 import numpy as np
 
+# TODO: migrate self-defined domains in lspi's domain module to here
+
 
 class MDP(ABC):
     # same as new in 3.4:

### 2020-02-24 codebase/rlessential/main.py

@@ -1,4 +1,72 @@
 
+import pandas as pd
+import numpy as np
+
+from rlessential.domains import MachineReplacementMDP
+from agents import MachineReplacementAgent
+from rlessential.solvers import ValueIteration
+
+
+def cartpole_iqr_compare():
+    print("loading samples")
+
+    samples = pickle.load(open('samples.p', 'rb'))
+
+    print("unifying samples")
+
+    ss = []
+    for s in samples:
+        ss.append(s[0])
+        # ss.append(s[3])
+    len(samples)
+    len(ss)
+
+    print("stacking")
+
+    smp = np.stack(ss, axis=0)
+    smp.shape
+
+    n = len(smp)
+    iq_range = iqr(smp, axis=0)
+    h = 2 * iq_range * (n ** (-1 / 3))
+
+    hi = np.max(smp, axis=0)
+    lo = np.min(smp, axis=0)
+
+    counts = (hi - lo) / h
+    counts = counts / 1
+    counts = np.round(counts)
+    counts = counts.astype(int)
+    num_buckets = tuple(counts)
+
+    print("agents")
+
+    agent_base = QLearningCartPoleAgent(num_buckets=(1, 2, 6, 12), num_episodes=1000, discount=0.99)
+    agent_base.train()
+    agent_base.run(False)
+
+    agent_hist = QLearningCartPoleAgent(num_buckets=num_buckets, num_episodes=1000, discount=0.99)
+    agent_hist.train()
+    agent_hist.run(False)
+
 
 if __name__ == '__main__':
-    pass
+
+    mdp = pd.read_csv('dataset/mdp/machine_replacement_mdp.csv')
+
+    domain_mr = MachineReplacementMDP(mdp)
+    agent_mr = MachineReplacementAgent(domain_mr)
+
+
+    gamma = 0.90
+    epsilon = 0.0000001
+    tau = (epsilon * (1 - gamma)) / (2 * gamma)
+
+    # even different initial values will end up with the same state values!
+    init_val = np.arange(10)
+    init_val = np.random.rand(10) * 10
+
+    vi = ValueIteration(domain_mr, discount=gamma, initial_values=init_val, threshold=tau, verbose=True)
+    vi.calculate_value()
+
+

### 2020-02-24 codebase/rlessential/solvers.py

@@ -10,11 +10,13 @@ from util import timing
 
 class ValueIteration:
 
-    def __init__(self, domain, discount, threshold, initial_values=None, verbose=False):
+    def __init__(self, domain, discount, threshold,
+                 num_iterations=0, initial_values=None, verbose=False):
         self.domain = domain
         self.discount = discount
         self.threshold = threshold
         self.initial_values = initial_values
+        self.num_iterations = num_iterations
         self.verbose = verbose
 
         # format: (#iter, V(s), distance)
@@ -41,7 +43,7 @@ class ValueIteration:
         dist = np.inf
         i = 1
         t = time.perf_counter()
-        while dist >= self.threshold:
+        while dist >= self.threshold or i < self.num_iterations:
 
             # noinspection PyCompatibility
             v = r + self.discount * p @ v_curr

### 2020-02-24 codebase/rlessential/agents.py
tiny fix

@@ -319,7 +319,7 @@ class MachineReplacementAgent(Agent):
         self.domain = domain
 
     def choose_action(self, state):
-
+        pass
 
     def train(self):
         pass

### 2020-02-24 minutia/paper/paper.tex
paper updated

@@
     \begin{abstract}
-        Promising advances in deep reinforcement learning suggests several application to the algorithms.
-        However robustness of these algorithms remains doubts to trust these algorithms in practice.
-        In this paper we study resistance of the state-of-the-art algorithms to environment dynamics.
-        We then show the amount of robustness we can achieve as a guarantee of performance in presence of those changes.
+        We present a state aggregation algorithm and a learning algorithm based on the former to solve a stochastic
+        problem.
+        The presented aggregation algorithm uses range of collected samples to provide the proper aggregation
+        \textit{buckets}.
+        Our approach is a batch learning by which we detect the \textit{splitting policy} using inter-quartile range
+        of data for each feature.
+        We then show the amount of robustness we can achieve as a guarantee of performance in presence of those
+        aggregation policies.
     \end{abstract}
@@
     Various approximation methods are proposed in the literature to reduce the resulted value function
-    dimensionality \cite{Sutton1998, Francois-Lavet2018}.
+    dimensionality\cite{Sutton1998, Francois-Lavet2018}.
     State aggregation is one of the simplest and most analytically transparent state representation approximation.
@@
     Although state aggregation may not generalize as well as neural network-based methods, the techniques has both
-    analytical and practical advantages \cite{Lagoudakis2003}.
+    analytical and practical advantages\cite{Lagoudakis2003}.
     In comparison, state aggregation takes less computation power to calculate representative features.
@@
     \section{Markov Decision Processes}
-    Reinforcement learning problems can be formulated as an Markov decision process (MDP)\cite{Puterman1994}.
+    Reinforcement learning problems can be formulated using a Markov Decision Processes (MDP)
+    framework\cite{Puterman1994, Bertsekas}.
+    MDPs facilitate handling sequential decision-making under uncertainty.
     Discrete discounted reward criterion is the center of focus for ease of exposition in this paper.
@@
     Now we formulate the aggregated problem by introducing the dynamics of the aggregate problem.
-    The transition probabilities between two representative states:
+    The transition probabilities between
+    two representative states:
 

### 2020-02-24 codebase/rlessential/agents.py
MacRep intermediate

@@ -315,17 +315,43 @@ class RAAMAgent:
 
 class MachineReplacementAgent(Agent):
 
-    def __init__(self, domain):
+    def __init__(self, domain, horizon):
         self.domain = domain
+        self.horizon = horizon
 
     def choose_action(self, state):
-        pass
+        state_probs, state_actions = self.domain.get_possible_actions(state)
+        assert (abs(1 - sum(state_probs)) < 0.01)
+        action = int(np.random.choice(state_actions, p=state_probs))
+        return action
+
+    # if self.probabilities is not None:
+    #     # find all relevant state ids
+    #     all_state_ids = np.where(self.state_ids == self.state_ids[state_index])[0]
+    #     all_probs = self.probabilities[all_state_ids]
+    #     all_actions = self.actions[all_state_ids]
+    #     assert (abs(1 - sum(all_probs)) < 0.01)
+    #     action = int(np.random.choice(all_actions, p=all_probs))
+    # else:
+    #     # assume that there is a single action for each state
+    #     # switch between q-values or the policy
+    #     if self.q_values is not None:
+    #         # TODO: np.where rather than .loc[]
+    #         state_q = self.q_values.loc[self.q_values['idstate'] == state_index]
+    #         action = int(state_q.loc[state_q['qvalue'] == state_q.max()[2]]['idaction'])
+    #     else:
+    #         action = int(self.actions[state_index])
+    # return action
 
     def train(self):
         pass
 
     def run(self):
-        pass
+        curr_state = 0
+        for h in tqdm.trange(self.horizon):
+            action = self.choose_action(curr_state)
+            next_state = curr_state * self.domain.num_actions + action
+
 
 
 # showcase the agent performance

### 2020-02-24 codebase/rlessential/domains.py

@@ -1,6 +1,8 @@
 from abc import ABC, abstractmethod
+import warnings
 
 import numpy as np
+
 
 # TODO: migrate self-defined domains in lspi's domain module to here
 
@@ -54,6 +56,8 @@ class BaseMDP(MDP):
         self.mdp = mdp
         self.num_states = self.state_count()
         self.num_actions = self.action_count()
+        self.probs = self.get_probabilities()
+        self.rewards = self.get_rewards()
 
     def state_count(self):
         unique_from = self.mdp['idstatefrom'].nunique()
@@ -64,15 +68,6 @@ class BaseMDP(MDP):
         unique_action = self.mdp['idaction'].nunique()
         return unique_action
 
-    def get_rewards(self):
-        rewards = np.zeros((self.num_states * self.num_actions, 1))
-        for index, row in self.mdp.iterrows():
-            s = row['idstatefrom'].astype(int)
-            a = row['idaction'].astype(int)
-            # any float value, no type casting
-            rewards[s * self.num_actions + a] = row['reward']
-        return rewards
-
     def get_probabilities(self):
         probs = np.zeros((self.num_states * self.num_actions, self.num_states))
         for index, row in self.mdp.iterrows():
@@ -82,6 +77,28 @@ class BaseMDP(MDP):
             # any float value, no type casting
             probs[s * self.num_actions + a][sp] = row['probability']
         return probs
+
+    def get_rewards(self):
+        rewards = np.zeros((self.num_states * self.num_actions, 1))
+        for index, row in self.mdp.iterrows():
+            s = row['idstatefrom'].astype(int)
+            a = row['idaction'].astype(int)
+            # any float value, no type casting
+            rewards[s * self.num_actions + a] = row['reward']
+        return rewards
+
+    def get_possible_actions(self, state):
+        if not self.probs:
+            warnings.warn('Failed to retrieve probabilities!')
+            return None
+        state_actions = list()
+        state_probs = list()
+        for a in range(self.num_actions):
+            p_s = self.probs[state * self.num_actions + a]
+            if p_s != 0:
+                state_actions.append(a)
+                state_probs.append(p_s)
+        return state_actions, state_probs
 
 
 class TwoStateMDP(BaseMDP):
@@ -185,6 +202,7 @@ class MachineReplacementMDP(BaseMDP):
     5 actions: 3 **repairs** and 2 **do nothings**
 
     """
+
     def __init__(self, mdp):
         # noinspection PyCompatibility
         super().__init__(mdp)

### 2020-02-24 codebase/rlessential/main.py

@@ -70,3 +70,5 @@ if __name__ == '__main__':
     vi.calculate_value()
 
 
+
+

### 2020-03-02 minutia/paper/paper.tex
MacRep IQR still in progress

@@
     \textit{RiverSwim} an MDP consists of six states and \textit{SixArms} one with seven states.
-In \textit{RiverSwim} flow of the river is to left and the swimmer picks probabilistically equal either state 1 or 2 as the start point.
-All states but two terminal states have zero reward to land at while the right most one rewards substantially higher than the other one.
+    In \textit{RiverSwim} flow of the river is to left and the swimmer picks probabilistically equal either state 1
+    or 2 as the start point.
+    All states but two terminal states have zero reward to land at while the right most one rewards substantially
+    higher than the other one.
     The agent in SixArms selects between six distinct actions which pulls different arms of a multi-armed bandit.
     By pulling an arm the agent traverses to a new state.
-Although the agent does not obtain reward by pulling the arms but going to the connected state to that arm is highly rewarding in which the transition probability is in inverse relationship with the reward value.
+    Although the agent does not obtain reward by pulling the arms but going to the connected state to that arm is
+    highly rewarding in which the transition probability is in inverse relationship with the reward value.
 
-    As both problems favor keeping the agent in states with lower reward, exploration is paramount to maximize the accumulated reward.
-This is also confirmed by the authors in \cite{Strehl2004} that the agent will fail to learn when a rudimentary $\epsilon$-greedy algorithm.
+    As both problems favor keeping the agent in states with lower reward, exploration is paramount to maximize the
+    accumulated reward.
+    This is also confirmed by the authors in \cite{Strehl2004} that the agent will fail to learn when a rudimentary
+    $\epsilon$-greedy algorithm.
+
+    We use a randomized policy to simulate trajectories and collect samples to attenuate bias in action selection
+    process.
+    As it has shown the literature, $\epsilon$-greedy has a bias in solving exploration-exploitation dilemma, we
+    explore the environment by following a randomized policy with a uniform distribution.
 

### 2020-03-02 codebase/lspi/domains.py

@@ -346,6 +346,7 @@ class RiverSwimDomain(Domain):
 
         next_state = np.array([curr_sample.values[2].astype(int)])
         reward = np.array([curr_sample.values[4].astype(int)])
+
         sample = Sample(self._state.copy(), action, reward, next_state)
 
         self._state = next_state

### 2020-03-02 codebase/rlessential/agents.py

@@ -4,6 +4,8 @@ import numpy as np
 import gym
 import tqdm
 
+from sample import Sample
+
 
 class Agent(object):
     """
@@ -310,48 +312,40 @@ class RAAMAgent:
                         for s_r in transition_reward:
                             sum_rewards += s_r[1]
                         # TODO: check to be correct
-                        self.r[a,b,s] = (1 / len(transition_reward)) * sum_rewards
+                        self.r[a, b, s] = (1 / len(transition_reward)) * sum_rewards
 
 
-class MachineReplacementAgent(Agent):
+class MachineReplacementMDPAgent(Agent):
 
-    def __init__(self, domain, horizon):
+    def __init__(self, domain, solver, discount, horizon=None):
         self.domain = domain
+        self.solver = solver
+        self.discount = discount
+
         self.horizon = horizon
+        self.state_ = 0
 
     def choose_action(self, state):
-        state_probs, state_actions = self.domain.get_possible_actions(state)
-        assert (abs(1 - sum(state_probs)) < 0.01)
-        action = int(np.random.choice(state_actions, p=state_probs))
-        return action
+        # TODO: try epsilon-greedy or Boltzmann distribution
 
-    # if self.probabilities is not None:
-    #     # find all relevant state ids
-    #     all_state_ids = np.where(self.state_ids == self.state_ids[state_index])[0]
-    #     all_probs = self.probabilities[all_state_ids]
-    #     all_actions = self.actions[all_state_ids]
-    #     assert (abs(1 - sum(all_probs)) < 0.01)
-    #     action = int(np.random.choice(all_actions, p=all_probs))
-    # else:
-    #     # assume that there is a single action for each state
-    #     # switch between q-values or the policy
-    #     if self.q_values is not None:
-    #         # TODO: np.where rather than .loc[]
-    #         state_q = self.q_values.loc[self.q_values['idstate'] == state_index]
-    #         action = int(state_q.loc[state_q['qvalue'] == state_q.max()[2]]['idaction'])
-    #     else:
-    #         action = int(self.actions[state_index])
-    # return action
+        allowed_actions = self.domain.get_allowed_actions()
+
+        # randomized action selection - uniform distribution
+        return np.random.choice(allowed_actions)
 
     def train(self):
-        pass
+        self.solver.calculate_value()
 
     def run(self):
-        curr_state = 0
-        for h in tqdm.trange(self.horizon):
+        curr_state = self.domain.state_
+        samples = list()
+        i = 0
[…]

### 2020-03-02 codebase/rlessential/domains.py

@@ -1,14 +1,20 @@
 from abc import ABC, abstractmethod
-import warnings
 
 import numpy as np
 
+from sample import Sample
 
 # TODO: migrate self-defined domains in lspi's domain module to here
 
+COL_STATE_FROM = 'idstatefrom'
+COL_ACTION = 'idaction'
+COL_STATE_TO = 'idstateto'
+COL_PROBABILITY = 'probability'
+COL_REWARD = 'reward'
+
 
 class MDP(ABC):
-    # same as new in 3.4:
+    # new in Python 3.4:
     # __metaclass__ = ABCMeta
 
     @abstractmethod
@@ -49,56 +55,115 @@ class MDP(ABC):
         """
         pass  # pragma: no cover
 
+    @abstractmethod
+    def step(self, action):
+        """
+        Applies the action and returns the observed sample.
+
+        :arg action Action index. Must be in range [0, num_actions)
+
+        :return: observed sample
+        """
+
+        pass  # pragma: no cover
+
+    @abstractmethod
+    def reset(self, initial_state):
+        """
+        Resets the domain to the initial state to start from.
+
+        :arg initial_state the starting state
+        """
+
+        pass  # pragma: no cover
+
 
 class BaseMDP(MDP):
 
-    def __init__(self, mdp):
+    def __init__(self, mdp, initial_state=None):
         self.mdp = mdp
         self.num_states = self.state_count()
         self.num_actions = self.action_count()
         self.probs = self.get_probabilities()
         self.rewards = self.get_rewards()
 
+        self.state_ = None
+        self.reset(initial_state)
+
     def state_count(self):
-        unique_from = self.mdp['idstatefrom'].nunique()
-        unique_to = self.mdp['idstateto'].nunique()
+        unique_from = self.mdp[COL_STATE_FROM].nunique()
+        unique_to = self.mdp[COL_STATE_TO].nunique()
         return max(unique_from, unique_to)
 
     def action_count(self):
-        unique_action = self.mdp['idaction'].nunique()
+        unique_action = self.mdp[COL_ACTION].nunique()
         return unique_action
 
     def get_probabilities(self):
         probs = np.zeros((self.num_states * self.num_actions, self.num_states))
         for index, row in self.mdp.iterrows():
-            s = row['idstatefrom'].astype(int)
-            a = row['idaction'].astype(int)
-            sp = row['idstateto'].astype(int)
+            s = row[COL_STATE_FROM].astype(np.int)
[…]

### 2020-03-02 codebase/rlessential/main.py

@@ -3,7 +3,7 @@ import pandas as pd
 import numpy as np
 
 from rlessential.domains import MachineReplacementMDP
-from agents import MachineReplacementAgent
+from agents import MachineReplacementMDPAgent
 from rlessential.solvers import ValueIteration
 
 
@@ -54,20 +54,19 @@ if __name__ == '__main__':
 
     mdp = pd.read_csv('dataset/mdp/machine_replacement_mdp.csv')
 
-    domain_mr = MachineReplacementMDP(mdp)
-    agent_mr = MachineReplacementAgent(domain_mr)
-
-
     gamma = 0.90
     epsilon = 0.0000001
     tau = (epsilon * (1 - gamma)) / (2 * gamma)
 
-    # even different initial values will end up with the same state values!
-    init_val = np.arange(10)
-    init_val = np.random.rand(10) * 10
+    domain_mr = MachineReplacementMDP(mdp)
+    solver_vi = ValueIteration(domain_mr, discount=gamma, threshold=tau, verbose=True)
+    agent_mr = MachineReplacementMDPAgent(domain_mr, solver_vi, discount=gamma, horizon=100)
+    agent_mr.run()
 
-    vi = ValueIteration(domain_mr, discount=gamma, initial_values=init_val, threshold=tau, verbose=True)
-    vi.calculate_value()
+    # # even different initial values will end up with the same state values!
+    # init_val = np.arange(10)
+    # init_val = np.random.rand(10) * 10
+
 
 
 

### 2020-03-02 codebase/rlessential/sample.py

@@ -1,0 +1,16 @@
+class Sample:
+
+    def __init__(self, current_state, action, reward, next_state, terminal=False):
+        self.current_state = current_state
+        self.action = action
+        self.reward = reward
+        self.next_state = next_state
+        self.terminal = terminal
+
+    def __repr__(self):
+        """Create string representation of tuple."""
+        return 'Sample({}, {}, {}, {}, {})'.format(self.current_state,
+                                                   self.action,
+                                                   self.reward,
+                                                   self.next_state,
+                                                   self.terminal)

### 2020-03-02 codebase/rlessential/solvers.py

@@ -22,6 +22,15 @@ class ValueIteration:
         # format: (#iter, V(s), distance)
         self.iter_values = list()
 
+    def v_table(self):
+        if not self.iter_values:
+            print('Value iteration has not been run yet!')
+            self.calculate_value()
+
+        num_s = self.domain.num_states
+        last_v = self.iter_values[-1][1].reshape(num_s, 1)
+        return last_v
+
     def calculate_value(self):
         num_s = self.domain.num_states
         num_a = self.domain.num_actions

### 2020-03-02 codebase/rlessential/agents.py
IQR  in progress

@@ -324,6 +324,7 @@ class MachineReplacementMDPAgent(Agent):
 
         self.horizon = horizon
         self.state_ = 0
+        self.samples = list()
 
     def choose_action(self, state):
         # TODO: try epsilon-greedy or Boltzmann distribution
@@ -338,13 +339,12 @@ class MachineReplacementMDPAgent(Agent):
 
     def run(self):
         curr_state = self.domain.state_
-        samples = list()
         i = 0
         while i < self.horizon:
             action = self.choose_action(curr_state)
             sample = self.domain.step(action)
             curr_state = sample.next_state
-            samples.append(sample)
+            self.samples.append(sample)
             i += 1
 
 

### 2020-03-02 codebase/rlessential/domains.py

@@ -122,7 +122,7 @@ class BaseMDP(MDP):
         if state is None:
             state = self.state_
 
-        candidate_transitions = self.mdp.iloc[state.item() * self.num_actions: (state.item() + 1) * self.num_actions, :]
+        candidate_transitions = self.mdp.loc[self.mdp[COL_STATE_FROM] == state.item()]
         candidate_transitions = candidate_transitions.loc[candidate_transitions[COL_PROBABILITY] != 0]
         if len(candidate_transitions) == 0:
             raise KeyError('Not able to find any candidates for state {}'.format(state))

### 2020-03-02 codebase/rlessential/main.py

@@ -1,6 +1,7 @@
 
 import pandas as pd
 import numpy as np
+from scipy.stats import iqr
 
 from rlessential.domains import MachineReplacementMDP
 from agents import MachineReplacementMDPAgent
@@ -50,6 +51,33 @@ def cartpole_iqr_compare():
     agent_hist.run(False)
 
 
+def samples_preprocess(samples, verbose=True):
+    state_samples = [sample_.current_state for sample_ in samples]
+    state_samples = np.stack(state_samples, axis=0)
+
+    if verbose:
+        print('{} state samples are processed.'.format(len(state_samples)))
+        print('output samples shape: {}'.format(state_samples.shape))
+
+    return state_samples
+
+
+def discretize_samples(samples):
+    iq_range = iqr(samples, axis=0)
+    bin_width = 2 * iq_range * (len(samples) ** (-1 / 3))
+
+    hi = np.max(samples, axis=0)
+    lo = np.min(samples, axis=0)
+
+    counts = (hi - lo) / bin_width
+    counts = np.round(counts)
+    counts = counts.astype(int)
+
+    num_buckets = tuple(counts)
+
+    return num_buckets
+
+
 if __name__ == '__main__':
 
     mdp = pd.read_csv('dataset/mdp/machine_replacement_mdp.csv')
@@ -58,16 +86,22 @@ if __name__ == '__main__':
     epsilon = 0.0000001
     tau = (epsilon * (1 - gamma)) / (2 * gamma)
 
+
     domain_mr = MachineReplacementMDP(mdp)
     solver_vi = ValueIteration(domain_mr, discount=gamma, threshold=tau, verbose=True)
     agent_mr = MachineReplacementMDPAgent(domain_mr, solver_vi, discount=gamma, horizon=100)
     agent_mr.run()
 
+    samples = agent_mr.samples
+    samples = discretize_samples(samples)
+
+
+
+
+
+
+
     # # even different initial values will end up with the same state values!
     # init_val = np.arange(10)
     # init_val = np.random.rand(10) * 10
-
-
-
-
 

### 2020-03-02 codebase/rlessential/sample.py

@@ -9,7 +9,7 @@ class Sample:
 
     def __repr__(self):
         """Create string representation of tuple."""
-        return 'Sample({}, {}, {}, {}, {})'.format(self.current_state,
+        return 'Sample({}, {}, {}, {}, {})\n'.format(self.current_state,
                                                    self.action,
                                                    self.reward,
                                                    self.next_state,

### 2020-03-09 minutia/paper/paper.tex
IQR  in progress
Paper updated befor the meeting

@@
 
-    where n is number of observations and IQR is the interquartile range of data.
+    where n is number of observations and $IQR(x)$ is the inter-quartile range of data.
 
@@
     policy-space solutions.
+
+    The objective is finding an optimal aggregate model which falls in the span of spectrum from a fine-grid model
+    to a rigid aggregate model.
+    Our presented method calculates a discretization policy for which we calculate a feasible value function.
+    Return of a policy shows the quality of the gauged policy.
+    Hence, by calculating the accumulative return (eq.\ref{eq:return}) over run of the \textit{true} model and in a
+    comparison with the original return we can measure the improvement.
+
+    \begin{equation}
+        R = \sum_{k=0}^{\infty} \gamma^{k} r_{k+1}. \label{eq:return}
+    \end{equation}
 

### 2020-03-09 codebase/rlessential/main.py
sample preprocessing

@@ -86,22 +86,18 @@ if __name__ == '__main__':
     epsilon = 0.0000001
     tau = (epsilon * (1 - gamma)) / (2 * gamma)
 
-
     domain_mr = MachineReplacementMDP(mdp)
     solver_vi = ValueIteration(domain_mr, discount=gamma, threshold=tau, verbose=True)
     agent_mr = MachineReplacementMDPAgent(domain_mr, solver_vi, discount=gamma, horizon=100)
     agent_mr.run()
 
     samples = agent_mr.samples
-    samples = discretize_samples(samples)
+    samples = samples_preprocess(samples)  # extract states from the samples
+    num_buckets = discretize_samples(samples)  # range calculation
+
+    vals = solver_vi()
 
 
 
 
-
-
-
-    # # even different initial values will end up with the same state values!
-    # init_val = np.arange(10)
-    # init_val = np.random.rand(10) * 10
 

### 2020-03-09 codebase/rlessential/main.py
Mac Rep MDP fixed

@@ -99,5 +99,3 @@ if __name__ == '__main__':
 
 
 
-
-

### 2020-03-10 codebase/rlessential/main.py
MacRep aggregation

@@ -86,7 +86,6 @@ if __name__ == '__main__':
     epsilon = 0.0000001
     tau = (epsilon * (1 - gamma)) / (2 * gamma)
 
-
     domain_mr = MachineReplacementMDP(mdp)
     solver_vi = ValueIteration(domain_mr, discount=gamma, threshold=tau, verbose=True)
     agent_mr = MachineReplacementMDPAgent(domain_mr, solver_vi, discount=gamma, horizon=100)

### 2020-03-11 minutia/paper/paper.tex
Paper updated and new paper initiated
paper agg state clustering

@@
     \begin{equation}
-        \text { Bin width }=2 \frac{\operatorname{IQR}(x)}{\sqrt[3]{n}}
+        |S_{agg}| = 2 \frac{\operatorname{IQR}(x)}{\sqrt[3]{n}}
     \end{equation}
 
-    where n is number of observations and $IQR(x)$ is the inter-quartile range of data.
+    where n is number of observations and $IQR(x)$ is the inter-quartile range of data.f
+
+    By having the inter-quartile range of data for each feature, a discretization policy can be calculated.
 
@@
     \end{equation}
+
+    The original states with the closest \textit{calculated} values forgather to build aggregate states.
+    For this purpose, state values are clustered by K-means analysis to calculate the relevant aggregate states.
 

### 2020-03-11 codebase/rlessential/domains.py
state values clustering

@@ -140,7 +140,7 @@ class BaseMDP(MDP):
         if len(candidate_transitions) == 0:
             raise KeyError('Not able to find any candidates for state-action pair ({}, {})'.format(self.state_, action))
         elif len(candidate_transitions) != 1:
-            candidate_transitions = self.mdp.sample(candidate_transitions, weights=COL_PROBABILITY)
+            candidate_transitions = candidate_transitions.sample(weights=COL_PROBABILITY)
 
         transition_reward = candidate_transitions[COL_REWARD].values.reshape(1, )
         transition_next_state = candidate_transitions[COL_STATE_TO].values.reshape(1, ).astype(np.int)

### 2020-03-11 codebase/rlessential/main.py

@@ -2,10 +2,12 @@
 import pandas as pd
 import numpy as np
 from scipy.stats import iqr
+from sklearn.cluster import KMeans
 
 from rlessential.domains import MachineReplacementMDP
 from agents import MachineReplacementMDPAgent
 from rlessential.solvers import ValueIteration
+from util import RANDOM_SEED
 
 
 def cartpole_iqr_compare():
@@ -51,7 +53,7 @@ def cartpole_iqr_compare():
     agent_hist.run(False)
 
 
-def samples_preprocess(samples, verbose=True):
+def preprocess_samples(samples, verbose=True):
     state_samples = [sample_.current_state for sample_ in samples]
     state_samples = np.stack(state_samples, axis=0)
 
@@ -69,13 +71,15 @@ def discretize_samples(samples):
     hi = np.max(samples, axis=0)
     lo = np.min(samples, axis=0)
 
-    counts = (hi - lo) / bin_width
-    counts = np.round(counts)
-    counts = counts.astype(int)
+    bin_count = (hi - lo) / bin_width
+    bin_count = np.round(bin_count).astype(int)
+    bin_count = bin_count.item()
 
-    num_buckets = tuple(counts)
+    return bin_count
 
-    return num_buckets
+
+def discretize_values(values):
+    pass
 
 
 if __name__ == '__main__':
@@ -85,17 +89,20 @@ if __name__ == '__main__':
     gamma = 0.90
     epsilon = 0.0000001
     tau = (epsilon * (1 - gamma)) / (2 * gamma)
+    steps = 1000
 
     domain_mr = MachineReplacementMDP(mdp)
     solver_vi = ValueIteration(domain_mr, discount=gamma, threshold=tau, verbose=True)
-    agent_mr = MachineReplacementMDPAgent(domain_mr, solver_vi, discount=gamma, horizon=100)
-    agent_mr.run()
+    agent_mr = MachineReplacementMDPAgent(domain_mr, solver_vi, discount=gamma, horizon=steps)
 
+    agent_mr.train()
+    values = solver_vi.get_v_table()
+
+    agent_mr.run()
     samples = agent_mr.samples
-    samples = samples_preprocess(samples)  # extract states from the samples
+    samples = preprocess_samples(samples)  # extract states from the samples
     num_buckets = discretize_samples(samples)  # range calculation
 
-    vals = solver_vi()
-
-
-
[…]

### 2020-03-11 codebase/rlessential/solvers.py

@@ -22,7 +22,7 @@ class ValueIteration:
         # format: (#iter, V(s), distance)
         self.iter_values = list()
 
-    def v_table(self):
+    def get_v_table(self):
         if not self.iter_values:
             print('Value iteration has not been run yet!')
             self.calculate_value()
@@ -88,7 +88,7 @@ class ValueIteration:
         p = self.domain.get_probabilities()
         sum_probs = np.sum(p, axis=1, keepdims=True)
 
-        last_v = self.iter_values[-1][1].reshape(num_s, 1)
+        last_v = self.get_v_table()
         # noinspection PyCompatibility
         last_v = r + self.discount * p @ last_v
         split_last_values = np.split(last_v, num_s)

### 2020-03-11 codebase/util.py

@@ -6,6 +6,7 @@ import time
 import pickle
 
 OUTPUT_DIR = 'outputs'
+RANDOM_SEED = 42
 
 
 def timing(t):

### 2020-03-16 codebase/rlessential/main.py
value function clustering done

@@ -106,3 +106,6 @@ if __name__ == '__main__':
     kmeans = KMeans(n_clusters=num_buckets, random_state=RANDOM_SEED)
     kmeans.fit(values)
     kmeans.labels_
+
+
+

### 2020-03-16 minutia/paper/paper.tex
paper: clustering for aggregation added

@@
 
+    \subsection{Aggregation}
     The original states with the closest \textit{calculated} values forgather to build aggregate states.
     For this purpose, state values are clustered by K-means analysis to calculate the relevant aggregate states.
+    State aggregation merge similar transitions based on a triple $<s, s', a>$ where $s$ indicates the starting state,
+    $s'$ is the ending state, and the take action in that transition represented by $a$.
+    In the process of stacking similar transitions, we here presume all inbound transitions as outbound transition of
+    the ending state.
+    This assumption render all transitions as outbound ones.
 

### 2020-03-17 codebase/rlessential/consts.py
aggregation completed

@@ -1,0 +1,7 @@
+# bookkeeping
+
+COL_STATE_FROM = 'idstatefrom'
+COL_ACTION = 'idaction'
+COL_STATE_TO = 'idstateto'
+COL_PROBABILITY = 'probability'
+COL_REWARD = 'reward'

### 2020-03-17 codebase/rlessential/domains.py

@@ -3,14 +3,9 @@ from abc import ABC, abstractmethod
 import numpy as np
 
 from sample import Sample
+from consts import *
 
 # TODO: migrate self-defined domains in lspi's domain module to here
-
-COL_STATE_FROM = 'idstatefrom'
-COL_ACTION = 'idaction'
-COL_STATE_TO = 'idstateto'
-COL_PROBABILITY = 'probability'
-COL_REWARD = 'reward'
 
 
 class MDP(ABC):

### 2020-03-17 codebase/rlessential/main.py

@@ -1,3 +1,4 @@
+from collections import defaultdict
 
 import pandas as pd
 import numpy as np
@@ -8,6 +9,7 @@ from rlessential.domains import MachineReplacementMDP
 from agents import MachineReplacementMDPAgent
 from rlessential.solvers import ValueIteration
 from util import RANDOM_SEED
+from consts import *
 
 
 def cartpole_iqr_compare():
@@ -78,20 +80,65 @@ def discretize_samples(samples):
     return bin_count
 
 
-def discretize_values(values):
-    pass
+def cluster_values(values, num_buckets, random_seed):
+    kmeans = KMeans(n_clusters=num_buckets, random_state=random_seed)
+    kmeans.fit(values)
+
+    return kmeans.labels_
+
+
+def aggregate(mdp, labels):
+    # mapping
+    agg_map = {i: v for i, v in enumerate(labels)}
+
+    # original rewards
+    rewards_orig = {i: r for i, r in zip(domain_mr.mdp[COL_STATE_TO], domain_mr.mdp[COL_REWARD])}
+
+    agg_state = list(set(agg_map.values()))
+
+    # aggregate rewards
+    rewards_agg = defaultdict(int)
+    for k, v in agg_map.items():
+        rewards_agg[v] += rewards_orig[k]
+
+    # get a deep copy
+    df_agg = mdp.copy(deep=True)
+
+    # state mapping
+    # should be done only once!
+    df_agg.loc[:, [COL_STATE_FROM, COL_STATE_TO]] = \
+        df_agg.loc[:, [COL_STATE_FROM, COL_STATE_TO]].replace(agg_map)
+
+    # reward mapping
+    for k, v in rewards_agg.items():
+        df_agg.loc[df_agg[COL_STATE_TO] == k, COL_REWARD] = v
+
+    # probability mapping
+    for s in agg_state:
+        for a in range(domain_mr.num_actions):
+            for sp in agg_state:
+                cond = ((df_agg[COL_STATE_FROM] == s) & (df_agg[COL_ACTION] == a) & (df_agg[COL_STATE_TO] == sp))
+                df = df_agg.loc[cond]
+                if len(df) != 0:
+                    prob = df[COL_PROBABILITY].mean()
+                    df_agg.loc[cond, COL_PROBABILITY] = prob
+
+    df_agg.drop_duplicates(inplace=True)
[…]

### 2020-03-17 minutia/paper/paper.tex
paper: aggregation part completed

@@
     State aggregation is an approximation approach to make a problem more tractable by a reduction in dimensionality in
-    state space \footnote{aggre
-    gation in action space although is possible, has less significance comparatively.}.
+    state space \footnote{aggregation in action space although is possible, has less significance comparatively.}.
     State aggregation, in particular, is a parametric feature-based approximation where the features are membership
@@
     the ending state.
-    This assumption render all transitions as outbound ones.
+    This assumption render all transitions as outbound ones, therefore we assign the average probability as the
+    transition probability in the aggregate model.
 

### 2020-07-12 codebase/rlessential/agents.py
essentialrl: visualization

@@ -42,13 +42,14 @@ class BaseAgent(Agent):
         self.samples = list()
 
     def choose_action(self, state, external_policy=None):
-        # randomized action selection - uniform distribution
-        if external_policy == 'randomized':
-            allowed_actions = self.domain.get_allowed_actions()
-            action = np.random.choice(allowed_actions)
-            return action
         # external policy is provided
-        elif external_policy is not None:
+        if external_policy is not None:
+            # randomized action selection - uniform distribution
+            if type(external_policy) is str and external_policy == 'randomized':
+                allowed_actions = self.domain.get_allowed_actions()
+                action = np.random.choice(allowed_actions)
+                return action
+
             if type(external_policy) is not np.ndarray:
                 raise TypeError('external policy is not a valid ndarray')
             if external_policy.shape[0] != self.domain.num_states:

### 2020-07-12 codebase/rlessential/main.py

@@ -1,4 +1,5 @@
 import numpy as np
+from matplotlib import pyplot as plt
 from scipy.stats import iqr
 from sklearn.cluster import KMeans
 
@@ -13,6 +14,20 @@ epsilon = 0.0001
 tau = (epsilon * (1 - gamma)) / (2 * gamma)
 
 steps = 3000
+
+
+def average_rolling(array):
+    """
+    Calculates moving(rolling) average.
+
+    :param array: input array
+    :return: calculated cumulative moving average
+    """
+    cumulative_sum = array.cumsum()
+    a = np.arange(array.size)
+    a = a + 1
+    rolling_mean = cumulative_sum / a
+    return rolling_mean
 
 
 def extract_states(samples):
@@ -182,6 +197,69 @@ def run_machine_replacement():
     print('original return:', rewards.sum())
     print('aggregate return:', rewards_aggregate.sum())
     print('adapted return:', rewards_aggregate_adapted.sum())
+
+
+def visualize_rewards(rewards, rewards_aggregate, rewards_aggregate_adapted):
+    titles = {'original': 'Original rewards', 'aggregate': 'Aggregate rewards', 'adapted': 'Adapted aggregate rewards'}
+
+    def plot_steps(original, aggregate, adapted):
+        fig, ax = plt.subplots()
+        ax.set_xticks([], [])
+        ax.set_yticks([], [])
+
+        ax1 = fig.add_subplot(311)
+        ax1.plot(original, color='r', label=titles['original'])
+        ax1.set_title(titles['original'])
+        ax1.set_xticks([], [])
+
+        ax2 = fig.add_subplot(312)
+        ax2.plot(aggregate, color='b', label=titles['aggregate'])
+        ax2.set_title(titles['aggregate'])
+        ax2.set_xticks([], [])
+        ax2.set_ylabel('Rewards')
+
+        ax3 = fig.add_subplot(313)
+        ax3.plot(adapted, color='g', label=titles['adapted'])
+        ax3.set_title(titles['adapted'])
+
+        plt.xlabel('Steps')
+        return plt
+
+    def plot_cumulative(original, aggregate, adapted):
+        plt.figure()
+        plt.plot(original.cumsum(),
+                 color='r', label=titles['original'])
+        plt.plot(aggregate.cumsum(),
+                 color='b', label=titles['aggregate'])
+        plt.plot(adapted.cumsum(),
+                 color='g', label=titles['adapted'])
+        plt.xlabel('Steps')
[…]

### 2020-07-12 codebase/util.py

@@ -1,12 +1,9 @@
 """
-Set of general utility functions in Python programming.
+Set of general utility functions in Python development.
 """
 import os
 import time
 import pickle
-
-OUTPUT_DIR = 'outputs'
-RANDOM_SEED = 42
 
 
 def timing(t):

### 2020-07-12 codebase/rlessential/main.py
more domains added

@@ -5,7 +5,7 @@ from sklearn.cluster import KMeans
 
 import consts
 from agents import BaseAgent
-from rlessential.domains import MachineReplacementMDPDomain
+from rlessential.domains import MachineReplacementMDPDomain, RiverSwimMDPDomain, SixArmsMDPDomain
 from rlessential.solvers import ValueIterationSolver
 
 # hyper-parameters
@@ -156,8 +156,8 @@ def aggregate_mdp(values, bin_count, domain):
     return df, aggregation_states
 
 
-def run_machine_replacement():
-    domain = MachineReplacementMDPDomain()
+def run(mdp_domain):
+    domain = mdp_domain()
     solver = ValueIterationSolver(domain,
                                   discount=gamma,
                                   threshold=tau,
@@ -176,7 +176,7 @@ def run_machine_replacement():
                                                        bin_count=bucket_count,
                                                        domain=domain)
 
-    domain_aggregate = MachineReplacementMDPDomain(mdp_aggregate)
+    domain_aggregate = mdp_domain(mdp_aggregate)
     solver_aggregate = ValueIterationSolver(domain=domain_aggregate,
                                             discount=gamma,
                                             threshold=tau,
@@ -197,6 +197,8 @@ def run_machine_replacement():
     print('original return:', rewards.sum())
     print('aggregate return:', rewards_aggregate.sum())
     print('adapted return:', rewards_aggregate_adapted.sum())
+
+    return rewards, rewards_aggregate, rewards_aggregate_adapted
 
 
 def visualize_rewards(rewards, rewards_aggregate, rewards_aggregate_adapted):
@@ -263,4 +265,7 @@ def visualize_rewards(rewards, rewards_aggregate, rewards_aggregate_adapted):
 
 
 if __name__ == '__main__':
-    run_machine_replacement()
+    machine_replacement_domain = MachineReplacementMDPDomain
+    river_swim_domain = RiverSwimMDPDomain
+    six_arms_domain = SixArmsMDPDomain
+    run(machine_replacement_domain)

### 2020-07-12 codebase/rlessential/main.py
piped rewards to visualization

@@ -268,4 +268,5 @@ if __name__ == '__main__':
     machine_replacement_domain = MachineReplacementMDPDomain
     river_swim_domain = RiverSwimMDPDomain
     six_arms_domain = SixArmsMDPDomain
-    run(machine_replacement_domain)
+    original_rewards, aggregate_rewards, adapted_rewards = run(machine_replacement_domain)
+    visualize_rewards(original_rewards, aggregate_rewards, adapted_rewards)

### 2020-07-12 minutia/icml/workshop.tex
minutia comprehensive clear up commit

%%%%%%%% ICML 2020 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%

\documentclass{article}

% Recommended, but optional, packages for figures and better typesetting:
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs} % for professional tables

% hyperref makes hyperlinks in the resulting PDF.
% If your build breaks (sometimes temporarily if a hyperlink spans a page)
% please comment out the following usepackage line and replace
% \usepackage{icml2020} with \usepackage[nohyperref]{icml2020} above.
\usepackage{hyperref}

% Attempt to make hyperref and algorithmic work together better:
\newcommand{\theHalgorithm}{\arabic{algorithm}}

% Use the following line for the initial blind version submitted for review:
%\usepackage{icml2020}

% If accepted, instead use the following line for the camera-ready submission:
\usepackage[accepted]{icml2020}
%\usepackage[preprint]{icml2020}

% SHi-ON packages
% argmin argmax begin
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{xcolor}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Submission and Formatting Instructions for ICML 2020}

\begin{document}

\twocolumn[
\icmltitle{Fast Uniform [Distributional] State Aggregation \\ in Reinforcement Learning}

% It is OKAY to include author information, even for blind
% submissions: the style file will automatically remove it for you
% unless you've provided the [accepted] option to the icml2020
% package.

% List of affiliations: The first argument should be a (short)
% identifier you will use later to specify author affiliations
% Academic affiliations should list Department, University, City, Region, Country
% Industry affiliations should list Company, City, Region, Country

% You can specify symbols, otherwise they are numbered in order.
% Ideally, you should not use this facility. Affiliations will be numbered
% in order of appearance and this is the preferred way.
\icmlsetsymbol{equal}{*}

\begin{icmlauthorlist}
\icmlauthor{Authors}{equal,to}
%\icmlauthor{Bauiu C.~Yyyy}{equal,to,goo}
%\icmlauthor{Cieua Vvvvv}{goo}
%\icmlauthor{Iaesut Saoeu}{ed}
%\icmlauthor{Fiuea Rrrr}{to}
%\icmlauthor{Tateu H.~Yasehe}{ed,to,goo}
%\icmlauthor{Aaoeu Iasoh}{goo}
%\icmlauthor{Buiui Eueu}{ed}
%\icmlauthor{Aeuia Zzzz}{ed}
%\icmlauthor{Bieea C.~Yyyy}{to,goo}
%\icmlauthor{Teoau Xxxx}{ed}
%\icmlauthor{Eee Pppp}{ed}
\end{icmlauthorlist}

\icmlaffiliation{to}{Department of Computer Science, University of New Hampshire}
%\icmlaffiliation{goo}{Googol ShallowMind, New London, Michigan, USA}
%\icmlaffiliation{ed}{School of Computation, University of Edenborrow, Edenborrow, United Kingdom}

\icmlcorrespondingauthor{Authors}{c.vvvvv@googol.com}
%\icmlcorrespondingauthor{Eee Pppp}{ep@eden.co.uk}

% You may provide any keywords that you
% find helpful for describing your paper; these are used to populate
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{Machine Learning, ICML}

\vskip 0.3in
]

% this must go after the closing bracket ] following \twocolumn[ ...

% This command actually creates the footnote in the first column
% listing the affiliations and the copyright notice.
% The command takes one argument, which is text to display at the start of the footnote.
% The \icmlEqualContribution command is standard text for equal contribution.
% Remove it (just {}) if you do not need this facility.

%\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.

\begin{abstract}
    Reinforcement learning (RL) problems often incur large scale state spaces needing sizable amount
    of computation to process.
    State aggregation as an analytically-transparent mean approximates and yields lower-dimension state spaces.
%   We present a state aggregation algorithm and a learning algorithm based on the former to solve a stochastic
%   problem.
    We present a state aggregation algorithm method in Markov decision processes (MDP) inspired by histograms.
    Our approach explores a \textit{splitting policy} in direct relationship with inter-quartile range
    of collected samples.
    We then accompanied an experiment by which we show the practical promises of the approach to achieve superior
    results.
\end{abstract}

\section{Introduction}
\label{sec:intro}
    Real-world problems embody high-dimensional state spaces where in computational issues
    arise.
    Various approximation methods are proposed in the literature to reduce the resulted value function
    dimensionality \cite{Sutton1998, Busoniu2010, Francois-Lavet2018}.
    State aggregation is one of the simplest and most transparent state representation approximation
    method.

    State aggregation, at the core level, is to discretize together similar states to ultimately yield a
    lower-dimension MDP. The reduced MDP requires considerably less amount of computational iterations or samples to
    solve, which improves convergence to the .

    Although state aggregation may not generalize as well as neural network-based methods, the techniques has both
    analytical and practical advantages \cite{Lagoudakis2003}.
    In comparison, state aggregation takes less computational power to calculate representative features.
    Moreover, analysis and troubleshooting in a state aggregation process is more obvious than a black-box neural
    network.
    Approximation is a corner stone in all successful RL methods to hold a "good" mapping between a large state
    space to a much smaller one where in value function is computable or at least feasible.
    These "good" approximators can end up with "good" features to represent the projected state space.

    The cross-fertilization in using state aggregation and deep neural networks in solving immense state
    space problems in RL is also profound.
    While the newly emerged methods are growing at a staggering pace, a parallel area of research has been formed
    wherein a variety of techniques have been utilized to make these methods contravene their promised functionality.

    Our contribution is a bifurcation.
    We first introduce an algorithm for aggregation in MDPs, then we move on to show advantages of taking such
    aggregation method to work.
    Hereby, we explain the paper structure in brief.
    We first lay down the core concepts of the MDP and the aggregation framework in Sections \ref{sec:mdp} and
    \ref{sec:state-aggregation} then continue to describe the methodology in Section \ref{sec:method}.

\section{Related Work}
    \label{sec:related-work}
    Aggregation was extensively appeared throughout the operations and computations research literature
    \cite{Chatelin1982, Rogers1991, Douglas1993}.
    It then introduced to the optimization by an iterative approach in linear programming \cite{Mendelssohn1982}.
    On the end of this spectrum, dynamic programming (DP) leveraged state aggregation to reduce computational size of
    the problem \cite{Bean1987}, as well as using an extension to state aggregation, \textit{soft state
    aggregation}, in an approximate value iteration method \cite{Singh1995}, and general approximate DP frameworks
    \cite{Gordon1995, Tsitsiklis1996}.

    In the recent body of work, state aggregation has applications in conjunction to other concepts such as
    temporal aggregation, so called \textit{options} \cite{Ciosek2015}, \textit{bottleneck simulator}
    \cite{Serban2018}, and also in continuous space optimal control \cite{Zhong2011}.

    Error analyses and regret bounds incured by state aggregation methods are comprehensively discussed by
    \citet{VanRoy2006} and \citet{Petrik2014}.
    Near optimality criterion in aggregation is also investigated in \citet{Bernstein2008}.

    It is noteworthy to point out that, state aggregation exists in the literature under different co-hyponyms.

\section{Markov Decision Processes}
\label{sec:mdp}
    Reinforcement learning problems can be formulated a Markov Decision Processes (MDP)
    framework \cite{Puterman1994, Sutton1998}.
    MDPs facilitate handling sequential decision-making under uncertainty.
    Discrete discounted reward criterion is the center of focus for ease of exposition in this paper.
    For an environment and a decision maker, consider the tuple $<>$ state space $S$ and action space $A$ are defined,
respectively.
    The transition probabilities $P$ describe how likely is the next state to incur by taking a specific action
    $a$ by the agent in a certain state $s$ which will later receive reward $r$.
    For a taken action $a_i \in A$ at the state $s_i \in S$, the decision maker will be ended up in state $s_{i+1}
    \in S$ with a likelihood expressed by the transition probabilities $P : S \times A \times S \rightarrow [0, 1]$
    and wil receive reward $r_{i+1}$ from the rewards $R : S \times A \rightarrow \mathbb{R}$.

%    \begin{equation}
%        s_{i+1} = P(s_{i+1}|s_i, a_i)
%    \end{equation}
%
%    \begin{equation}
%        r_{i+1} = R(s_i, a_i)
%    \end{equation}

\section{State Aggregation}
\label{sec:state-aggregation}
    Value function approximation alleviates the computational hurdle of a large state-space problem.
    We first introduce an approximate value iteration method to solve an aggregate problem.
    State aggregation is an approximation approach to make a problem more tractable by a reduction in dimensionality in
    state space \footnote{aggregation in action space although is possible, has less significance comparatively.}.
    State aggregation, in particular, is a parametric feature-based approximation where the features are membership
    functions of 0-1 form.
    Aggregation consolidates similar states together, so called representative states, which allows the
    \textit{aggregate} problem to be solved by an exact method \cite{Bertsekas2019}
    The computed value function from the exact solution of the aggregate problem can be then used in the original
    problem.

    Now we formulate the aggregation by introducing the dynamics of the aggregate problem.
    The transition probabilities between two representative states:

    \begin{equation}
        \hat{P}_{s s^\prime}(a)=\sum_{j=1}^{n} P_{s j}(a) \phi_{j y},
    \end{equation}

    where $p_ {s j}$ is the transition probabilities from a representative state, $s$, to an original state, $j$
    and $\phi_{j s^\prime}$ is the aggregation probabilities from an original state, $j$, to a representative state,
    $s^\prime$.
    The rewards are defined as:

    \begin{equation}
        \hat{r}(s, a)=\sum_{j=1}^{n} P_{s j}(a) r(s, a, j),
    \end{equation}

    where $\hat{r}(s, a)$ is the received reward at state $s$ by taking action $a$.

    Aggregation then helps us to derive the value function.
    Approximated optimal value function is a weighted sum of optimal rewards and can be computed by
    Equation \ref{approx val func} where $r_{s^\prime}^{\ast}$ is optimal reward at represented state $s$.

    \begin{equation}
        \label{approx val func}
        \tilde{V}(j)=\sum_{s^\prime \in \mathcal{U}} \phi_{j s^\prime} r_{s^\prime}^{\ast},
    \end{equation}

\section{Method}
\label{sec:method}
    Splitting the observations into spans of similar observations is one way to interpret an aggregation
    problem.
    Histograms as classical and simple density estimators were around in the literature for decades \cite{Scott1979,
    Scott2015}.
    Inspired by histograms and the statistics behind finding the proper number of bins for a given distribution, we
    build our proposed aggregation method on top of such bases.
    Proper \textit{bin width}, according to the definition, is to find a trade-off to capture similarities in data
    features and diminish variations caused by random sampling \cite{Knuth2019}.

    Finding the bin width for a given distribution could be calculated by Freedman-Diaconis rule
    \cite{Freedman1981}.
    As the rule states:

    \begin{equation}
        |S_{agg}| = 2 \frac{\operatorname{IQR}(x)}{\sqrt[3]{n}}
    \end{equation}

    where n is number of observations and $IQR(x)$ is the inter-quartile range of data.
    This segmentation rule exhibits robustness to sample distribution due to low sensitivity of IQR to outliers and
    heteroscedasticity in the given distribution compared to conventional dispersion measures such as variance or
    standard deviation.
    State values, in this paper, are considered as the similarity metric \textcolor{red}{[measure]} in state
    aggregation.

    By having the inter-quartile range of data for each feature, a discretization policy can be calculated.

    For the sake of clarity, we elaborate on some preliminary definitions here to have a consistent set of notations
    to the end of this paper. We use model-free LSPI and LSTDQ \citep{Lagoudakis2003} to update and
    evaluate the policy of actions, respectively.

    We developed an aggregation scheme to lower the state space dimensionality by assigning same action to the state
    in a neighbourhood, so called nearest neighbour/ piecewise linear aggregation.
    We then solve directly the lower dimension problem which relatively takes less computation.

    The aggregate problem is stochastic even if the original problem is deterministic.
    Once we lay out the aggregate problem framework we can solve the problem by any exact methods either value- or
    policy-space solutions.

    The objective is finding an optimal aggregate model which falls in the span of spectrum from a fine-grid model
    to a rigid aggregate model.
    Our presented method calculates a discretization policy for which we calculate a feasible value function.
    Return of a policy shows the quality of the gauged policy.
    Hence, by calculating the accumulative return (eq.\ref{eq:return}) over run of the \textit{true} model and in a
    comparison with the original return we can measure the improvement.

    \begin{equation}
        R = \sum_{k=0}^{\infty} \gamma^{k} r_{k+1}. \label{eq:return}
    \end{equation}

    \subsection{Aggregation}
    The original states with the closest \textit{calculated} values forgather to build aggregate states.
    For this purpose, state values are clustered by K-means analysis to calculate the relevant aggregate states.
    State aggregation merge similar transitions based on a triple $<s, s', a>$ where $s$ indicates the starting state,
    $s'$ is the ending state, and the take action in that transition represented by $a$.
    In the process of stacking similar transitions, we here presume all inbound transitions as outbound transition of
    the ending state.
    This assumption render all transitions as outbound ones, therefore we assign the average probability as the
    transition probability in the aggregate model.

\section{Experiment}
    We apply the approach to two MDPs proposed by \cite{Strehl2004}.
    \textit{RiverSwim} an MDP consists of six states and \textit{SixArms} one with seven states.
    In \textit{RiverSwim} flow of the river is to left and the swimmer picks probabilistically equal either state 1
    or 2 as the start point.
    All states but two terminal states have zero reward to land at while the right most one rewards substantially
    higher than the other one.
    The agent in SixArms selects between six distinct actions which pulls different arms of a multi-armed bandit.
    By pulling an arm the agent traverses to a new state.
    Although the agent does not obtain reward by pulling the arms but going to the connected state to that arm is
    highly rewarding in which the transition probability is in inverse relationship with the reward value.

    As both problems favor keeping the agent in states with lower reward, exploration is paramount to maximize the
    accumulated reward.
    This is also confirmed by the authors in \cite{Strehl2004} that the agent will fail to learn when a rudimentary
    $\epsilon$-greedy algorithm.

    We use a randomized policy to simulate trajectories and collect samples to attenuate bias in action selection
    process.
    As it has shown the literature, $\epsilon$-greedy has a bias in solving exploration-exploitation dilemma, we
    explore the environment by following a randomized policy with a uniform distribution.

    \subsection{Machine Replacement}
    This domain as a \textit{non-episodic} (continuous) task maintains a set of machines.
    The goal is to find a policy to incur the least repair needed.
    incure

\section{Future Works}
    We developed the presented approach with having reproducibility in mind.
    It could be served as a launching pad for the future contributors who want to surf this newly emerged field deeper.
    Future ideas to pursue is a generative model to be able to generate adversaries from a network.

\bibliography{library}
\bibliographystyle{icml2020}

\end{document}

### 2020-07-12 minutia/paper/paper.tex
minutia comprehensive clear up commit

@@
     \begin{abstract}
-        We present a state aggregation algorithm and a learning algorithm based on the former to solve a stochastic
-        problem.
-        The presented aggregation algorithm uses range of collected samples to provide the proper aggregation
-        \textit{buckets}.
-        Our approach is a batch learning by which we detect the \textit{splitting policy} using inter-quartile range
-        of data for each feature.
-        We then show the amount of robustness we can achieve as a guarantee of performance in presence of those
-        aggregation policies.
+        Real world problems often incur large scale state spaces needing sizable amount of computation to process.
+        State aggregation as an analytically-transparent mean approximates and yields lower-dimension state spaces.
+%        We present a state aggregation algorithm and a learning algorithm based on the former to solve a stochastic
+%        problem.
+        We present a state aggregation algorithm method in Markov decision processes (MDP) inspired by histograms.
+        Our approach explores a \textit{splitting policy} in direct relationship with inter-quartile range
+        of collected samples.
+        We then accompanied an experiment by which we show the practical promises of the approach to achieve superior
+        results.
     \end{abstract}
@@
     Reinforcement learning problems can be formulated using a Markov Decision Processes (MDP)
-    framework\cite{Puterman1994, Bertsekas}.
+    framework\cite{Puterman1994,Sutton1998}.
     MDPs facilitate handling sequential decision-making under uncertainty.
@@
 
+    \subsection{Machine Replacement}
+    This domain as a \textit{non-episodic} (continuous) task maintains a set of machines.
+    The goal is to find a policy to incur the least repair needed.
+    incure
     \section{Future Works}

### 2020-07-12 presentation/GradTalk/main.tex
minutia comprehensive clear up commit

% Golden Spiral can be approximated with Golden ratio or Fibonacci sequence.
% Golden Ratio: a+b/a = a/b

\documentclass{beamer}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}

\usetheme{Cuerna}
\usecolortheme{default}
% default, bluesimplex, lettuce, brick

% slide (frames) numbering
\addtobeamertemplate{navigation symbols}{}{
    \usebeamerfont{footline}%
    \usebeamercolor[fg]{footline}%
    \hspace{1em}%
    \insertframenumber/\inserttotalframenumber
}

\title{TITLE}
\author{Shayan Amani}

\date{DATE 2019}
\institute{Department of Computer Science, University of New Hampshire}

\begin{document}

  \begin{frame}
    \titlepage
  \end{frame}

%%%%%% Slides begin

\begin{frame}
\frametitle{They have said...}

\Huge{``Our virtues and our failings are \textbf{inseparable}, like force and matter. When they separate, \textbf{man is no more}.''}

% \begin{rightflush}
% Nikola Tesla
% \end{rightflush}
\hfill \LARGE{\textit{Nikola Tesla}}

\end{frame}

\begin{frame}
    \frametitle{Who Am I?}

    I...
    \begin{itemize}
        \item finished my bachelor's in electrical engineering in 2015
        \item aspired to develop for my own ideas --at a certain point, I had 200k monthly active users on one of my applications--
        \item started my Ph.D. program at UNH in fall 2017
        \item have been through a journey to figure out what I want
        \item have been experimenting with quite a few areas
        \item ended up working with Prof. Marek Petrik on RL
    \end{itemize}

    \note{journey of discovering new things}
    \note{tiptoeing different waters}
\end{frame}

\begin{frame}
    \frametitle{What Have I Been Dealing With}
    \includegraphics[width=1\columnwidth]{wordcloud.png}

\end{frame}

\begin{frame}
    \frametitle{Outline}

    \begin{itemize}
        \item Problem
        \item Examples
        \item Motivation
        \item Outline
        \item BODY
        \item Q&A
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{State Aggregation}
    \label{fm:sta}
    \framesubtitle{Problem}

    \begin{itemize}
        \item Problem \textit{approximation} 
        \item based on \textit{aggregation}.
        \item in value space
        \item feature-based parametric approximation
    \end{itemize}

\end{frame}

\begin{frame}
    \frametitle{Why State Aggregation?}
    \framesubtitle{Problem}

    \begin{itemize}
        \item less computational power
        \item more tractable problem
        \item analytically transparent approximation
            \begin{itemize}
                \item compared to Neural Networks
            \end{itemize}
    \end{itemize}

\end{frame}

\begin{frame}
    \frametitle{Aggregate State}
    \framesubtitle{Problem}

    \begin{itemize}
        \item special group of multiple [original] states
        \item 
    \end{itemize}

\end{frame}

\begin{frame}
    \frametitle{Nomenclature}
    \framesubtitle{Problem}

    \begin{itemize}
        \item Original State
            \begin{itemize}
                \item 
            \end{itemize}
        \item Aggregate State
    \end{itemize}

\end{frame}

\begin{frame}
    \frametitle{Idea}

\end{frame}

\begin{frame}
    \frametitle{How Many Bins?}
    \framesubtitle{Idea}

    Number of buckets to split original states into:
    \begin{itemize}
        \item range split aggregation
    \end{itemize}

\end{frame}

\begin{frame}
    \frametitle{}

\end{frame}

\begin{frame}
    \frametitle{}

\end{frame}

\begin{frame}
    \frametitle{}

\end{frame}

\begin{frame}
    \frametitle{}

\end{frame}

\begin{frame}
    \frametitle{}

\end{frame}

\begin{frame}
    \frametitle{}

\end{frame}

\end{document}

### 2020-08-20 codebase/rlessential/main.py
rlessential: Updated methods

@@ -1,38 +1,31 @@
 import numpy as np
-from matplotlib import pyplot as plt
-from scipy.stats import iqr, skew
 from sklearn.cluster import KMeans
 
 import consts
+import methods
 from agents import BaseAgent
-from rlessential.domains import MachineReplacementMDPDomain, RiverSwimMDPDomain, SixArmsMDPDomain
-from rlessential.solvers import ValueIterationSolver
+from domains import MachineReplacementMDPDomain, RiverSwimMDPDomain, SixArmsMDPDomain
+from solvers import ValueIterationSolver
+from visualization.plot import visualize_rewards
 
 # hyper-parameters
 GAMMA = 0.90
 EPSILON = 0.0001
 TAU = (EPSILON * (1 - GAMMA)) / (2 * GAMMA)
 
-STEPS = 5000
+# ['sqrt', 'sturge', 'rice', 'doane', 'scott', 'freedman-diaconis', 'shimazaki-shinomoto']
+ALGORITHMS = [{'method': 'sqrt', 'steps': 90},
+              {'method': 'sturge', 'steps': 500},
+              {'method': 'rice', 'steps': 120},
+              {'method': 'doane', 'steps': 125},
 
-# ['scott', 'fd', 'ss']
-WIDTH_METHOD = 'ss'
-# ['base', 'sqrt', 'sturge', 'rice', 'doane']
-COUNT_METHOD = 'base'
+              {'method': 'scott', 'steps': 2600},
+              {'method': 'freedman-diaconis', 'steps': 3750},
+              {'method': 'shimazaki-shinomoto', 'steps': 10000}]
 
-
-def average_rolling(array):
-    """
-    Calculates moving(rolling) average.
-
-    :param array: input array
-    :return: calculated cumulative moving average
-    """
-    cumulative_sum = array.cumsum()
-    a = np.arange(array.size)
-    a = a + 1
-    rolling_mean = cumulative_sum / a
-    return rolling_mean
+i = 6
+METHOD = ALGORITHMS[i]['method']
+STEPS = ALGORITHMS[i]['steps']
 
 
 def extract_states(samples):
@@ -51,122 +44,40 @@ def extract_states(samples):
     return states
 
 
-def calculate_bin_width(samples):
-    def scott(array):
-        width = 3.49 * array.std() / (len(array) ** (1 / 3))
-        return width
+7
 
-    def freedman_diaconis(array):
-        inter_quartile_range = iqr(array, axis=0)
[…]

### 2020-08-20 codebase/rlessential/methods.py

@@ -1,0 +1,95 @@
+import matplotlib.pyplot as plt
+import numpy as np
+from scipy.stats import iqr
+
+
+def base_count(array, width):
+    count = np.ceil((array.max() - array.min()) / width)
+    count = int(count.item())
+    return count
+
+
+def sqrt_choice(array):
+    count = np.ceil(np.sqrt(len(array)))
+    return count
+
+
+def sturge(array):
+    count = np.ceil(np.log2(len(array))) + 1
+    return count
+
+
+def rice(array):
+    count = np.ceil(2 * np.cbrt(len(array)))
+    return count
+
+
+def doane(array):
+    """
+    Doane's method
+
+    :param array: data point samples
+    :type array: np.ndarray
+
+    :return: bin count
+    :rtype: int
+    """
+    n = len(array)
+
+    m_2 = (1 / n) * np.sum((array - array.mean()) ** 2)
+    m_3 = (1 / n) * np.sum((array - array.mean()) ** 3)
+    b_1 = m_3 / (m_2 ** (3 / 2))
+
+    sig_b_1 = np.sqrt(6 * (n - 2) / ((n + 1) * (n + 3)))
+
+    count = 1 + np.ceil(np.log2(n)) + np.ceil(np.log2(1 + (np.abs(b_1) / sig_b_1)))
+    return count
+
+
+def scott(array):
+    width = 3.49 * array.std() / (len(array) ** (1 / 3))
+    return width
+
+
+def freedman_diaconis(array):
+    inter_quartile_range = iqr(array, axis=0)
+    width = 2 * inter_quartile_range / (len(array) ** (1 / 3))
+    width = width.item()
+    return width
+
+
+def shimazaki_shinomoto(array):
+    """
+    Shimazaki and Shinomoto's choice. Biased variance is recommended.
+
+    :param array: sample array
+    :type array: np.ndarray
+
+    :return: bin-width
+    :rtype: float
+    """
+
+    def l2_risk_function(points):
+        lo = points.min()
+        hi = points.max()
+
+        n_min = lo + 1  # min number of splits
+        n_max = hi + 1  # max number of splits
+
+        counts = np.arange(n_min, n_max)  # number of bins vector
+        widths = (hi - lo) / counts  # width (size) of bins vector
+
+        costs = np.zeros((widths.size, 1))  # cost function values
+        for i in range(counts.size):
+            edges = np.linspace(lo, hi, counts[i] + 1)
[…]

### 2020-08-20 codebase/rlessential/visualization/aggregation_distribution.py
Visualizations added

@@ -1,0 +1,38 @@
+import matplotlib.pyplot as plt
+import numpy as np
+from scipy.stats import norm
+
+theme_blue = '#0C2B36'
+theme_red = '#E04D4F'
+theme_green = '#00F900'
+
+mu = 5
+sig = 1
+
+# Generate some data for this demonstration.
+data = norm.rvs(10.0, 2.5, size=500)
+
+# Fit a normal distribution to the data:
+mu, std = norm.fit(data)
+
+# Plot the histogram.
+plt.hist(data, bins=9, density=True, alpha=0.8, color=theme_blue, edgecolor='gray')
+
+# Plot the PDF.
+xmin, xmax = plt.xlim()
+x = np.linspace(xmin, xmax, 100)
+p = norm.pdf(x, mu, std)
+plt.plot(x, p, theme_red, linewidth=4)
+
+plt.gca().spines['left'].set_visible(False)
+plt.gca().spines['top'].set_visible(False)
+plt.gca().spines['bottom'].set_visible(False)
+plt.gca().spines['right'].set_visible(False)
+
+plt.xticks([], [])
+plt.yticks([], [])
+
+# plt.savefig('dist_initial.svg', format='svg')
+plt.savefig('dist_final.svg', format='svg')
+
+plt.show()

### 2020-08-20 codebase/rlessential/visualization/plot.py

@@ -1,0 +1,94 @@
+import matplotlib.pyplot as plt
+import numpy as np
+
+import consts
+
+# slides theme colors
+theme_blue = '#0C2B36'
+theme_red = '#E04D4F'
+theme_green = '#00F900'
+
+titles = {'original': 'Original reward', 'aggregate': 'Aggregate reward', 'adapted': 'Adapted aggregate reward'}
+
+
+def average_rolling(array):
+    """
+    Calculates moving(rolling) average.
+
+    :param array: input array
+    :return: calculated cumulative moving average
+    """
+    cumulative_sum = array.cumsum()
+    a = np.arange(array.size)
+    a = a + 1
+    rolling_mean = cumulative_sum / a
+    return rolling_mean
+
+
+def plot_steps(original, aggregate, adapted):
+    fig, ax = plt.subplots()
+    ax.set_xticks([], [])
+    ax.set_yticks([], [])
+
+    ax1 = fig.add_subplot(311)
+    ax1.plot(original, color=theme_blue, label=titles['original'], lw=2)
+    ax1.set_title(titles['original'])
+    ax1.set_xticks([], [])
+
+    ax2 = fig.add_subplot(312)
+    ax2.plot(aggregate, color=theme_red, label=titles['aggregate'], lw=2)
+    ax2.set_title(titles['aggregate'])
+    ax2.set_xticks([], [])
+    ax2.set_ylabel('rewards')
+
+    ax3 = fig.add_subplot(313)
+    ax3.plot(adapted, color=theme_green, label=titles['adapted'], lw=2)
+    ax3.set_title(titles['adapted'])
+
+    plt.xlabel('steps')
+    return plt
+
+
+def plot_cumulative(original, aggregate, adapted):
+    plt.figure()
+    plt.plot(original.cumsum(),
+             color=theme_blue, label=titles['original'], lw=2)
+    plt.plot(aggregate.cumsum(),
+             color=theme_red, label=titles['aggregate'], lw=2)
+    plt.plot(adapted.cumsum(),
+             color=theme_green, label=titles['adapted'], lw=2)
+    plt.title('Cumulative Reward')
+    plt.xlabel('steps')
+    plt.ylabel('rewards')
+    plt.legend()
+    return plt
+
+
+def plot_rolling(original, aggregate, adapted):
+    rolling_original = average_rolling(original)
+    rolling_aggregate = average_rolling(aggregate)
[…]

### 2020-08-20 codebase/rlessential/visualization/shimazaki_hitogram.py

@@ -7,6 +7,7 @@ from scipy import linspace
 # slides theme colors
 theme_blue = '#0C2B36'
 theme_red = '#E04D4F'
+theme_green = '#00FA92'
 
 x = normal(4, 1, 100)  # Generate n pseudo-random numbers with (mu,sigma,n)
 x = [4.37, 3.87, 4.00, 4.03, 3.50, 4.08, 2.25, 4.70, 1.73, 4.93, 1.73, 4.62, \

### 2020-08-20 presentation/GradTalk/abstract.tex
Graduate talk presentation added

% Preamble
\documentclass[11pt]{article}

% Packages
\usepackage{amsmath}
\usepackage{hyperref}

% sans-serif typeface
\renewcommand{\familydefault}{\sfdefault}

% hyperlink setup
\hypersetup{
    colorlinks=true,       % false: boxed links; true: colored links
    urlcolor=blue        % color of external links
}

% suppress page numbering
\pagenumbering{gobble}

% Document
\begin{document}
    \begin{flushleft}
        \LARGE \textbf{Distributional State Aggregation in Reinforcement Learning}
        \newline \break
        \Large \textbf{Shayan Amani}
        \newline \break
        \textit{MS Project Presentation}
        \newline \break

        Tuesday, August 25, 2020 at 10:00 AM EST \\
        via Zoom at \href{https://unh.zoom.us/j/98881910790}{https://unh.zoom.us/j/98881910790}

    \end{flushleft}
    \section*{Abstract}

    Despite the theoretical promises, real-world problems introduce computational challenges in the adaption of reinforcement learning algorithms.
    Such problems are often accompanied by an enormous state space, large enough to surpass the conventional memory and processing limits.
    State aggregation encompasses ideas and methods to downscale a state space with a bounded error to yield a compact instance of the original problem ultimately.
    Being around for decades, the recent advancement in state aggregation went in the shadow of approaches based on neural networks, though.
    As an analytically-transparent method, state aggregation still has superior advantages over incorporating neural networks. \\

    While a variety of approaches are available in the literature, the majority of the solutions are whether highly dependent on problem mechanics or make strong theoretical assumptions that render the results impractical.
    Universal state aggregation methods are deemed to be necessary regardless of the low- level assumptions and properties.
    This work proposes a class of non-parametric state aggregation methods merely based on the collected sample distribution.
    Inspired by histograms as the density estimator, a myriad of practices are investigated, then experiments are performed to scrutinize the core hypothesis.
    \newline \break
    \textbf{Advisor}: \\
    Marek Petrik

\end{document}

### 2020-08-20 presentation/GradTalk/main.tex
Graduate talk presentation added

@@
 
-\documentclass{beamer}
+\documentclass[mathserif]{beamer}
 \usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
+%\usepackage[T1]{fontenc}
 \usepackage{graphicx}
+\usepackage{amsmath}
+\usepackage{subfig}
 
+\usepackage[style=verbose,sorting=none, backend=bibtex]{biblatex}
+\addbibresource{library.bib}  % instead of \bibliography{library}
+
+% slide color theme
 \usetheme{Cuerna}
+% default, bluesimplex, lettuce, brick
 \usecolortheme{default}
-% default, bluesimplex, lettuce, brick
+
+%\usefonttheme[onlymath]{serif}
+
+% slide notes
+\usepackage{pgfpages}
+\usepackage{xcolor}
+\usepackage{hyperref}
+\setbeameroption{hide notes} % Only slides
+%\setbeameroption{show only notes} % Only notes
+%\setbeameroption{show notes on second screen=right} % Both
 
@@
 \usebeamerfont{footline}%
+
 \usebeamercolor[fg]{footline}%
@@
 
-\title{TITLE}
+\definecolor{navydarkblue}{rgb}{0.2156, 0.2784, 0.3647}
+\definecolor{maraschino}{rgb}{0.8784, 0.3019, 0.3098}
+
+\title{Distributional State Aggregation in Reinforcement Learning}
 \author{Shayan Amani}
 
-\date{DATE 2019}
+\date{August XX, 2020}
 \institute{Department of Computer Science, University of New Hampshire}
@@
     \begin{frame}
-\frametitle{They have said...}
+        \frametitle{whoami}
 
-\Huge{``Our virtues and our failings are \textbf{inseparable}, like force and matter. When they separate, \textbf{man is no more}.''}
+        \begin{columns}
+            \column{0.42\linewidth}
+            \centering
+            \includegraphics[width=1\columnwidth]{res/wordcloud.png}
 
-% \begin{rightflush}
-% Nikola Tesla
-% \end{rightflush}
-\hfill \LARGE{\textit{Nikola Tesla}}
-
-\end{frame}
-
-\begin{frame}
-    \frametitle{Who Am I?}
-
-    I...
+            \column{0.58\linewidth}
             \begin{itemize}
-        \item finished my bachelor's in electrical engineering in 2015
-        \item aspired to develop for my own ideas --at a certain point, I had 200k monthly active users on one of my applications--
-        \item started my Ph.D. program at UNH in fall 2017
-        \item have been through a journey to figure out what I want
-        \item have been experimenting with quite a few areas
-        \item ended up working with Prof. Marek Petrik on RL
+                \item Finished bachelor's in EE in spring 2015
+                \item Aspired to develop for my own ideas
+                \begin{itemize}
+                    \item at a certain point, I had 200k monthly active users
                 \end{itemize}
+                \item Started Ph.D.\ at UNH in fall 2017
+                \item Have been experimenting with quite a few areas to discover what I like to pursue
+            \end{itemize}
+        \end{columns}
 
@@
     \begin{frame}
-    \frametitle{What Have I Been Dealing With}
-    \includegraphics[width=1\columnwidth]{wordcloud.png}
+        \Huge \textbf{Introduction}
+    \end{frame}
+
+    \begin{frame}
+        \frametitle{Problem}
+
+        \includegraphics[width=0.9\columnwidth]{res/aggregation-distribution-before-after.pdf}
+
+        \begin{itemize}
+            \item Reduce model into low resolution clusters.
+            \item Same distribution, different discritization in state space.
+        \end{itemize}
+
+        \note{cluster the original states into the group of states, called aggregate states}
 
@@
     \begin{frame}
-    \frametitle{Outline}
+        \frametitle{State Aggregation}
+        \framesubtitle{Problem}
 
         \begin{itemize}
-        \item Problem
-        \item Examples
-        \item Motivation
-        \item Outline
-        \item BODY
-        \item Q&A
+            \item No general rule applicable to various aggregation problems.
+            \item Domain-dependent solutions.
+            \item feature-based aggregation and more engineering.
         \end{itemize}
@@
     \begin{frame}
-    \frametitle{State Aggregation}
-    \label{fm:sta}
-    \framesubtitle{Problem}
+        \frametitle{Example}
+
+        \center \includegraphics[scale=0.9]{res/manhattan-taxi.png}\let\thefootnote\relax\footcite{Duan2019}
+
+        \note{100 Taxi trips in the picture.}
 
         \begin{itemize}
-        \item Problem \textit{approximation} 
-        \item based on \textit{aggregation}.
-        \item in value space
-        \item feature-based parametric approximation
+            \item 11 million taxi trips in Manhattan, NYC.
+            \item Clustered into 1922 divisions.
         \end{itemize}
@@
     \begin{frame}
-    \frametitle{Why State Aggregation?}
-    \framesubtitle{Problem}
+        \frametitle{State Aggregation}
+        \framesubtitle{Motivation}
 
@@
         \end{itemize}
+    \end{frame}
+
+    \begin{frame}
+        \frametitle{State Aggregation}
+        \framesubtitle{Motivation}
+
+        {\Large \textbf{Steps toward an aggregation framework:}}
+
+        \begin{itemize}
+            \item Non-parametric
+            \item Domain-agnostic
+            \item Sample-based
+        \end{itemize}
+
+        \note{non-parametric: automatic}
 
@@
     \begin{frame}
-    \frametitle{Aggregate State}
-    \framesubtitle{Problem}
+        \frametitle{Outline}
 
         \begin{itemize}
-        \item special group of multiple [original] states
-        \item 
+            \item Introduction
+            \item Density Estimation
+            \item Histograms
+            \item Metrics
+            \item Methods
+            \begin{itemize}
+                \item Environment
+                \item Interval Count
+                \item Interval Width
+            \end{itemize}
+            \item Discussion
+            \item Future Work
+            \item Q\&A
+        \end{itemize}
+    \end{frame}
+
+    \begin{frame}
+        \Huge \textbf{Density Estimation}
+    \end{frame}
+
+    \begin{frame}
+        \frametitle{Density Estimators}
+
+        {\Large \textbf{Definition:}}
+
+        Density estimation is fitting an estimate, based on \textbf{observed data} of an unobservable
+        underlying \textbf{probability density function (PDF)}.
+
+        \vspace{0.3in}
+
+        {\Large \textbf{Estimators:}}
+        \begin{itemize}
+            \item histogram
+            \item kernel density estimation
+            \item wavelets thresholding
+            \item smoothing splines
+        \end{itemize}
+    \end{frame}
+
+    \begin{frame}
+        \frametitle{Histogram}
+        \framesubtitle{Density Estimators}
+
+        \begin{columns}
+            \column{0.60\linewidth}
+            \centering
+            \includegraphics[width=1\columnwidth]{res/histogram_pdf.pdf}
+
+            \column{0.40\linewidth}
+            $ \hat{f}(x; w) = \frac{\left|B_j\right|}{n w}, \quad x \in B_{j}$
+        \end{columns}
+
+    \end{frame}
+
+    \begin{frame}
+        \Huge \textbf{Histograms}
+    \end{frame}
+
+    \begin{frame}
+        \frametitle{Pros}
+        \framesubtitle{Histogram}
+
+        {\Large \textbf{Advantages:}}
+        \begin{itemize}
+            \item As density estimators, histograms have been studied thoroughly for decades.
+            \item Computational advantages compared to kernel-based methods.
+            \item Non-parametric estimation to avoid domain-specific feature engineering.
+        \end{itemize}
+
+        \note{2. compared to Kernel Density Estimators and other complex methods}
+
+    \end{frame}
+
+    \begin{frame}
+        \frametitle{Cons}
+        \framesubtitle{Histogram}
+
+        {\Large \textbf{Drawbacks:}}
+        \begin{itemize}
+            \item No universal optimality conditions on parameters ($k, w$) and asymptotic considerations.
+            \item Cost of being non-parametric: slow convergence rate.
+            \item Strong theoretical assumptions render theorems and results impractical.
+            \item Computational complexity in case of non-normal distributions.
+        \end{itemize}
+
+        \note{No Free Lunch\ldots this time in histogram bin width calculation}
+        \note{Strong \ldots theoretical assumptions [about the underlying distribution]}
+    \end{frame}
+
+    \begin{frame}
+        \frametitle{Goal}
+        \framesubtitle{Histogram}
+
+        {\Large \textbf{State aggregation based on histograms:}}
+        \begin{itemize}
+            \item Automatic and efficient method of choosing the number of bins.
+            \item Based on the characteristics of the underlying distribution.
+        \end{itemize}
+    \end{frame}
+
+    \vspace{0.3in}
+
+    \begin{frame}
+        \Huge \textbf{Metrics}
+    \end{frame}
+
+    \begin{frame}
+        \frametitle{Measure of Fit}
+        \framesubtitle{Histogram}
+
+        \begin{itemize}
+            \item \textbf{risk} or integrated mean squared error (\textbf{IMSE}).
+            \begin{itemize}
+                \item No direct solution to IMSE, underlying distribution is unknown
+            \end{itemize}
+            \item \textbf{estimated risk} or cross-validation estimator of risk.
+        \end{itemize}
+    \end{frame}
+
+    \begin{frame}
+        \frametitle{IMSE}
+        \framesubtitle{Measure of Fit}
+
+        \begin{equation}
+            \label{eq:histogram_mse}
+            \begin{split}
+                \textrm{MSE}\{\hat{f}(.; w)\} & = \mathbb{E}\left[\hat{f}(x)-f(x)\right]^{2} \\
+                & = \frac{1}{w k} \hat{f}(x)-\frac{1}{k} \hat{f}(x)^{2}+\left[\hat{f}(x)-f(x)\right]^{2} \\
+                & = \text{\textcolor{maraschino}{Variance}} + \text{\textcolor{navydarkblue}{Bias}}^{2}
+            \end{split}
+        \end{equation}
+
+        \begin{itemize}
+            \item The histogram converges in mean squared to $f(x)$ if $w \rightarrow 0$ and $nw \rightarrow \infty$.
+        \end{itemize}
+    \end{frame}
+
+    \begin{frame}
+        \frametitle{IMSE}
+        \framesubtitle{Measure of Fit}
+
+        \begin{equation}
+            \label{eq:histogram_imse}
+            \begin{split}
+                \textrm{IMSE}\{\hat{f}(.; w)\} = \int \mathbb{E}\{\hat{f}(x)-f(x)\}^{2} dx
+            \end{split}
+        \end{equation}
+
+        \begin{itemize}
+            \item Global error measure of a histogram estimate.
+            \item Slower convergence to a fixed-point than parametric estimators\footcite{Wasserman2004}: $n^{-2/3} > n^{-1}$
+        \end{itemize}
+
+        \note{slower: price of being non-parametric}
+
+    \end{frame}
+
+    \begin{frame}
+        \Huge \textbf{Metrics} \\
+        \qquad \LARGE \textbf{How many bins?}
+
+        \note{widely asked question in the literature.}
+    \end{frame}
+
+    \begin{frame}
+        \frametitle{Bias-Variance Tradeoff}
+
+        {\Large \textbf{Number of buckets of discritization:}}
+        \begin{itemize}
+            \item range split aggregation
+        \end{itemize}
+
+        \LARGE \textbf{\textcolor{maraschino}{Variance}}
+        \hfill
+        \LARGE \textbf{\textcolor{navydarkblue}{Bias}}
+
+        \includegraphics[width=1\columnwidth]{res/slider.pdf}
+
+        \LARGE \textbf{\textcolor{maraschino}{Bin-count}}
+        \hfill
+        \LARGE \textbf{\textcolor{navydarkblue}{Bin-width}}
+
+        \center \LARGE $\Longleftrightarrow$
+
+    \end{frame}
+
+    \begin{frame}
+        \frametitle{Smoothing}
+        \framesubtitle{Bias-Variance Tradeoff}
+
+        \begin{center}
+            \includegraphics[scale=0.55]{res/over-under-smoothed.pdf}
+        \end{center}
+
+    \end{frame}
+
+    \begin{frame}
+        \frametitle{Optimal Smoothing}
+        \framesubtitle{Bias-Variance Tradeoff}
+
+        \includegraphics[width=1\columnwidth]{res/bias-variance-smoothing.png}\let\thefootnote\relax\footcite{Wasserman2004}
+    \end{frame}
+
+    \begin{frame}
+        \frametitle{Bin Count vs. Bin Width}
+
+        For equally-spaced intervals:
+
+        \begin{equation}
+            \label{eq:ordinary_count_width}
+            k=\left\lceil\frac{\max(x)-\min (x)}{w}\right\rceil
+        \end{equation}
+
+        \begin{itemize}
+            \item depends on the distribution range and interval size: $k = f(R, w)$
+
+        \end{itemize}
+    \end{frame}
+
+    \begin{frame}
+        \Huge \textbf{Methods}
+    \end{frame}
+
+    \begin{frame}
+        \frametitle{Environment}
+        \framesubtitle{Method}
+
+        \large {\textbf{Machine Replacement}}
+
+        \center \includegraphics[scale=0.25]{res/machine-replacement-mdp.pdf}\let\thefootnote\relax\footcite{Delage2010}
+
+    \end{frame}
+
+    \begin{frame}
+        \Huge \textbf{Methods} \\
+        \qquad \LARGE \textbf{Interval Count Methods}
+    \end{frame}
+
+    \begin{frame}
+        \frametitle{Square-root Rule}
+        \framesubtitle{Interval Count}
+
+        \begin{equation}
+            \label{eq:square_root_method_width}
+            w = \frac{\max(x) - \min(x)}{\sqrt{n}}
+        \end{equation}
+
+        \begin{equation}
+            \label{eq:square_root_method_count}
+            k = \lceil \sqrt{n} \rceil
+        \end{equation}
+
+        \begin{itemize}
+            \item Intuitive: no restricting assumptions
+            \item Used by Excel\footcite{ExcelHistogram}.
+        \end{itemize}
+
+        \note{rounds up to the next integer}
+
+    \end{frame}
+
+    \begin{frame}
+        \frametitle{Aggregation}
+        \framesubtitle{Square-root Rule}
+
+        \begin{figure}
+            \label{fig:square-root-rewards}
+            \centering
+            \includegraphics[width=0.46\columnwidth]{res/experiments/sqrt_steps_rewards.pdf}
+            \qquad
+            \includegraphics[width=0.46\columnwidth]{res/experiments/sqrt_rolling_rewards.pdf}
+        \end{figure}
+
+    \end{frame}
+
+    \begin{frame}
+        \frametitle{Sturge's Rule}
+        \framesubtitle{Interval Count}
+
+        \begin{equation}
+            \label{eq:sturge_method_width}
+            w = \frac{\max(x) - \min(x)}{1 + \lg{n}}
+        \end{equation}
+
+        \begin{equation}
+            \label{eq:sturge_method_count}
+            k = 1 + \lceil \lg{n} \rceil
+        \end{equation}
+
+        \begin{itemize}
+            \item Estimates the original distribution with a series of binomial coefficients\footcite{Sturges1926}.
+            \item Normal distribution is implied.
+        \end{itemize}
+    \end{frame}
+
+    \begin{frame}
+        \frametitle{Aggregation}
+        \framesubtitle{Sturge's Formula}
+
+        \begin{figure}
+            \label{fig:sturge-rewards}
+            \centering
+            \includegraphics[width=0.46\columnwidth]{res/experiments/sturge_steps_rewards.pdf}
+            \qquad
+            \includegraphics[width=0.46\columnwidth]{res/experiments/sturge_rolling_rewards.pdf}
+        \end{figure}
+
+    \end{frame}
+
+    \begin{frame}
+        \frametitle{Rice Rule}
+        \framesubtitle{Interval Count}
+
+        \begin{equation}
+            \label{eq:rice_method_width}
+            w = \frac{\max(x) - \min(x)}{2 \sqrt[3]{n}}
+        \end{equation}
+
+        \begin{equation}
+            \label{eq:rice_method_count}
+            k = \lceil 2 \sqrt[3]{n} \rceil
+        \end{equation}
+
+        \begin{itemize}
+            \item The intuitive successor to Sturges' rule.
+            \item Recommended by trial and errors\footcite{Lane2003}.
+        \end{itemize}
+    \end{frame}
+
+    \begin{frame}
+        \frametitle{Aggregation}
+        \framesubtitle{Rice Rule}
+
+        \begin{figure}
+            \label{fig:rice-rewards}
+            \centering
+            \includegraphics[width=0.46\columnwidth]{res/experiments/rice_steps_rewards.pdf}
+            \qquad
+            \includegraphics[width=0.46\columnwidth]{res/experiments/rice_rolling_rewards.pdf}
+        \end{figure}
+
+    \end{frame}
+
+    \begin{frame}
+        \frametitle{Doane's Rule}
+        \framesubtitle{Interval Count}
+
+        \[
+            k = \underbrace{1 + \textcolor{maraschino}{\lceil} \lg n\textcolor{maraschino}{\rceil}}_{\text{Sturges' classes}} + \underbrace{ \textcolor{maraschino}{\lceil} \lg \left(1 +
+            \frac{\textcolor{maraschino}{|}\sqrt{\beta_{1}}\textcolor{maraschino}{|}}
+            {\sigma_{\sqrt{\beta_{1}}}}\right)\textcolor{maraschino}{\rceil}}_{\text{Doane's extended classes}}
+        \]
+
+        \note{beta distribution}
+
+        \begin{equation}
+            \label{eq:fisher_pearson_coefficient}
+            \sqrt{\beta_{1}} = \frac{m_{3}}{m_{2}^{3/2}}
+        \end{equation}
+
+        \begin{equation}
+            \label{eq:pearson_moments}
+            m_{2} = \frac{1}{n} \sum_{i=1}^{n}\left(X_{i}-\bar{X}\right)^{2} ,
+            m_{3}=\frac{1}{n} \sum_{i=1}^{n}\left(X_{i}-\bar{X}\right)^{3}
+        \end{equation}
+
+        \begin{equation}
+            \label{eq:pearson_estimated}
+            \sigma_{\sqrt{\beta_{1}}} = \sqrt{\frac{6(n-2)}{(n+1)(n+3)}}
+        \end{equation}
+    \end{frame}
+
+    \begin{frame}
+        \frametitle{Doane's Formula}
+        \framesubtitle{Interval Count}
+
+        \begin{itemize}
+            \item Modified Sturges' formula to reflect distribution characteristics\footcite{Doane1976}.
+            \item Hence, performs better in case of skewed distributions\footcite{Doane2011}.
+            \item Inspired by coding in the information theory: increasing entropy by introducing more symbols for
+            coding.
+        \end{itemize}
+
+        \note{negative value that \sqrt{\beta_{1}} might take and the variable name.}
+        \note{skewed distribution = non-normal distribution}
+    \end{frame}
+
+    \begin{frame}
+        \frametitle{Aggregation}
+        \framesubtitle{Doane's Formula}
+
+        \begin{figure}
+            \label{fig:doane-rewards}
+            \centering
+            \includegraphics[width=0.46\columnwidth]{res/experiments/doane_steps_rewards.pdf}
+            \qquad
+            \includegraphics[width=0.46\columnwidth]{res/experiments/doane_rolling_rewards.pdf}
+        \end{figure}
+
+    \end{frame}
+
+    \begin{frame}
+        \Huge \textbf{Methods} \\
+        \qquad \LARGE \textbf{Interval Width Methods}
+    \end{frame}
+
+    \begin{frame}
+        \frametitle{Scott's Formula}
+        \framesubtitle{Interval Width}
+
+        \begin{equation}
+            \label{eq:scott_method_imse}
+            \textrm{IMSE} = \int E\{\hat{f}(x)-f(x)\}^{2} dx
+        \end{equation}
+
+        using Taylor expansion:
+        \begin{equation}
+            \label{eq:scott_method_imse_taylor}
+            \textrm{IMSE} = 1 /\left(n w\right)+\frac{1}{12} w^{2} \int_{-\infty}^{\infty} f^{\prime}(x)^{2} dx +
+            O\left(1 /n + w^{3}\right)
+        \end{equation}
+
+        \begin{equation}
+            \label{eq:scott_method_optimal_width}
+            w^{*} = \left\{6 \middle/ \int_{-\infty}^{\infty} f^{\prime}(x)^{2} d x\right\}^{1 / 3}
+            n^{-1 / 3}
+        \end{equation}
+
+        \begin{equation}
+            \label{eq:scott_method_optimal_width_gaussian}
+            \begin{split}
+                w^{*} & = 2 \sqrt[6]{9 \pi} \: \sigma \: n^{-1 / 3} \\
+                & = 3.49083 \: \sigma \: n^{-1 / 3}
+            \end{split}
+        \end{equation}
+
+    \end{frame}
+
+    \begin{frame}
+        \frametitle{Scott's Formula}
+        \framesubtitle{Interval Width}
+
+        \begin{itemize}
+            \item Calculates the $w^{*}$ by minimizing \textrm{IMSE}\footcite{Scott1979}.
+            \item Derived the estimate for a Gaussian distribution.
+        \end{itemize}
+    \end{frame}
+
+    \begin{frame}
+        \frametitle{Aggregation}
+        \framesubtitle{Scott's Formula}
+
+        \begin{figure}
+            \label{fig:scott-rewards}
+            \centering
+            \includegraphics[width=0.46\columnwidth]{res/experiments/scott_steps_rewards.pdf}
+            \qquad
+            \includegraphics[width=0.46\columnwidth]{res/experiments/scott_rolling_rewards.pdf}
+        \end{figure}
+
+    \end{frame}
+
+    \begin{frame}
+        \frametitle{Freedman-Diaconis' Rule}
+        \framesubtitle{Interval Width}
+
+        \begin{equation}
+            \label{eq:freedman-diaconis-rule}
+            w = 2 \: \textrm{IQR}(x) \: n^{-1/3}
+        \end{equation}
+
+        \begin{itemize}
+            \item $3.5 \: \sigma \approx 2 \: \textrm{IQR}$
+            \item More robustness to outliers than standard deviation
+        \end{itemize}
+    \end{frame}
+
+    \begin{frame}
+        \frametitle{IQR}
+        \framesubtitle{Freedman-Diaconis' Rule}
+
+        \let\thefootnote\relax\footnote{Jhguch $|$ \href{https://creativecommons.org/licenses/by-sa/2.5}{CC BY-SA}}
+
+        \begin{figure}
+            \label{fig:iqr}
+            \includegraphics[scale=0.5]{res/Boxplot_vs_PDF.pdf}
+        \end{figure}
+
+    \end{frame}
+
+    \begin{frame}
+        \frametitle{Aggregation}
+        \framesubtitle{Freedman-Diaconis' Rule}
+
+        \begin{figure}
+            \label{fig:freedman-diaconis-rewards}
+            \centering
+            \includegraphics[width=0.46\columnwidth]{res/experiments/freedman-diaconis_steps_rewards.pdf}
+            \qquad
+            \includegraphics[width=0.46\columnwidth]{res/experiments/freedman-diaconis_rolling_rewards.pdf}
+        \end{figure}
+
+    \end{frame}
+
+    \begin{frame}
+        \frametitle{Shimazaki-Shinomoto's Choice}
+        \framesubtitle{Interval Width}
+
+        \begin{equation}
+            \label{eq:shimazaki-rule}
+            w^{*} = \operatorname*{arg\,min}_{w} C(w) = \operatorname*{arg\,min}_{w} \frac{2 \bar{X} - \sigma^{2}}{w^{2}}
+        \end{equation}
+
+        \[
+            \bar{X} = \frac{1}{n} \sum_{i=1}^{n} X_{i}; \: \underbrace{\sigma^{2} = \frac{1}{n} \sum_{i=1}^{n} (X_{i} - \bar{X})^{2}}_{\text{biased variance}}
+        \]
+
+        \begin{itemize}
+            \item Minimizes cost function for a set of proposed bin-widths\footcite{Shimazaki2007}.
+            \item Looks over the dispersion of data points count falling into bins.
         \end{itemize}
@@
     \begin{frame}
-    \frametitle{Nomenclature}
-    \framesubtitle{Problem}
+        \frametitle{Cost Analysis}
+        \framesubtitle{Shimazaki-Shinomoto's Choice}
 
-    \begin{itemize}
-        \item Original State
-            \begin{itemize}
-                \item 
-            \end{itemize}
-        \item Aggregate State
-    \end{itemize}
+        \begin{figure}
+            \label{fig:shimazaki-cost-histogram}
+            \centering
+            \subfloat{{\includegraphics[width=0.46\columnwidth]{res/cost_shimazaki.pdf} }}
+            \qquad
+            \subfloat{{\includegraphics[width=0.46\columnwidth]{res/histogram_shimazaki.pdf} }}
+        \end{figure}
 
@@
     \begin{frame}
-    \frametitle{Idea}
+        \frametitle{Aggregation}
+        \framesubtitle{Shimazaki-Shinomoto's Choice}
+
+        \begin{figure}
+            \label{fig:shimazaki-rewards}
+            \centering
+            \includegraphics[width=0.46\columnwidth]{res/experiments/shimazaki-shinomoto_steps_rewards.pdf}
+            \qquad
+            \includegraphics[width=0.46\columnwidth]{res/experiments/shimazaki-shinomoto_rolling_rewards.pdf}
+        \end{figure}
 
@@
     \begin{frame}
-    \frametitle{How Many Bins?}
-    \framesubtitle{Idea}
+        \frametitle{Discussion}
 
-    Number of buckets to split original states into:
         \begin{itemize}
-        \item range split aggregation
+            \item Incorporating more characteristics of underlying distribution leads to a better aggregation policy.
+            \item High dimensional spaces may take more advantage of this approach.
+            \item Sampling distribution decently explains the candidacy of states for aggregation.
         \end{itemize}
-
     \end{frame}
@@
     \begin{frame}
-    \frametitle{}
-
+        \frametitle{Ideas for Future}
+        \begin{enumerate}
+            \item Adjusted Fisher-Pearson estimation of skewness to take sample size into
+            account.
+            \item Employ studentized test instead of kurtosis in Doane's formula\footcite{Geary1936},~\footcite{Tracy2005}.
+            \item Variable bin-width aggregation.
+        \end{enumerate}
     \end{frame}
@@
     \begin{frame}
-    \frametitle{}
-
+        \Huge \textbf{Questions}
     \end{frame}
@@
     \begin{frame}
-    \frametitle{}
+        \frametitle{They have said...}
 
+        \Huge{``Our virtues and our failings are \textbf{inseparable}, like force and matter. \textbf{When they
+        separate, man is no more}.''}
+
+        \hfill \LARGE{\textit{Nikola Tesla}}
     \end{frame}
 
-\begin{frame}
-    \frametitle{}
-
-\end{frame}
-
-\begin{frame}
-    \frametitle{}
-
-\end{frame}
-
-\begin{frame}
-    \frametitle{}
-
+    \begin{frame}[allowframebreaks]
+        \printbibliography
     \end{frame}

### 2020-08-20 minutia/icml/workshop.tex
Aggregation paper added

@@
 \twocolumn[
-\icmltitle{Fast Uniform [Distributional] State Aggregation \\ in Reinforcement Learning}
+\icmltitle{ Density-based [Distributional] State Aggregation \\ in Reinforcement Learning}
 
@@
 
-    Our contribution is a bifurcation.
+    Compared to other kernel-based density estimators such as \citep{HongLim2019}, histograms steer clear of feature
+    selection step which is more of an experimental process than a framework of rules to follow.
+    Our histogram-based state abstraction model yield a state space model by collecting samples in a domain-agnostic
+    way.
+
+    Our contribution is a two-fold.
     We first introduce an algorithm for aggregation in MDPs, then we move on to show advantages of taking such
@@
 
-    Finding the bin width for a given distribution could be calculated by Freedman-Diaconis rule
+    \subsection{Bin width}
+    Finding the bin width for a given distribution could be calculated by Freedman-Diaconis's choice
     \cite{Freedman1981}.
@@
     where n is number of observations and $IQR(x)$ is the inter-quartile range of data.
-    This segmentation rule exhibits robustness to sample distribution due to low sensitivity of IQR to outliers and
+    The segmentation rule exhibits robustness to sample distribution due to low sensitivity of IQR to outliers and
     heteroscedasticity in the given distribution compared to conventional dispersion measures such as variance or
@@
     By having the inter-quartile range of data for each feature, a discretization policy can be calculated.
-
-    For the sake of clarity, we elaborate on some preliminary definitions here to have a consistent set of notations
-    to the end of this paper. We use model-free LSPI and LSTDQ \citep{Lagoudakis2003} to update and
-    evaluate the policy of actions, respectively.
 
